### 2023-01-27 Buffer_CBD.qmd
Commit.

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.28.2-Firenze">
  <identifier></identifier>
  <parentidentifier></parentidentifier>
  <language></language>
  <type>dataset</type>
  <title></title>
  <abstract></abstract>
  <links/>
  <fees></fees>
  <encoding></encoding>
  <crs>
    <spatialrefsys nativeFormat="Wkt">
      <wkt>PROJCRS["ETRS89 / UTM zone 33N",BASEGEOGCRS["ETRS89",ENSEMBLE["European Terrestrial Reference System 1989 ensemble",MEMBER["European Terrestrial Reference Frame 1989"],MEMBER["European Terrestrial Reference Frame 1990"],MEMBER["European Terrestrial Reference Frame 1991"],MEMBER["European Terrestrial Reference Frame 1992"],MEMBER["European Terrestrial Reference Frame 1993"],MEMBER["European Terrestrial Reference Frame 1994"],MEMBER["European Terrestrial Reference Frame 1996"],MEMBER["European Terrestrial Reference Frame 1997"],MEMBER["European Terrestrial Reference Frame 2000"],MEMBER["European Terrestrial Reference Frame 2005"],MEMBER["European Terrestrial Reference Frame 2014"],ELLIPSOID["GRS 1980",6378137,298.257222101,LENGTHUNIT["metre",1]],ENSEMBLEACCURACY[0.1]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],ID["EPSG",4258]],CONVERSION["UTM zone 33N",METHOD["Transverse Mercator",ID["EPSG",9807]],PARAMETER["Latitude of natural origin",0,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8801]],PARAMETER["Longitude of natural origin",15,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8802]],PARAMETER["Scale factor at natural origin",0.9996,SCALEUNIT["unity",1],ID["EPSG",8805]],PARAMETER["False easting",500000,LENGTHUNIT["metre",1],ID["EPSG",8806]],PARAMETER["False northing",0,LENGTHUNIT["metre",1],ID["EPSG",8807]]],CS[Cartesian,2],AXIS["(E)",east,ORDER[1],LENGTHUNIT["metre",1]],AXIS["(N)",north,ORDER[2],LENGTHUNIT["metre",1]],USAGE[SCOPE["Engineering survey, topographic mapping."],AREA["Europe between 12°E and 18°E: Austria; Denmark - offshore and offshore; Germany - onshore and offshore; Norway including Svalbard - onshore and offshore."],BBOX[46.4,12,84.42,18]],ID["EPSG",25833]]</wkt>
      <proj4>+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs</proj4>
      <srsid>2106</srsid>
      <srid>25833</srid>
      <authid>EPSG:25833</authid>
      <description>ETRS89 / UTM zone 33N</description>
      <projectionacronym>utm</projectionacronym>
      <ellipsoidacronym>EPSG:7019</ellipsoidacronym>
      <geographicflag>false</geographicflag>
    </spatialrefsys>
  </crs>
  <extent/>
</qgis>

### 2023-01-27 Dist_Mal_Dentist.qmd
Commit.

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.28.2-Firenze">
  <identifier></identifier>
  <parentidentifier></parentidentifier>
  <language></language>
  <type>dataset</type>
  <title></title>
  <abstract></abstract>
  <contact>
    <name></name>
    <organization></organization>
    <position></position>
    <voice></voice>
    <fax></fax>
    <email></email>
    <role></role>
  </contact>
  <links/>
  <fees></fees>
  <encoding></encoding>
  <crs>
    <spatialrefsys nativeFormat="Wkt">
      <wkt>PROJCRS["ETRS89 / UTM zone 33N",BASEGEOGCRS["ETRS89",ENSEMBLE["European Terrestrial Reference System 1989 ensemble",MEMBER["European Terrestrial Reference Frame 1989"],MEMBER["European Terrestrial Reference Frame 1990"],MEMBER["European Terrestrial Reference Frame 1991"],MEMBER["European Terrestrial Reference Frame 1992"],MEMBER["European Terrestrial Reference Frame 1993"],MEMBER["European Terrestrial Reference Frame 1994"],MEMBER["European Terrestrial Reference Frame 1996"],MEMBER["European Terrestrial Reference Frame 1997"],MEMBER["European Terrestrial Reference Frame 2000"],MEMBER["European Terrestrial Reference Frame 2005"],MEMBER["European Terrestrial Reference Frame 2014"],ELLIPSOID["GRS 1980",6378137,298.257222101,LENGTHUNIT["metre",1]],ENSEMBLEACCURACY[0.1]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],ID["EPSG",4258]],CONVERSION["UTM zone 33N",METHOD["Transverse Mercator",ID["EPSG",9807]],PARAMETER["Latitude of natural origin",0,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8801]],PARAMETER["Longitude of natural origin",15,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8802]],PARAMETER["Scale factor at natural origin",0.9996,SCALEUNIT["unity",1],ID["EPSG",8805]],PARAMETER["False easting",500000,LENGTHUNIT["metre",1],ID["EPSG",8806]],PARAMETER["False northing",0,LENGTHUNIT["metre",1],ID["EPSG",8807]]],CS[Cartesian,2],AXIS["(E)",east,ORDER[1],LENGTHUNIT["metre",1]],AXIS["(N)",north,ORDER[2],LENGTHUNIT["metre",1]],USAGE[SCOPE["Engineering survey, topographic mapping."],AREA["Europe between 12°E and 18°E: Austria; Denmark - offshore and offshore; Germany - onshore and offshore; Norway including Svalbard - onshore and offshore."],BBOX[46.4,12,84.42,18]],ID["EPSG",25833]]</wkt>
      <proj4>+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs</proj4>
      <srsid>2106</srsid>
      <srid>25833</srid>
      <authid>EPSG:25833</authid>
      <description>ETRS89 / UTM zone 33N</description>
      <projectionacronym>utm</projectionacronym>
      <ellipsoidacronym>EPSG:7019</ellipsoidacronym>
      <geographicflag>false</geographicflag>
    </spatialrefsys>
  </crs>
  <extent>
    <spatial maxx="0" crs="EPSG:25833" dimensions="2" minz="0" maxy="0" minx="0" miny="0" maxz="0"/>
    <temporal>
      <period>
        <start></start>
        <end></end>
      </period>
    </temporal>
  </extent>
</qgis>

### 2023-01-27 MSB106_Assignment.qmd
Commit.

---
title: "MSB106_Assignment"
author: "Sindre H. Øveraas, Alen Colakovic, Mona Lisa Jones & Sebastian M. Fløysand"
format:
 pdf:
  fontsize: "10"
  documentclass: article
  margin-left: 2cm
  margin-right: 2cm
papersize: A4
editor: visual
---

```{r, message=FALSE}
library(rgdal)
library(dplyr)
library(RSQLite)
library(sf)
library(tidyverse)
library(readr)
```

```{r, message=FALSE}
NOR_CBD <- read_csv("NOR_CBD.csv")
Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
Buffer_CBD <- read_csv("Buffer_CBD.csv")
Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
Buffer_den <- read.csv("Buffer_den.csv")
NOR_KOMM <- read_csv("NOR_KOMM.csv")
NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
```

```{r}
Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning)
```

```{r}

```

## Theoretical Foundation of Hypothesis

## Data description

## Econometric approach

### 2023-01-27 NOR_OSM_SHOP_MAL.qmd
Commit.

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.28.2-Firenze">
  <identifier></identifier>
  <parentidentifier></parentidentifier>
  <language></language>
  <type>dataset</type>
  <title></title>
  <abstract></abstract>
  <contact>
    <name></name>
    <organization></organization>
    <position></position>
    <voice></voice>
    <fax></fax>
    <email></email>
    <role></role>
  </contact>
  <links/>
  <fees></fees>
  <encoding></encoding>
  <crs>
    <spatialrefsys nativeFormat="Wkt">
      <wkt>PROJCRS["Slovenia 1996 / UTM zone 33N",BASEGEOGCRS["Slovenia 1996",DATUM["Slovenia Geodetic Datum 1996",ELLIPSOID["GRS 1980",6378137,298.257222101,LENGTHUNIT["metre",1]]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],ID["EPSG",4765]],CONVERSION["UTM zone 33N",METHOD["Transverse Mercator",ID["EPSG",9807]],PARAMETER["Latitude of natural origin",0,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8801]],PARAMETER["Longitude of natural origin",15,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8802]],PARAMETER["Scale factor at natural origin",0.9996,SCALEUNIT["unity",1],ID["EPSG",8805]],PARAMETER["False easting",500000,LENGTHUNIT["metre",1],ID["EPSG",8806]],PARAMETER["False northing",0,LENGTHUNIT["metre",1],ID["EPSG",8807]]],CS[Cartesian,2],AXIS["(E)",east,ORDER[1],LENGTHUNIT["metre",1]],AXIS["(N)",north,ORDER[2],LENGTHUNIT["metre",1]],USAGE[SCOPE["Military survey."],AREA["Slovenia - onshore and offshore."],BBOX[45.42,13.38,46.88,16.61]],ID["EPSG",8687]]</wkt>
      <proj4>+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs</proj4>
      <srsid>60321</srsid>
      <srid>520000321</srid>
      <authid>EPSG:8687</authid>
      <description>Slovenia 1996 / UTM zone 33N</description>
      <projectionacronym>utm</projectionacronym>
      <ellipsoidacronym>EPSG:7019</ellipsoidacronym>
      <geographicflag>false</geographicflag>
    </spatialrefsys>
  </crs>
  <extent>
    <spatial maxx="0" crs="EPSG:8687" dimensions="2" minz="0" maxy="0" minx="0" miny="0" maxz="0"/>
    <temporal>
      <period>
        <start></start>
        <end></end>
      </period>
    </temporal>
  </extent>
</qgis>

### 2023-01-27 regnskapstall_Tann.qmd
Commit.

<!DOCTYPE qgis PUBLIC 'http://mrcc.com/qgis.dtd' 'SYSTEM'>
<qgis version="3.28.2-Firenze">
  <identifier></identifier>
  <parentidentifier></parentidentifier>
  <language></language>
  <type>dataset</type>
  <title></title>
  <abstract></abstract>
  <contact>
    <name></name>
    <organization></organization>
    <position></position>
    <voice></voice>
    <fax></fax>
    <email></email>
    <role></role>
  </contact>
  <links/>
  <fees></fees>
  <encoding></encoding>
  <crs>
    <spatialrefsys nativeFormat="Wkt">
      <wkt>PROJCRS["Slovenia 1996 / UTM zone 33N",BASEGEOGCRS["Slovenia 1996",DATUM["Slovenia Geodetic Datum 1996",ELLIPSOID["GRS 1980",6378137,298.257222101,LENGTHUNIT["metre",1]]],PRIMEM["Greenwich",0,ANGLEUNIT["degree",0.0174532925199433]],ID["EPSG",4765]],CONVERSION["UTM zone 33N",METHOD["Transverse Mercator",ID["EPSG",9807]],PARAMETER["Latitude of natural origin",0,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8801]],PARAMETER["Longitude of natural origin",15,ANGLEUNIT["degree",0.0174532925199433],ID["EPSG",8802]],PARAMETER["Scale factor at natural origin",0.9996,SCALEUNIT["unity",1],ID["EPSG",8805]],PARAMETER["False easting",500000,LENGTHUNIT["metre",1],ID["EPSG",8806]],PARAMETER["False northing",0,LENGTHUNIT["metre",1],ID["EPSG",8807]]],CS[Cartesian,2],AXIS["(E)",east,ORDER[1],LENGTHUNIT["metre",1]],AXIS["(N)",north,ORDER[2],LENGTHUNIT["metre",1]],USAGE[SCOPE["Military survey."],AREA["Slovenia - onshore and offshore."],BBOX[45.42,13.38,46.88,16.61]],ID["EPSG",8687]]</wkt>
      <proj4>+proj=utm +zone=33 +ellps=GRS80 +towgs84=0,0,0,0,0,0,0 +units=m +no_defs</proj4>
      <srsid>60321</srsid>
      <srid>520000321</srid>
      <authid>EPSG:8687</authid>
      <description>Slovenia 1996 / UTM zone 33N</description>
      <projectionacronym>utm</projectionacronym>
      <ellipsoidacronym>EPSG:7019</ellipsoidacronym>
      <geographicflag>false</geographicflag>
    </spatialrefsys>
  </crs>
  <extent>
    <spatial maxx="0" crs="" dimensions="2" minz="0" maxy="0" minx="0" miny="0" maxz="0"/>
    <temporal>
      <period>
        <start></start>
        <end></end>
      </period>
    </temporal>
  </extent>
</qgis>

### 2023-01-30 MSB106_Assignment.qmd
Commit.

@@
 ```{r}
+Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
+  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y)
+```
 
-```
+## Introduction
+
+Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
+More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
+As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
 

### 2023-01-31 Retail Agglomeration.R
Utkast og stuktur

@@ -1,0 +1,153 @@
+---
+title: "Regional and Urban Economics"
+subtitle: Retail Agglomeration
+author: "Sindre Halvorsen Øveraas , Sebastian Mena Fløysand, Alen Colakovic & Mona Lisa Jones"
+format: pdf
+  html: default
+  pdf: default
+editor: visual
+execute: 
+  echo: true
+---
+
+## Abstract
+
+## Introduction
+
+**Hypothesis**
+
+The location choice of firms providing service to end
+consumers, significantly determine their ultimate growth potential.
+
+**Research question**
+
+What is the most profitable location for dental practice in
+years to come in the Haugaland-region?
+
+## Literature Review
+
+#### Shopping Mall literature
+
+#### 
+Theoretical Framework
+
+**Central place Theory**
+
+**Retail Agglomeration and Economies of Shopping Centres**
+
+Town and cities have long had shopping districts in which stores are concentrated.
+To some extent such centre points were the result of uncoordinated store-location decisions (Brueckner, 2011).
+Nevertheless, not coincidentally as the concentration of retail stores can be explained by a market desire to create a Nash equilibrium.
+At this point all retailers expected outcome cannot be improved by changing position (Hotelling 1929).
+Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
+Shopping malls are often located
+outside the city centre.
+Sometimes they challenge the existence of the city centre.
+In extraordinary circumstances they can be seeds for new urban centres.
+Not so hard to understand taken into consideration that shopping malls provide most goods and services given downtown.
+Not to mention, the convenience of free parking, longer opening hours and easy road access.
+Especially, if the malls
+cater for cinemas, restaurants, bars and other attributes in conjunction with downtown.
+Occasionally, perceived as \"predatory malls\" which are purposely built to take over the local retail market.
+Owners of shopping malls \"orchestrate\" the process of retail agglomeration that happens naturally in towns and cities (Brueckner, 2011).
+The agglomeration in centre points contributes several benefits for the consumers.
+First of all, they save time and money.
+When the consumers visit multiple stores to make a variety of purchases, the cost of the overall shopping trip is subsequently reduced (Brueckner, 2011).
+Secondly, comparison shopping as consumers prefer to compare products to achieve a better purchase decision. 
+Again, resulting in clusters in order to achieve higher shopping benefits than isolated stores.
+Consequently, leading to more customer traffic and increased price competition between stores selling
+similar products (Brueckner, 2011).
[…]

### 2023-01-31 Retail Agglomeration.R
commiting update

@@ -104,14 +104,11 @@ In order to construct the model, the following assumptions are taken into accoun
 
 The adopted bid-rent model includes one major difference, whereas the original model the land and non-land production factors relationships are fixes, land and non-land production factors are assumed to be mutually substitutable inputs, irrespective of whether the firm produces agriculture or a manufactured good (McCann, 2013).
 
-Bilde av modellen
-
 Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues.
-The rent gradient is a negatively sloped straight line and has an intercept with
-the vertical axis of maximum value of rent.
+The rent gradient is a negatively sloped straight line and has an intercept with the vertical axis of maximum value of rent.
 Furthermore, it has an intercept with the horizontal axis whereas land value is nil.
 The slope is reflecting a trade-off between transport cost and rent.
- The llustration shows three categories: retail-, manufacturing- and residential units, each serving a specific purpose.
+The llustration shows three categories: retail-, manufacturing- and residential units, each serving a specific purpose.
 The rent realized by the landowner is the envelope of the three rent supply curves.
 
 **Analysis**

### 2023-01-31 Retail Agglomeration.R
ingjen

@@ -26,7 +26,7 @@ years to come in the Haugaland-region?
 
 ## Literature Review
 
-#### Shopping Mall literature
+#### Previous literature on Shopping malls
 
 #### 
 Theoretical Framework
@@ -77,7 +77,7 @@ The bid-rent model has since been associated with the works of Alsonso (1964,197
 Further development has been implemented by authors such as Mills (1969, 1970), Muth (1969), and Evans (1973).
 Ensuing the model can be applied in a broader setting than the original Thunen model (McCann, 2013). 
 
-**The Bid-rent model**
+**Assumtions**
 
 The bid rent model is a simple theory to explain urban structure.
 In order to construct the model, the following assumptions are taken into account.
@@ -104,6 +104,8 @@ In order to construct the model, the following assumptions are taken into accoun
 
 The adopted bid-rent model includes one major difference, whereas the original model the land and non-land production factors relationships are fixes, land and non-land production factors are assumed to be mutually substitutable inputs, irrespective of whether the firm produces agriculture or a manufactured good (McCann, 2013).
 
+*(Bilde / illustasjon av mdellen)*
+
 Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues.
 The rent gradient is a negatively sloped straight line and has an intercept with the vertical axis of maximum value of rent.
 Furthermore, it has an intercept with the horizontal axis whereas land value is nil.
@@ -111,7 +113,7 @@ The slope is reflecting a trade-off between transport cost and rent.
 The llustration shows three categories: retail-, manufacturing- and residential units, each serving a specific purpose.
 The rent realized by the landowner is the envelope of the three rent supply curves.
 
-**Analysis**
+**Interpretation of the model**
 
 A more central location implies a willingness to pay higher rents due to lower transport costs and higher revenues.
 Likewise, a suburban location yields the same profit if and only if less rent is paid for land.

### 2023-01-31 MSB106_Assignment.qmd
Commit.

@@
 ---
-title: "MSB106_Assignment"
+title: "MSB106 - Assignment"
 author: "Sindre H. Øveraas, Alen Colakovic, Mona Lisa Jones & Sebastian M. Fløysand"
@@
 library(readr)
+library(ggplot2)
+library(kableExtra)
 ```
@@
 Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
-  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning)
+  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
+  rename("knr" = kommunenummer, "DistMal" = HubDist, "HubNameMal" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
 ```
@@
   select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
-  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y)
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
 ```
@@
 
+```{r}
+kable(summary(Commune_Data))
+```
+
+```{r}
+kable(summary(Dentist_Data[3:12]))
+```
+
+```{r}
+Commune_Data |>
+  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
+  theme_bw()
+```
+
+Might be more appropriate to explore the distance from CBDs to Malls compared to turnover capita in retail per inhabitant, rather than the distance from the geographical centroid of Norwegian Communes, because of the tendency of bigger population density around CBDs.
+
+```{r}
+Dentist_Data |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data_LIMIT <- Dentist_Data |>
+  filter(Sum.salgsi < 100000000, DistCBD < 250,
+         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
+```
+
+```{r}
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data |>
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "EBIT") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "EBIT") +
+  theme_bw()
+```
+
 ## Econometric approach

### 2023-02-01 MSB106_Assignment.qmd
Commit.

@@
 Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
-Buffer_CBD <- read_csv("Buffer_CBD.csv")
+Zonal_CBD <- read_csv("Zonal_CBD.csv")
 Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
 NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
-Buffer_den <- read.csv("Buffer_den.csv")
+Zonal_Den <- read.csv("Zonal_Den.csv")
 NOR_KOMM <- read_csv("NOR_KOMM.csv")
@@
 ```{r}
-kable(summary(Dentist_Data[3:12]))
+kable(summary(Dentist_Data[3:6]))
+```
+
+```{r}
+kable(summary(Dentist_Data[8:12]))
 ```

### 2023-02-08 Retail Agglomeration.qmd
Add files via upload
not able to merge

---
title: "Regional and Urban Economics"
subtitle: Dental industry
author: "Sindre Halvorsen Øveraas , Sebastian Mena Fløysand, Alen Colakovic & Mona Lisa Jones"
format:
 pdf:
  fontsize: "10"
  documentclass: article
  margin-left: 2cm
  margin-right: 2cm
papersize: A4
editor: visual
bibliography: references.bib
---

## Abstract

## Introduction

Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"

**Hypothesis**

The location choice of firms providing service to end consumers, significantly determine their ultimate growth potential.

**Research question**

What is the most profitable location for dental practice?

## Literature Review

#### Previous literature on Shopping malls

####

#### Theoretical Framework

**Marshall's agglomeration principles**

Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
Firstly, *knowledge spill over* within industries or product specific technological knowledge.
Furthermore, *market transactions* in terms of value chain transactions with industry-specialized buyers and suppliers.
Lastly, *competition* for specialized production factors such as labour and product market competition @nielsen2021.

**The hotelling model of spatial competition**

Hotelling's (1929) theory of market area model will be applied to look at how the dental industry geographical location placements in the market area, to reach the profit-maximizing equilibrium @mccann2013.

**Alonso model**

The model demonstrate geographical locations tied up to the Alonso's model (1960), which is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
Moreover, indicating where the most profitable location for dental practise might be.
A more central location implies a willingness to pay higher rents due to lower transport costs and higher revenues @capello2015.
Likewise, non-central location yields the same profit if and only if less rent is paid for land @capello2015.
The bid-rent curve will in this case demonstrate the prise of rent within financial locations, Retail locations, and industrial locations.

**Gravitational models ?**

**Retail Agglomeration and Economies of Shopping Centres**

Town and cities have long had shopping districts in which stores are concentrated.
To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
Nevertheless, not coincidentally according to Hotelling theory (1929).
He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.

Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
That being, intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
In this case the dental industry information spillovers are bounded in geographical space.
Hence, the need to locate in close proximity.
That being the latest equipment, in order to work faster or the right personnel to attract more patients.
Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.

Another important externalities expressed in Marshall (1920) theory is market transactions.
Whereas, geographical proximity reduces transaction cost through several mechanism @nielsen2021.
The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." @baker Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
In both cases growth potential would be compromised.

Marshal also points out the competition within the geographical location.
Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
The fact that owners of stores seem to prefer highly concentrated shopping.
More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.

**Shopping Malls**

Owners of shopping malls "orchestrate" the process of retail agglomeration that happens naturally in towns and cities @brueckner2011.
Naturally, the same retail agglomeration and Nash equilibrium may occur in shopping malls if not more intensely as the shops are in closer proximity.
Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.

When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
Much so, because the price of rent is dependent on the overall revenue of a shopping mall.
Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.

The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013.
For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
Nevertheless, they are limited in numbers and needs government approval.
Each municipality must apply to get the wine monopoly in their area.
One of the factors considered in this process is the death of city centre.
A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
Especially if the mall cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
Not to mention the upper hand of easy road access and free parking.

**Urban location of dentist; The Alonso model (1960)**

The bid rent model is a simple theory to explain urban structure @capello2015.
The model assumes a uniformed space where all land is equally fertile, and transport is identical in all directions, with a single specific marketplace the town or city centre in this case @capello2015.
According to @capello2015 the model also assumes that demand is unlimited in other words it is supply-orientated.
Additionally, the production factors are perfectly distributed in space, not uneven spatial distribution @capello2015.
Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale @capello2015.
The last assumptions are perfect competition in the marketplace and that the unit transport costs are constant in space @capello2015.
Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.

![](images/bidrent-01.png)

The illustration shows three categories: closest to the centre the bank and finance districts with the most expensive rent.
Followed by the retail agglomeration areas and industrial sites furthest away from the central point.
Dental practices are service orientated thus depending on population and central activities.
Opening hours are relay outside of office hours.
Production cost does not appear to be important in this case since dentist is a public service and not dependent of regular transportation of sales goods.
The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
From a Marshall's way of thinking dentists should cluster somewhere in this pink area to reap maximum benefits from the agglomeration.
The attendances to break out of the clusters support Hotelling game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013.

**Teori om utvalgt region?**

## Methodology

In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
If the data was collected by the authors this section will describe how the sample was collected.
If specific instruments were constructed to collect the data they will be described here and included in the appendix.
Procedures used to measure variables are often described in sufficient detail to permit replication.

In economic papers, this section describes the empircal model and estimation strategy used by the authors.

#### Data description

(Data and how it was collected)

**SEDA (Spatial Explanatory Data Analyses)**

Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
The Journal of the statistics society define such data as being objects in space, which objects have physical location.
Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
"Invariably the examination of the data involves the examination of some form of map; often it will involve little more. Graphical techniques for examining such maps are thus a central part of the methodology. A number of such techniques are discussed." @haslett1992

#### Empirical Specification

#### Econometric Approach

#### Estimation strategy

#### Analytical framework

## Results

## Discussion

## Conclusion

## Acknowledgements

## References

### 2023-02-09 MSB106_Assignment.qmd
Commit.

@@
 NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
-Zonal_Den <- read.csv("Zonal_Den.csv")
+Zonal_Den <- read.csv("Zonal_Den.csv") |>
+  select(fid, X_sum)
 NOR_KOMM <- read_csv("NOR_KOMM.csv")
@@
   rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
+```
+
+```{r}
+Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
 ```

### 2023-02-14 MSB106_Assignment.qmd
Commit.

@@
 library(kableExtra)
+library(huxtable)
 ```
@@
 
+```{r}
+lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
+lm2 <- lm(Sum.salgsi ~ DistCBD + DistMal + X_sum, data = Dentist_Data_LIMIT)
+```
+
+```{r}
+huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+```
+
 ## Econometric approach

### 2023-02-15 Retail Agglomeration.qmd
updat

@@
 
-#### Theoretical Framework
+### Theoretical Framework
 
@@
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories.
 Firstly, *knowledge spill over* within industries or product specific technological knowledge.
@@
 
-Hotelling's (1929) theory of market area model will be applied to look at how the dental industry geographical location placements in the market area, to reach the profit-maximizing equilibrium @mccann2013.
+Hotelling's (1929) theory of market area model will be applied to look at how the dental industry geographical location placements in the market area, to reach the profit-maximizing equilibrium.
 
@@
 
-The model demonstrate geographical locations tied up to the Alonso's model (1960), which is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
-Moreover, indicating where the most profitable location for dental practise might be.
+Furthermore, geographical locations tied up to **Alonso's model (1960)**, an urbanized formulation of the nineteenth century von Thünen model.
+Moreover, indicating where the optimal location for dental practise might be.
 A more central location implies a willingness to pay higher rents due to lower transport costs and higher revenues @capello2015.
@@
 He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
-At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
+At this point all retailers expected outcome cannot be improved by changing position.
 Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
@@
 Hence, the need to locate in close proximity.
-That being the latest equipment, in order to work faster or the right personnel to attract more patients.
-Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
 
@@
 The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." @baker Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
-Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
-In both cases growth potential would be compromised.
+Presumably, larger cities would be more attractive to dentists compromising cost and quality of specialised labour, should demand in smaller towns or cities not be met.
 
@@
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
+More-so, it suggests that the beneficial outweigh the loss attributed to greater price competition.
 
@@
 
-**Teori om utvalgt region?**
+#### Teori om utvalgt region?
 
-## Methodology
-
-In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
-If the data was collected by the authors this section will describe how the sample was collected.
-If specific instruments were constructed to collect the data they will be described here and included in the appendix.
-Procedures used to measure variables are often described in sufficient detail to permit replication.
-
-In economic papers, this section describes the empircal model and estimation strategy used by the authors.
-
-#### Data description
-
-(Data and how it was collected)
-
-**SEDA (Spatial Explanatory Data Analyses)**
-
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
-The Journal of the statistics society define such data as being objects in space, which objects have physical location.
-Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
-"Invariably the examination of the data involves the examination of some form of map; often it will involve little more.
-Graphical techniques for examining such maps are thus a central part of the methodology.
-A number of such techniques are discussed." @haslett1992
-
-#### Empirical Specification
-
-#### Econometric Approach
-
-#### Estimation strategy
-
-#### Analytical framework
+## Data
 

### 2023-02-15 Retail Agglomeration.qmd
up date

@@
 
+Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
+It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
+Furthermore, the theory uses the concept of externalities as disparities in the spatial distubution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
+
+Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
+It also involves generatating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+
+Harald Hotelling's locational equilibrium is determined by a logic of profit maximization whereby each producer controls its own market area.
+Productivity advantages of cities and urban clusters with a high density of firms increase profit by attracting a larger number of potential customers, and more productive workers @capello2011.
+Furthermore, the attractiveness of a central location increase the cost of rent.
+Alonso's bid rent model indicates the most profitable location for firms.
+Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
+In Reilly's approach, the attractiveness of the retail location, represent the size of the retail centre.
+Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
+
+A generalized Models based on Newton law of gravity can be applied to estimate flows between two points such as, cities or other points of interest @capello2015.
+The model of potential has the capacity to measure the potential of attractiveness to a place.
+Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
+
+**Interdependent location choice, the Hotelling model (1929)**
+
+The model assume that given the location of producers, and given demand uniformly distributed geographically (in linear or circular form) the market is divided into areas within each of which there operates a single firm in a duopoly environment.
+Furthermore, no relocation costs and demand only depend on location choice @capello2015.
+
+The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
+One firm starts relocating closer to the other to take some of the customers in the other market area @capello2015.
+The other respond by doing the same and the game continues until both end up in the middle of the market on the broader of AB.
+The end of the game is the position where neither can increase sales volume by moving position @capello2015.
+
+A simple explanation of why two dentists providing the same service, at the same price might locate next to each other.
+Nevertheless, despite increasing the transportation cost for patients.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the excistance of larger agglomerations.
+
 **Marshall's agglomeration principles**
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories.
-Firstly, *knowledge spill over* within industries or product specific technological knowledge.
-Furthermore, *market transactions* in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, *competition* for specialized production factors such as labour and product market competition @nielsen2021.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially driving sales.
+Firstly, knowledge spill over within industries or product specific technological knowledge.
+Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
+Lastly, competition for specialized production factors such as labour and product market competition Nielsen et al. (2021).
 
-**The hotelling model of spatial competition**
+There are solidly established conclusions regarding the exsistanse of agglomeration economies Puga (2010).
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
+Nevertheless, there is a large theoretical literature that develops these mechanisms Puga (2010).
+Duranton and Puga (2004) discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
 
-Hotelling's (1929) theory of market area model will be applied to look at how the dental industry geographical location placements in the market area, to reach the profit-maximizing equilibrium.
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills Puga (2010).
+Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
+The attraction for the consumers and users of public facilities is overall cost reduction Puga (2010).
+Hence, the larger the population sharing facilities the lower the cost per user Puga (2010).
+Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
 
-**Alonso model**
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches Puga (2010).
+Shopping malls accure skilled workers to drive sales.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities Davis and Dingel (2020).
+More-so, cites provide a constant market for specialized skills and more productive work force Puga (2010).
+Perhaps, a possibility for higher wages being compensated by more efficiant workers in bigger agglomeration economies.
+That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice Puga (2010).
 
-Furthermore, geographical locations tied up to **Alonso's model (1960)**, an urbanized formulation of the nineteenth century von Thünen model.
-Moreover, indicating where the optimal location for dental practise might be.
-A more central location implies a willingness to pay higher rents due to lower transport costs and higher revenues @capello2015.
-Likewise, non-central location yields the same profit if and only if less rent is paid for land @capello2015.
-The bid-rent curve will in this case demonstrate the prise of rent within financial locations, Retail locations, and industrial locations.
+kladd
 
-**Gravitational models ?**
+*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* Puga (2010)
+
+**Urban location of dentist; The Alonso model**
+
+The model demonstrate geographical locations tied up to location.
+It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
+Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
+Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
+Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
+As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
+More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
+@capello2015.
+The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
+Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
+The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.3
+
+![](images/bidrent.png)
+
+The illustration shows three categories: The most expensive rent closest to the centre is the Bank and finance district, followed by the retail agglomeration areas and industrial sites furthest away from the central point.
+Dental practices are service orientated thus depending on population density and central activities.
+Opening hours are relay outside of office hours.
+Production cost does not appear to be important in this case since dentist is a public service.
+Hence, not dependent of regular transportation of sales goods.
+The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
+From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
+Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
+The attractiveness might be high in these places due to economies of scope.
+As the positive benefits from the Marshallian agglomeration attributes play out.
+Subsequently, increasing profit by trading higher transportation cost for lower rent.
+In addition fewer dentists to compete with.
+The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
+
+**Gravitational models**
+
+The centre attracts firms, people and activities.
+Furthermore, it influence the central points in diverse ways such as commuter movements, disfuction of knowledge, corporations and network and personal relationship @capello2015.
+More-so, they are generated by the intensity of the flows generated between places @capello2015.
+The gravitational pull to a place of interest for a dentist will determine how much sales can be generated by locating there.
+Several locations can then be measures against each other to determine profit maximisation.
+These models has the potential to measures the future attractiveness of a place.
+Changes in infrastructure such easier access to a place might change the attractiveness of a place or move the flow of people.
+An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
+Subserviently, increasing the gravitationally pull between the two places.
+Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
+Rogfast might move the population of Haugesund futher south and a new town centre might occur competing with other retail clusters in years to come.
+On the other hand the flow of people towards Raglamyr might increase untill the potental new centre place has grown.
+These models have predictive capacity and can be used to estimate the potential impact.
+In particular, location of new productive activity in an area @capello2015.
+For instance predict where to build a shopping centre @capello2015.
+A new shopping centre further south in Haugaland or expantion of excisting retail agglomeration areas might be of interest.
+As a result of increased population.
+By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
 
@@
 Town and cities have long had shopping districts in which stores are concentrated.
-To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
-Nevertheless, not coincidentally according to Hotelling theory (1929).
+To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011. Nevertheless, not coincidentally according to Hotelling's theory (1929).
 He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
-At this point all retailers expected outcome cannot be improved by changing position.
+At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
 Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
-
 Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
-That being, intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
+That being,intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
 In this case the dental industry information spillovers are bounded in geographical space.
 Hence, the need to locate in close proximity.
+That being the latest equipment, in order to work faster or the right personnel to attract more patients.
+Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
 
@@
 The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-Presumably, larger cities would be more attractive to dentists compromising cost and quality of specialised labour, should demand in smaller towns or cities not be met.
+"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
+Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
+In both cases growth potential would be compromised.
 
 Marshal also points out the competition within the geographical location.
-Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
-The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the beneficial outweigh the loss attributed to greater price competition.
+Hence, the marketplace and business districts give consumers low search and switching costs @capello2015. The fact that owners of stores seem to prefer highly concentrated shopping.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
 
@@
 Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.
-
 When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
@@
 Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
-
 The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
-Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013.
+Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013Lai.
 For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
@@
 A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
-Especially if the mall cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
+Perhaps more so if malls cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
 Not to mention the upper hand of easy road access and free parking.
-
-**Urban location of dentist; The Alonso model (1960)**
-
-The bid rent model is a simple theory to explain urban structure @capello2015.
-The model assumes a uniformed space where all land is equally fertile, and transport is identical in all directions, with a single specific marketplace the town or city centre in this case @capello2015.
-According to @capello2015 the model also assumes that demand is unlimited in other words it is supply-orientated.
-Additionally, the production factors are perfectly distributed in space, not uneven spatial distribution @capello2015.
-Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale @capello2015.
-The last assumptions are perfect competition in the marketplace and that the unit transport costs are constant in space @capello2015.
-Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
-
-![](images/bidrent-01.png)
-
-The illustration shows three categories: closest to the centre the bank and finance districts with the most expensive rent.
-Followed by the retail agglomeration areas and industrial sites furthest away from the central point.
-Dental practices are service orientated thus depending on population and central activities.
-Opening hours are relay outside of office hours.
-Production cost does not appear to be important in this case since dentist is a public service and not dependent of regular transportation of sales goods.
-The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
-From a Marshall's way of thinking dentists should cluster somewhere in this pink area to reap maximum benefits from the agglomeration.
-The attendances to break out of the clusters support Hotelling game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013.
-
-#### Teori om utvalgt region?
 

### 2023-02-15 MSB106_Assignment.qmd
teori update

@@
 ## Theoretical Foundation of Hypothesis
+
+Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
+It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
+Furthermore, the theory uses the concept of externalities as disparities in the spatial distubution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
+
+Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
+It also involves generatating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+
+Harald Hotelling's locational equilibrium is determined by a logic of profit maximization whereby each producer controls its own market area.
+Productivity advantages of cities and urban clusters with a high density of firms increase profit by attracting a larger number of potential customers, and more productive workers @capello2011.
+Furthermore, the attractiveness of a central location increase the cost of rent.
+Alonso's bid rent model indicates the most profitable location for firms.
+Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
+In Reilly's approach, the attractiveness of the retail location, represent the size of the retail centre.
+Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
+
+A generalized Models based on Newton law of gravity can be applied to estimate flows between two points such as, cities or other points of interest @capello2015.
+The model of potential has the capacity to measure the potential of attractiveness to a place.
+Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
+
+**Interdependent location choice, the Hotelling model (1929)**
+
+The model assume that given the location of producers, and given demand uniformly distributed geographically (in linear or circular form) the market is divided into areas within each of which there operates a single firm in a duopoly environment.
+Furthermore, no relocation costs and demand only depend on location choice @capello2015.
+
+The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
+One firm starts relocating closer to the other to take some of the customers in the other market area @capello2015.
+The other respond by doing the same and the game continues until both end up in the middle of the market on the broader of AB.
+The end of the game is the position where neither can increase sales volume by moving position @capello2015.
+
+A simple explanation of why two dentists providing the same service, at the same price might locate next to each other.
+Nevertheless, despite increasing the transportation cost for patients.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the excistance of larger agglomerations.
+
+**Marshall's agglomeration principles**
+
+Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially driving sales.
+Firstly, knowledge spill over within industries or product specific technological knowledge.
+Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
+Lastly, competition for specialized production factors such as labour and product market competition Nielsen et al. (2021).
+
+There are solidly established conclusions regarding the exsistanse of agglomeration economies Puga (2010).
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
+Nevertheless, there is a large theoretical literature that develops these mechanisms Puga (2010).
+Duranton and Puga (2004) discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills Puga (2010).
+Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
+The attraction for the consumers and users of public facilities is overall cost reduction Puga (2010).
+Hence, the larger the population sharing facilities the lower the cost per user Puga (2010).
+Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
+
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches Puga (2010).
+Shopping malls accure skilled workers to drive sales.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities Davis and Dingel (2020).
+More-so, cites provide a constant market for specialized skills and more productive work force Puga (2010).
+Perhaps, a possibility for higher wages being compensated by more efficiant workers in bigger agglomeration economies.
+That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice Puga (2010).
+
+kladd
+
+*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* Puga (2010)
+
+**Urban location of dentist; The Alonso model**
+
+The model demonstrate geographical locations tied up to location.
+It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
+Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
+Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
+Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
+As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
+More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
+@capello2015.
+The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
+Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
+The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.3
+
+![](images/bidrent.png)
+
+The illustration shows three categories: The most expensive rent closest to the centre is the Bank and finance district, followed by the retail agglomeration areas and industrial sites furthest away from the central point.
+Dental practices are service orientated thus depending on population density and central activities.
+Opening hours are relay outside of office hours.
+Production cost does not appear to be important in this case since dentist is a public service.
+Hence, not dependent of regular transportation of sales goods.
+The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
+From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
+Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
+The attractiveness might be high in these places due to economies of scope.
+As the positive benefits from the Marshallian agglomeration attributes play out.
+Subsequently, increasing profit by trading higher transportation cost for lower rent.
+In addition fewer dentists to compete with.
+The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
+
+**Gravitational models**
+
+The centre attracts firms, people and activities.
+Furthermore, it influence the central points in diverse ways such as commuter movements, disfuction of knowledge, corporations and network and personal relationship @capello2015.
+More-so, they are generated by the intensity of the flows generated between places @capello2015.
+The gravitational pull to a place of interest for a dentist will determine how much sales can be generated by locating there.
+Several locations can then be measures against each other to determine profit maximisation.
+These models has the potential to measures the future attractiveness of a place.
+Changes in infrastructure such easier access to a place might change the attractiveness of a place or move the flow of people.
+An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
+Subserviently, increasing the gravitationally pull between the two places.
+Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
+Rogfast might move the population of Haugesund futher south and a new town centre might occur competing with other retail clusters in years to come.
+On the other hand the flow of people towards Raglamyr might increase untill the potental new centre place has grown.
+These models have predictive capacity and can be used to estimate the potential impact.
+In particular, location of new productive activity in an area @capello2015.
+For instance predict where to build a shopping centre @capello2015.
+A new shopping centre further south in Haugaland or expantion of excisting retail agglomeration areas might be of interest.
+As a result of increased population.
+By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
+
+**Retail Agglomeration and Economies of Shopping Centres**
+
+Town and cities have long had shopping districts in which stores are concentrated.
+To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
+Nevertheless, not coincidentally according to Hotelling's theory (1929).
+He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
+At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
+Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
+Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
+That being,intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
+In this case the dental industry information spillovers are bounded in geographical space.
+Hence, the need to locate in close proximity.
+That being the latest equipment, in order to work faster or the right personnel to attract more patients.
+Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
+
+Another important externalities expressed in Marshall (1920) theory is market transactions.
+Whereas, geographical proximity reduces transaction cost through several mechanism @nielsen2021.
+The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
+"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
+Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
+In both cases growth potential would be compromised.
+
+Marshal also points out the competition within the geographical location.
+Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
+The fact that owners of stores seem to prefer highly concentrated shopping.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
+
+**Shopping Malls**
+
+Owners of shopping malls "orchestrate" the process of retail agglomeration that happens naturally in towns and cities @brueckner2011.
+Naturally, the same retail agglomeration and Nash equilibrium may occur in shopping malls if not more intensely as the shops are in closer proximity.
+Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.
+When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
+Much so, because the price of rent is dependent on the overall revenue of a shopping mall.
+Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
+The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
+Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013Lai.
+For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
+Nevertheless, they are limited in numbers and needs government approval.
+Each municipality must apply to get the wine monopoly in their area.
+One of the factors considered in this process is the death of city centre.
+A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
+Perhaps more so if malls cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
+Not to mention the upper hand of easy road access and free parking.
 

### 2023-02-15 MSB106_Assignment.qmd
...

@@
 editor: visual
+bibliography: references.bib
 ---
@@
 
+**Abstract**
+
 ## Introduction
@@
 
+**Hypothesis & Research question**
+
+The location choice of firms providing service to end consumers, significantly determine their ultimate growth potential.
+
+What is the most profitable location for dental practice?
+
+**Location Theory**
+
 Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
 It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities as disparities in the spatial distubution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
+Furthermore, the theory uses the concept of externalities as disparities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
 
 Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generatating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+It also involves generating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
 
@@
 
-**Interdependent location choice, the Hotelling model (1929)**
+**Interdependent location choice, the Hotelling's model (1929)**
 
@@
 Nevertheless, despite increasing the transportation cost for patients.
-Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the excistance of larger agglomerations.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomerations.
 
@@
 
-There are solidly established conclusions regarding the exsistanse of agglomeration economies Puga (2010).
-However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
-Nevertheless, there is a large theoretical literature that develops these mechanisms Puga (2010).
-Duranton and Puga (2004) discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+There are solidly established conclusions regarding the excistanse of agglomeration economies @puga2010.
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010.
+Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010.
+@duranton2004 discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills Puga (2010).
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010..
 Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction Puga (2010).
-Hence, the larger the population sharing facilities the lower the cost per user Puga (2010).
+The attraction for the consumers and users of public facilities is overall cost reduction @puga2010.
+Hence, the larger the population sharing facilities the lower the cost per user @puga2010.
 Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
 
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches Puga (2010).
-Shopping malls accure skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities Davis and Dingel (2020).
-More-so, cites provide a constant market for specialized skills and more productive work force Puga (2010).
-Perhaps, a possibility for higher wages being compensated by more efficiant workers in bigger agglomeration economies.
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010.
+Shopping malls require skilled workers to drive sales.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities @davis2020Davis.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010.
+Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
 That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice Puga (2010).
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010.
 
@@
 *"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* Puga (2010)
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* @puga2010.
 
@@
 
+`In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
+If the data was collected by the authors this section will describe how the sample was collected.
+If specific instruments were constructed to collect the data they will be described here and included in the appendix.
+Procedures used to measure variables are often described in sufficient detail to permit replication.
+In economic papers, this section describes the empirical model and estimation strategy used by the authors.`
+
+**Data description**
+
+(Data and how it was collected)
+
+**Empirical Specification**
+
+**Econometric Approach**
+
+**Estimation strategy**
+
+**Analytical framework**
+
+SEDA (Spatial Explanatory Data Analyses)
+
+Spatial data analysis is a rapidly growing area in Statistics @haslett1992. The Journal of the statistics society define such data as being objects in space, which objects have physical location.
+Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
+
+The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
+Graphical techniques for examining such maps are thus a central part of the methodology.
+A number of such techniques are discussed."` @haslett1992.
+
+`What data are used?`
+
+`What has been calculated?`
+
+`What was not used and why?`
+
+`Outlines?`
+
 ```{r}
@@
 ## Econometric approach
+
+**Result**
+
+**Discussion**
+
+**Conclusion**

### 2023-02-15 Retail Agglomeration.qmd
...

@@
 
-####
-
 ### Theoretical Framework
@@
 
-There are solidly established conclusions regarding the exsistanse of agglomeration economies Puga (2010).
+There are solidly established conclusions regarding the excistanse of agglomeration economies Puga (2010).
 However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
@@
 Town and cities have long had shopping districts in which stores are concentrated.
-To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011. Nevertheless, not coincidentally according to Hotelling's theory (1929).
+To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
+Nevertheless, not coincidentally according to Hotelling's theory (1929).
 He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
@@
 Marshal also points out the competition within the geographical location.
-Hence, the marketplace and business districts give consumers low search and switching costs @capello2015. The fact that owners of stores seem to prefer highly concentrated shopping.
+Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
+The fact that owners of stores seem to prefer highly concentrated shopping.
 More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
@@
 
+`In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
+If the data was collected by the authors this section will describe how the sample was collected.
+If specific instruments were constructed to collect the data they will be described here and included in the appendix.
+Procedures used to measure variables are often described in sufficient detail to permit replication.
+In economic papers, this section describes the empirical model and estimation strategy used by the authors.`
+
+**Data description**
+
+(Data and how it was collected)
+
+**Empirical Specification**
+
+**Econometric Approach**
+
+**Estimation strategy**
+
+**Analytical framework**
+
+SEDA (Spatial Explanatory Data Analyses)
+
+Spatial data analysis is a rapidly growing area in Statistics @haslett1992. The Journal of the statistics society define such data as being objects in space, which objects have physical location.
+Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
+
+The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
+Graphical techniques for examining such maps are thus a central part of the methodology.
+A number of such techniques are discussed."` @haslett1992.
+
+`What data are used?`
+
+`What has been calculated?`
+
 ## Results

### 2023-02-16 MSB106_Assignment.qmd
Skal vi ha flere overskrifter? Skal vi bruke hoveddokumentet?
...

@@
 Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially driving sales.
-Firstly, knowledge spill over within industries or product specific technological knowledge.
-Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition Nielsen et al. (2021).
+Firstly, *knowledge spill over* within industries or product specific technological knowledge.
+Furthermore, *market transactions* in terms of value chain transactions with industry-specialized buyers and suppliers.
+Lastly, *competition* for specialized production factors such as labour and product market competition Nielsen et al. (2021).
 
@@
 
+*Sharing*
+
 A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010..
-Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
+Dspite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
 The attraction for the consumers and users of public facilities is overall cost reduction @puga2010.
@@
 Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
+
+*Matching*
 
@@
 
-kladd
+*Learning*
 
@@
 
-**Data description**
-
 (Data and how it was collected)
@@
 
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992. The Journal of the statistics society define such data as being objects in space, which objects have physical location.
+Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
+The Journal of the statistics society define such data as being objects in space, which objects have physical location.
 Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.

### 2023-02-16 Retail Agglomeration.qmd
...

@@
 It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities as disparities in the spatial distubution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
+Furthermore, the theory uses the concept of externalities as disparities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
 
 Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generatating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+It also involves generating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
 
@@
 Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
-In Reilly's approach, the attractiveness of the retail location, represent the size of the retail centre.
+In gravitational models, the attractiveness of the retail location, represent the size of the retail centre.
 Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
@@
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially driving sales.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
 Firstly, knowledge spill over within industries or product specific technological knowledge.
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition Nielsen et al. (2021).
+Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021Nielsen.
 
-There are solidly established conclusions regarding the excistanse of agglomeration economies Puga (2010).
-However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
-Nevertheless, there is a large theoretical literature that develops these mechanisms Puga (2010).
-Duranton and Puga (2004) discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+There are solidly established conclusions regarding the excistanse of agglomeration economies @puga2010Puga.
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
+Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
+@duranton2004Duranton discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills Puga (2010).
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
 Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction Puga (2010).
-Hence, the larger the population sharing facilities the lower the cost per user Puga (2010).
+The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
+Hence, the larger the population sharing facilities the lower the cost per user @puga2010Puga.
 Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
 
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches Puga (2010).
-Shopping malls accure skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities Davis and Dingel (2020).
-More-so, cites provide a constant market for specialized skills and more productive work force Puga (2010).
-Perhaps, a possibility for higher wages being compensated by more efficiant workers in bigger agglomeration economies.
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010Puga.
+Shopping malls require skilled workers to drive sales.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020Davis.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010Puga.
+Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
 That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice Puga (2010).
-
-kladd
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010Puga.
 
 *"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* Puga (2010)
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* @puga2010Puga
 
@@
 Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
-The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.3
+The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
 
@@
 The centre attracts firms, people and activities.
-Furthermore, it influence the central points in diverse ways such as commuter movements, disfuction of knowledge, corporations and network and personal relationship @capello2015.
+Furthermore, it influence the central points in diverse ways such as commuter movements, disfustion of knowledge, corporations and network and personal relationship @capello2015.
 More-so, they are generated by the intensity of the flows generated between places @capello2015.
@@
 Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
-Rogfast might move the population of Haugesund futher south and a new town centre might occur competing with other retail clusters in years to come.
-On the other hand the flow of people towards Raglamyr might increase untill the potental new centre place has grown.
+Rogfast might move the population of Haugesund further south and a new town centre might occur competing with other retail clusters in years to come.
+On the other hand the flow of people towards Raglamyr might increase until the potential new centre place has grown.
 These models have predictive capacity and can be used to estimate the potential impact.
@@
 For instance predict where to build a shopping centre @capello2015.
-A new shopping centre further south in Haugaland or expantion of excisting retail agglomeration areas might be of interest.
+A new shopping centre further south in Haugaland or expansion of excising retail agglomeration areas might be of interest.
 As a result of increased population.

### 2023-02-16 MSB106_Assignment.qmd
added to Mena

--- 2023-02-15 MSB106_Assignment.qmd
@@
 library(kableExtra)
-library(huxtable)
 ```
@@
 NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
-Zonal_Den <- read.csv("Zonal_Den.csv") |>
-  select(fid, X_sum)
+Zonal_Den <- read.csv("Zonal_Den.csv")
 NOR_KOMM <- read_csv("NOR_KOMM.csv")
@@
 
-```{r}
-Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
-```
-
 ## Introduction
@@
 ## Theoretical Foundation of Hypothesis
-
-Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
-It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities as disparities in the spatial distubution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
-
-Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generatating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
-
-Harald Hotelling's locational equilibrium is determined by a logic of profit maximization whereby each producer controls its own market area.
-Productivity advantages of cities and urban clusters with a high density of firms increase profit by attracting a larger number of potential customers, and more productive workers @capello2011.
-Furthermore, the attractiveness of a central location increase the cost of rent.
-Alonso's bid rent model indicates the most profitable location for firms.
-Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
-In Reilly's approach, the attractiveness of the retail location, represent the size of the retail centre.
-Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
-
-A generalized Models based on Newton law of gravity can be applied to estimate flows between two points such as, cities or other points of interest @capello2015.
-The model of potential has the capacity to measure the potential of attractiveness to a place.
-Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
-
-**Interdependent location choice, the Hotelling model (1929)**
-
-The model assume that given the location of producers, and given demand uniformly distributed geographically (in linear or circular form) the market is divided into areas within each of which there operates a single firm in a duopoly environment.
-Furthermore, no relocation costs and demand only depend on location choice @capello2015.
-
-The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
-One firm starts relocating closer to the other to take some of the customers in the other market area @capello2015.
-The other respond by doing the same and the game continues until both end up in the middle of the market on the broader of AB.
-The end of the game is the position where neither can increase sales volume by moving position @capello2015.
-
-A simple explanation of why two dentists providing the same service, at the same price might locate next to each other.
-Nevertheless, despite increasing the transportation cost for patients.
-Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the excistance of larger agglomerations.
-
-**Marshall's agglomeration principles**
-
-Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially driving sales.
-Firstly, knowledge spill over within industries or product specific technological knowledge.
-Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition Nielsen et al. (2021).
-
-There are solidly established conclusions regarding the exsistanse of agglomeration economies Puga (2010).
-However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult Puga (2010).
-Nevertheless, there is a large theoretical literature that develops these mechanisms Puga (2010).
-Duranton and Puga (2004) discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
-
-A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills Puga (2010).
-Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction Puga (2010).
-Hence, the larger the population sharing facilities the lower the cost per user Puga (2010).
-Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
-
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches Puga (2010).
-Shopping malls accure skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities Davis and Dingel (2020).
-More-so, cites provide a constant market for specialized skills and more productive work force Puga (2010).
-Perhaps, a possibility for higher wages being compensated by more efficiant workers in bigger agglomeration economies.
-That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice Puga (2010).
-
-kladd
-
-*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* Puga (2010)
-
-**Urban location of dentist; The Alonso model**
-
-The model demonstrate geographical locations tied up to location.
-It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
-Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
-Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
-Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
-As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
-More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
-@capello2015.
-The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
-Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
-The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.3
-
-![](images/bidrent.png)
-
-The illustration shows three categories: The most expensive rent closest to the centre is the Bank and finance district, followed by the retail agglomeration areas and industrial sites furthest away from the central point.
-Dental practices are service orientated thus depending on population density and central activities.
-Opening hours are relay outside of office hours.
-Production cost does not appear to be important in this case since dentist is a public service.
-Hence, not dependent of regular transportation of sales goods.
-The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
-From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
-Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
-The attractiveness might be high in these places due to economies of scope.
-As the positive benefits from the Marshallian agglomeration attributes play out.
-Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with.
-The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
-
-**Gravitational models**
-
-The centre attracts firms, people and activities.
-Furthermore, it influence the central points in diverse ways such as commuter movements, disfuction of knowledge, corporations and network and personal relationship @capello2015.
-More-so, they are generated by the intensity of the flows generated between places @capello2015.
-The gravitational pull to a place of interest for a dentist will determine how much sales can be generated by locating there.
-Several locations can then be measures against each other to determine profit maximisation.
-These models has the potential to measures the future attractiveness of a place.
-Changes in infrastructure such easier access to a place might change the attractiveness of a place or move the flow of people.
-An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
-Subserviently, increasing the gravitationally pull between the two places.
-Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
-Rogfast might move the population of Haugesund futher south and a new town centre might occur competing with other retail clusters in years to come.
-On the other hand the flow of people towards Raglamyr might increase untill the potental new centre place has grown.
-These models have predictive capacity and can be used to estimate the potential impact.
-In particular, location of new productive activity in an area @capello2015.
-For instance predict where to build a shopping centre @capello2015.
-A new shopping centre further south in Haugaland or expantion of excisting retail agglomeration areas might be of interest.
-As a result of increased population.
-By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
-
-**Retail Agglomeration and Economies of Shopping Centres**
-
-Town and cities have long had shopping districts in which stores are concentrated.
-To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
-Nevertheless, not coincidentally according to Hotelling's theory (1929).
-He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
-At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
-Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
-Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
-That being,intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
-In this case the dental industry information spillovers are bounded in geographical space.
-Hence, the need to locate in close proximity.
-That being the latest equipment, in order to work faster or the right personnel to attract more patients.
-Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
-
-Another important externalities expressed in Marshall (1920) theory is market transactions.
-Whereas, geographical proximity reduces transaction cost through several mechanism @nielsen2021.
-The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
-Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
-In both cases growth potential would be compromised.
-
-Marshal also points out the competition within the geographical location.
-Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
-The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
-
-**Shopping Malls**
-
-Owners of shopping malls "orchestrate" the process of retail agglomeration that happens naturally in towns and cities @brueckner2011.
-Naturally, the same retail agglomeration and Nash equilibrium may occur in shopping malls if not more intensely as the shops are in closer proximity.
-Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.
-When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
-Much so, because the price of rent is dependent on the overall revenue of a shopping mall.
-Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
-The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
-Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013Lai.
-For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
-Nevertheless, they are limited in numbers and needs government approval.
-Each municipality must apply to get the wine monopoly in their area.
-One of the factors considered in this process is the death of city centre.
-A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
-Perhaps more so if malls cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
-Not to mention the upper hand of easy road access and free parking.
 
@@
 
-```{r}
-lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
-lm2 <- lm(Sum.salgsi ~ DistCBD + DistMal + X_sum, data = Dentist_Data_LIMIT)
-```
+## Econometric approach
 
-```{r}
-huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
-```
+After a quixk view of some of our findings where we used the dentist income (y) and in comparison to the Distance from Mall (x) we already see some signs of a pattern where the income is grater when the dentist is placed closer to the shopping mall and decreases as the distance gets bigger.
+However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
+The theory from Keynesian suggest that this pattern might take place as the transportation cost for individuals further from the Shopping mall does not increase to much by travelling the extra distance.
+Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
 
-## Econometric approach
+\
+
+We also find that most dentist are either located close to a City, Central Business District or a shopping mall.
+
+\
+
+Dentist are also located close to echeother making for a sort of cluster, indicating that it might be som advantages beeing located in a cluster.
+Taken for granted that or data on Dentist includes both \"normal Dentis\" and Spescialist, there is reason to believe that it could be an advantage beeing located close together.
+As found in Marshalls Theory their ar several advantages by locating in clusters.
+Where in this case a \"normal dentist\" would for example recommend one of their customers to go see a specialist to get braces.
+The customer would then probably prefer to go to the closes specialist.
+The access to work force in these areas would also be good and wll be whitin the same geographic area.
+
+\

### 2023-02-17 MSB106_Assignment.qmd
more info

@@
 Dentist are also located close to echeother making for a sort of cluster, indicating that it might be som advantages beeing located in a cluster.
-Taken for granted that or data on Dentist includes both \"normal Dentis\" and Spescialist, there is reason to believe that it could be an advantage beeing located close together.
+Taken for granted that or data on Dentist includes both "normal Dentis" and Spescialist, there is reason to believe that it could be an advantage beeing located close together.
 As found in Marshalls Theory their ar several advantages by locating in clusters.
-Where in this case a \"normal dentist\" would for example recommend one of their customers to go see a specialist to get braces.
+Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
 The customer would then probably prefer to go to the closes specialist.
@@
 \
+as mentioned earlier, dentis should place them self in the pink area of the Alons model.
+however, a dentist beeing placed further out in to other retail agglomeration areas can also be a strategy, but the dentist would then rely on the area having less competition so the rate between competition and population density are proposinal.
+Our findings in QGIS shows that in areas where it is a high population density, the number of dentist also is higher.
+The user/pasient then has more to choose between and the dentist also has a bigger scale of potential customers.
+Where in smaller areas it is less competition and the dentis would appear to have somewhat of a monopoly since the transportation cost to the second closes dentis are noticeable higher.
+
+\
+
+it is important to be aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentis based on their living address.
+This also takes away som of the competition.
+Another thing that is important to notice is that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
+a suggestion would be that for a dentist to be placed close CBD areas, shopping malls and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
+So by beeing placed close CBD areas, shopping malls and industry the transportation cost for the customer could be reduced as they are already located in these areas even if they dont live here
+
+\

### 2023-02-17 MSB106_Assignment.qmd
Commit.

--- 2023-02-16 MSB106_Assignment.qmd
@@
 
+```{r}
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = X_sum, y = Sum.salgsi, colour = Sum.salgsi)) +
+  geom_point(lwd = .9) +
+  labs(x = "Pop density", y = "Income") +
+  theme_bw()
+```
+
+```{r}
+lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
+lm2 <- lm(Sum.salgsi ~ DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
+```
+
+```{r}
+huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+```
+
 ## Econometric approach
-
-After a quixk view of some of our findings where we used the dentist income (y) and in comparison to the Distance from Mall (x) we already see some signs of a pattern where the income is grater when the dentist is placed closer to the shopping mall and decreases as the distance gets bigger.
-However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
-The theory from Keynesian suggest that this pattern might take place as the transportation cost for individuals further from the Shopping mall does not increase to much by travelling the extra distance.
-Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
-
-\
-
-We also find that most dentist are either located close to a City, Central Business District or a shopping mall.
-
-\
-
-Dentist are also located close to echeother making for a sort of cluster, indicating that it might be som advantages beeing located in a cluster.
-Taken for granted that or data on Dentist includes both \"normal Dentis\" and Spescialist, there is reason to believe that it could be an advantage beeing located close together.
-As found in Marshalls Theory their ar several advantages by locating in clusters.
-Where in this case a \"normal dentist\" would for example recommend one of their customers to go see a specialist to get braces.
-The customer would then probably prefer to go to the closes specialist.
-The access to work force in these areas would also be good and wll be whitin the same geographic area.
-
-\

### 2023-02-18 Retail Agglomeration.qmd
Commit.

@@
 
-(Data and how it was collected)
+For this assignment we where provided with different geospatial data in the geographic information software QGIS to calculate the relationship of the different variables that might determine firm activity.
+This geospatial data included Norwegian commune data which contained the administrative boundaries of the Communes, presented in the software as different divisions of lines that distinguish the different commune boundaties.
+We where also provided with different point layers, such as the geometric centers of communes, urban city centers or Central Business Districts (CBD), all businesses in Norway which are regarded as dental services, and all Norwegian shopping malls.
+We where also provided with a QGIS layer with population density estimates, visualized with 100x100 raster graphics.
+
+To carry out a full-flegded data analysis of the influencing factors of the activity of Norwegian dentist services, we collected some more data than what we already had been presented with.
+We firstly utilized QGIS to calculate different other statistics and variables.
+Regarding the Norwegian shopping malls layer we added the variable \"Wine-monopoly\" as a false/ true variable to determine which Norwegian malls, and also the number of stores which are in each mall.
+The malls opening hours have also been included in this layer.
+From SSB, we gathered data on the average municipal sales per inhabitant in retailing for each Norwegian commune.
+We also created 2 km buffer zones around dentists and CBDs to determine population density in these areas.
 

### 2023-02-18 Retail Agglomerationextra.qmd
overført
...

--- 2023-02-16 Retail Agglomeration.qmd
@@
 
-A generalized Models based on Newton law of gravity can be applied to estimate flows between two points such as, cities or other points of interest @capello2015.
 The model of potential has the capacity to measure the potential of attractiveness to a place.
@@
 
+```{r}
+NOR_CBD <- read_csv("NOR_CBD.csv")
+Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
+Buffer_CBD <- read_csv("Buffer_CBD.csv")
+Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
+NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
+Buffer_den <- read.csv("Buffer_den.csv")
+NOR_KOMM <- read_csv("NOR_KOMM.csv")
+NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
+```
+
+```{r}
+Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
+  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning)
+```
+
+```{r}
+Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
+  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y)
+```
+
+```{r}
+kable(summary(Commune_Data))
+```
+
+```{r}
+kable(summary(Dentist_Data[3:6]))
+```
+
+```{r}
+kable(summary(Dentist_Data[8:12]))
+```
+
+```{r}
+Commune_Data |>
+  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data_LIMIT <- Dentist_Data |>
+  filter(Sum.salgsi < 100000000, DistCBD < 250,
+         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
+```
+
+```{r}
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data |>
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "EBIT") +
+  theme_bw()
+```
+
+```{r}
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "EBIT") +
+  theme_bw()
+```
+
 `In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.

### 2023-02-18 Retail Agglomerationextra.qmd
...

@@
 Furthermore, the attractiveness of a central location increase the cost of rent.
+
 Alonso's bid rent model indicates the most profitable location for firms.
@@
 
-There are solidly established conclusions regarding the excistanse of agglomeration economies @puga2010Puga.
+There are solidly established conclusions regarding the excitanse of agglomeration economies @puga2010Puga.
 However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
@@
 The centre attracts firms, people and activities.
-Furthermore, it influence the central points in diverse ways such as commuter movements, disfustion of knowledge, corporations and network and personal relationship @capello2015.
+Furthermore, it influence the central points in diverse ways such as commuter movements, diffusion of knowledge, corporations and network and personal relationship @capello2015.
 More-so, they are generated by the intensity of the flows generated between places @capello2015.
@@
 These models has the potential to measures the future attractiveness of a place.
-Changes in infrastructure such easier access to a place might change the attractiveness of a place or move the flow of people.
+Changes in infrastructure such as easier access to a place might change the attractiveness of a place or move the flow of people.
 An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
@@
 Rogfast might move the population of Haugesund further south and a new town centre might occur competing with other retail clusters in years to come.
-On the other hand the flow of people towards Raglamyr might increase until the potential new centre place has grown.
+On the other hand, the flow of people towards Raglamyr might increase until the potential new centre place has grown.
 These models have predictive capacity and can be used to estimate the potential impact.
@@
 kable(summary(Commune_Data))
+View(commune_data)
 ```
@@
 
+The question addressed are related to the geographical space of Norway.
+Norway is a country with relatively small population and large heterogeneous land area.
+Furthermore, the population is decentralized with the majority of population living in urban areas.
+This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
+Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
+Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
+For exsampel faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
+As distance increases, communication networks become important allowing the growth of sub centres to functions as self organising systems.
+
+The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
+Having an urban system of multiple centres allowed applying the same logic of the model to other retail agglomeration areas.
+Resulting in finding relationships based on three different types of distance measurements.
+Firstly, the distance between Dental Pratice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+
+Furthermore, The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
+The cost of dental care in Norway can vary depending on a number of factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
+The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
+Bertrand model also take price as competitive advantage in to concideration.
+
+Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
+Prior to the most retail activities took place in the city centre.
+Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway.
+The largest shopping mall in Norway is currently Sandvika Storsenter, located just outside of Oslo with 205 shops, restaurants and entrainment venue @home-n.
+
+However, the growth of shopping malls in Norway has not been without controversy.
+Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments @kanvin2014.
+
+Information about Wine monopoly locations has also been studied in this paper.
+The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
+Subsequently, the government may take in to account the potential impact of exciting retail centres in the region.
+This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
+As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
+
 **Estimation strategy**

### 2023-02-18 MSB106_Assignment.qmd

--- 2023-02-17 MSB106_Assignment.qmd
@@
 \
+
+![](https://lh3.googleusercontent.com/zSwLHynV_h_vnX66jchxqoZgBDHH_LVE5PAOiiSLtM3fgSrVBOPT4lT3G7jP-OlsOwg65bJhrAj9rPTicCosZdGAYHySGihdbgtSzj7UICa8fQfX7oMPtO-xdHGtzP_SnGVXlWLrUUPKDKTWfJN2l00)
+
+\
+\
+
+The model shows that moving further away from a CBD and a Mall has a negative effect on te income for dentist.
+moving 1 km away from CBD reduce the income by aprox.
+528 kr and moving 1 km further away from a shopping mall reduces the income by 25 725 kr.
+This tells a story where it is more profitable beeing located close or in a Mal or a CBD where the Mall is the best option out from this model.
+The result also indicates that the best placing for a dentis would be in the second circle from the Alonso model as stated earlier, as this is the area where we are most likely to find Shopping malls.
+
+\
+
+The reasoning for a more profitable business close to a CBD can be that this is more crowded areas, gravitation pull.
+More people live here and it also contains more business and activities which can be combined with a dentist appointment.
+Somewhat of the same pattern goes for shopping malls where customers can combine their dentis visit with shopping afterwards, maybe in a way to threat them self.
+
+\
+\

### 2023-02-18 Retail Agglomeration.qmd
Commit.

@@
 
-To carry out a full-flegded data analysis of the influencing factors of the activity of Norwegian dentist services, we collected some more data than what we already had been presented with.
+To carry out a full-flegded data analysis of the influencing factors of the activity of Norwegian dentist services, we had to collect more data than what we already had been presented with.
 We firstly utilized QGIS to calculate different other statistics and variables.
@@
 
-**Empirical Specification**
+After calculations in QGIS we transfered our data to RSTUDIO to do the primary statistical analysis of our data.
+The data is then merged into two different data frames.
+The first one containing Norwegian communes as the unit of observation, which currently includes geometric commune centers to the nearest mall, the retail per capita per commune and the different commune numbers.
+The second data frame use dentists as the unit of observations.
+This data frame include data on the distances from CBD to dentists, distances from malls to dentists and different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
+
+**mpirical Specification**
 

### 2023-02-18 MSB106_Assignment.qmd
Commit.

--- 2023-02-17 MSB106_Assignment.qmd
@@
   select(fid, X_sum)
+Dist_Comp_Den <- read.csv("Dist_Comp_Den.csv") |>
+  select(fid, distance)
 NOR_KOMM <- read_csv("NOR_KOMM.csv")
@@
 Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
+```
+
+```{r}
+Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
+  rename("DistComp" = distance)
 ```
@@
 ```{r}
-kable(summary(Dentist_Data[8:12]))
+kable(summary(Dentist_Data[8:11]))
+```
+
+```{r}
+kable(summary(Dentist_Data[11:14]))
 ```
@@
 lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
-lm2 <- lm(Sum.salgsi ~ DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
+lm2 <- lm(Sum.salgsi ~ DistComp + DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
 ```

### 2023-02-18 Retail Agglomerationextra.qmd
seb

@@
 The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
-
-![](images/bidrent.png)
 
@@
 Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
-For exsampel faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
+For example faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
 As distance increases, communication networks become important allowing the growth of sub centres to functions as self organising systems.
@@
 Firstly, the distance between Dental Pratice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
+Additionally, dentist also located close to eachother making some sort of industrial cluster.
+Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
 
-Furthermore, The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
-The cost of dental care in Norway can vary depending on a number of factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
+Subsequently, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
+Furthermore, decreasing as the distance gets bigger.
+However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
+The Keynesian theory suggest that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
+Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
+
+According to the Alons model dentist should locate close to the CBD for higher volume of sales.
+Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.
+However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
+Our findings in QGIS shows that in areas where it is a high population density, the number of dentist also is higher.
+The user/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
+Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation
+cost to the second closes dentist are noticeable higher.
+
+Furthermore, it is important to be aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
+This also takes away some of the competition and would support placement in more rural areas.
+More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
+Suggesting some dentist might seek closer to CBD areas, whereas business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
+Reducing the transportation cost for the customer as they are already located in these areas even if they dont live here.
+
+The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
+Nevertheless, the cost of dental care in Norway can vary depending on a number of factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
 The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
-Bertrand model also take price as competitive advantage in to concideration.
+Bertrand model also take price as competitive advantage in to consideration.
+
+Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
+
+Another Mashiallian point to be made could be the advantage of industry clusters.
+Taken for granted that the data on Dentist includes both "normal Dentists" and specialist.
+Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
+The customer would then probably prefer to go to the closes specialist.
+The access to work force in these areas would also be good and will be within the same geographic area.
 
@@
 Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
-In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway.
-The largest shopping mall in Norway is currently Sandvika Storsenter, located just outside of Oslo with 205 shops, restaurants and entrainment venue @home-n.
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway @home-n.
 
@@
 As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
+
+Hva viser informasjonen om vinmonopolet?
 

### 2023-02-18 Retail Agglomerationextra.qmd
..

@@
 Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
-Additionally, dentist also located close to eachother making some sort of industrial cluster.
+Additionally, dentist also located close to each other making some sort of industrial cluster.
 Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
@@
 
-According to the Alons model dentist should locate close to the CBD for higher volume of sales.
+According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
 Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.

### 2023-02-18 Retail Agglomerationextra.qmd
.

@@
 Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
+
 For example faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
@@
 The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
-Having an urban system of multiple centres allowed applying the same logic of the model to other retail agglomeration areas.
+Having an urban system of multiple centres and applying the same logic of the model to other retail agglomeration areas.
 Resulting in finding relationships based on three different types of distance measurements.
-Firstly, the distance between Dental Pratice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
 Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
@@
 Furthermore, decreasing as the distance gets bigger.
+
 However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.

### 2023-02-18 Retail Agglomerationextra.qmd
..

@@
 
-For example faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
-As distance increases, communication networks become important allowing the growth of sub centres to functions as self organising systems.
-
 The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
@@
 
-Subsequently, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
+Furthermore, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
 Furthermore, decreasing as the distance gets bigger.
 
-However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
+Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
+As distance increases, communication networks become important allowing the growth of sub centres to functions as self organising systems.
+
+Subsequently, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
 The Keynesian theory suggest that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.

### 2023-02-18 Retail Agglomerationextra.qmd
..

@@
 Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
-As distance increases, communication networks become important allowing the growth of sub centres to functions as self organising systems.
-
-Subsequently, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
-The Keynesian theory suggest that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
+Hence supporting findings; when the distance reach around 10km from the mall the income starts to increase before it decrease again around 20km.
+Also in line with the Keynesian theory suggesting that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
 Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
@@
 However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
-Our findings in QGIS shows that in areas where it is a high population density, the number of dentist also is higher.
-The user/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
-Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation
-cost to the second closes dentist are noticeable higher.
+Our findings in QGIS shows that in areas of high population density, the number of dentist also is increase.
+The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
+Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation cost to the second closes dentist are noticeable higher.
 
-Furthermore, it is important to be aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
+Furthermore, being aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
 This also takes away some of the competition and would support placement in more rural areas.
 More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
-Suggesting some dentist might seek closer to CBD areas, whereas business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
+Suggesting some dentist might seek closer to CBD areas to capture most of the remaining part of the population.
+Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
 Reducing the transportation cost for the customer as they are already located in these areas even if they dont live here.
@@
 Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
+Tjener tannleger i større byer mer?
 

### 2023-02-18 Retail Agglomerationextra.qmd
...

@@
 
-The question addressed are related to the geographical space of Norway.
+The questions addressed are related to the geographical space of Norway.
 Norway is a country with relatively small population and large heterogeneous land area.
@@
 This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
+Whereas one of the assumptions are homogeneous landscape.
 Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
@@
 Furthermore, decreasing as the distance gets bigger.
-
 Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
@@
 Information about Wine monopoly locations has also been studied in this paper.
+As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
 The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
-Subsequently, the government may take in to account the potential impact of exciting retail centres in the region.
+More importanly, the government may take in to account the potential impact of exciting retail centres in the region.
 This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
-As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
 

### 2023-02-18 Retail Agglomerationextra.qmd
...

@@
 
-Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
-More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
+Our main task in this assignment is to perform an analysis of geo-spatial determinants of firm activity.
+More specifically we are to focus on the Norwegian dental industry in this regard, and see how geo-spatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
@@
 The model of potential has the capacity to measure the potential of attractiveness to a place.
+Subsequently, the gravitational force is higher in bigger cities or more heavily populated areas.
 Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
@@
 However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
-Our findings in QGIS shows that in areas of high population density, the number of dentist also is increase.
+Our findings in QGIS shows that in areas of high population density, the number of dentist also increase.
 The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
@@
 Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
-Reducing the transportation cost for the customer as they are already located in these areas even if they dont live here.
+Reducing the transportation cost for the customer as they are already located in these areas even if they do not live here.
 
 The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
-Nevertheless, the cost of dental care in Norway can vary depending on a number of factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
+Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
 The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
@@
 Tjener tannleger i større byer mer?
+Her kan vi også skrive om gravitasjons modeller.
 
@@
 The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
-More importanly, the government may take in to account the potential impact of exciting retail centres in the region.
+More importantly, the government may take in to account the potential impact of exciting retail centres in the region.
 This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.

### 2023-02-19 Retail Agglomerationextra.qmd
..
forslag

@@
 A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
-Dispite higer rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patiants to drive sales.
+Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
 The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
@@
 
+Included why and why not outliner????
+
 **Econometric Approach**

### 2023-02-19 Retail Agglomeration_Theory.qmd
Add files via upload
Har laget nytt dokument med Teori hadde problemer med å merge... Forslag her.

--- 2023-02-19 Retail Agglomerationextra.qmd
@@
 
-Our main task in this assignment is to perform an analysis of geo-spatial determinants of firm activity.
-More specifically we are to focus on the Norwegian dental industry in this regard, and see how geo-spatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
+Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
+More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
@@
 
-## Literature Review
-
-#### Previous literature on Shopping malls
-
-### Theoretical Framework
+## Theoretical Framework
 
@@
 It also involves generating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+
+Keynesian economics also emphasizes the importance of consumer demand in driving economic growth.
+In regional economics, this may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
+Subsequently, By increasing consumer demand, supporting business in a specific region and promoting economic growth or preventing economic decline @capello2016.
 
@@
 
-```{r}
-NOR_CBD <- read_csv("NOR_CBD.csv")
-Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
-Buffer_CBD <- read_csv("Buffer_CBD.csv")
-Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
-NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
-Buffer_den <- read.csv("Buffer_den.csv")
-NOR_KOMM <- read_csv("NOR_KOMM.csv")
-NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
-```
-
-```{r}
-Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
-  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning)
-```
-
-```{r}
-Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
-  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
-  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y)
-```
-
-```{r}
-kable(summary(Commune_Data))
-View(commune_data)
-```
-
-```{r}
-kable(summary(Dentist_Data[3:6]))
-```
-
-```{r}
-kable(summary(Dentist_Data[8:12]))
-```
-
-```{r}
-Commune_Data |>
-  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
-  theme_bw()
-```
-
-```{r}
-Dentist_Data |>
-  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from CBD", y = "Income") +
-  theme_bw()
-```
-
-```{r}
-Dentist_Data_LIMIT <- Dentist_Data |>
-  filter(Sum.salgsi < 100000000, DistCBD < 250,
-         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
-```
-
-```{r}
-Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
-  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
-  theme_bw()
-```
-
-```{r}
-Dentist_Data |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "EBIT") +
-  theme_bw()
-```
-
-```{r}
-Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "EBIT") +
-  theme_bw()
-```
-
 `In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
@@
 
-**Econometric Approach**
+**Estimation strategy**
+
+**Analytical framework**
+
+SEDA (Spatial Explanatory Data Analyses)
+
+Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
+The Journal of the statistics society define such data as being objects in space, which objects have physical location.
+Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
+
+The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
+Graphical techniques for examining such maps are thus a central part of the methodology.
+A number of such techniques are discussed."` @haslett1992.
+
+`What data are used?`
+
+`What has been calculated?`
+
+## **Econometric Approach**
 
@@
 
-**Estimation strategy**
-
-**Analytical framework**
-
-SEDA (Spatial Explanatory Data Analyses)
-
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
-The Journal of the statistics society define such data as being objects in space, which objects have physical location.
-Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
-
-The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
-Graphical techniques for examining such maps are thus a central part of the methodology.
-A number of such techniques are discussed."` @haslett1992.
-
-`What data are used?`
-
-`What has been calculated?`
-
-## Results
-
-## Discussion
-
 ## Conclusion
 
-## Acknowledgements
-
 ## References

### 2023-02-19 Retail Agglomeration_Theory.qmd
stuff added

@@
 Nevertheless, despite increasing the transportation cost for patients.
-Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomerations.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomeration economies.
 

### 2023-02-19 MSB106_Assignment.qmd
...

--- 2023-02-17 MSB106_Assignment.qmd
@@
 ---
-title: "MSB106 - Assignment"
+title: "MSB106_Assignment"
 author: "Sindre H. Øveraas, Alen Colakovic, Mona Lisa Jones & Sebastian M. Fløysand"
@@
 library(readr)
-library(ggplot2)
-library(kableExtra)
-library(huxtable)
 ```
@@
 Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
-Zonal_CBD <- read_csv("Zonal_CBD.csv")
+Buffer_CBD <- read_csv("Buffer_CBD.csv")
 Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
 NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
-Zonal_Den <- read.csv("Zonal_Den.csv") |>
-  select(fid, X_sum)
+Buffer_den <- read.csv("Buffer_den.csv")
 NOR_KOMM <- read_csv("NOR_KOMM.csv")
@@
 Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
-  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
-  rename("knr" = kommunenummer, "DistMal" = HubDist, "HubNameMal" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
+  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning)
 ```
@@
   select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
-  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
-```
-
-```{r}
-Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y)
 ```
@@
 
+Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
+It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
+Furthermore, the theory uses the concept of externalities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approach to economic growth @capello2011.
+
+Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
+It also involves generating geographical advantages, in terms of easy or difficult access to a particular area @capello2011.
+
+Furthermore, Keynesian economics emphasizes the importance of consumer demand in driving economic growth.
+This may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
+Subsequently, increasing consumer demand, supporting business in a specific region.
+Promoting economic growth or preventing economic decline @capello2016.
+
+Harald Hotelling's locational equilibrium is determined by a logic of profit maximization whereby each producer controls its own market area.
+Productivity advantages of cities and urban clusters with a high density of firms increase profit by attracting a larger number of potential customers, and more productive workers @capello2011.
+Furthermore, the attractiveness of a central location increase the cost of rent.
+
+Alonso's bid rent model indicates the most profitable location for firms.
+Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
+In gravitational models, the attractiveness of the retail location, represent the size of the retail centre.
+Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
+
+The model of potential has the capacity to measure the potential of attractiveness to a place.
+Bigger cities or more heavily populated areas have a stronger gravitational force.
+A possible indicator to predict places of growth potential for dentist practice or shopping malls @capello2015.
+
+**Interdependent location choice, the Hotelling's model (1929)**
+
+The model assume that given the location of producers, and given demand uniformly distributed geographically (in linear or circular form) the market is divided into areas within each of which there operates a single firm in a duopoly environment.
+Furthermore, no relocation costs and demand only depend on location choice @capello2015.
+
+The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
+One firm starts relocating closer to the other to take some of the customers in the other market area @capello2015.
+The other respond by doing the same and the game continues until both end up in the middle of the market on the broader of AB.
+The end of the game is the position where neither can increase sales volume by moving position @capello2015.
+
+A simple explanation of why two dentists providing the same service, at the same price might locate next to each other.
+Nevertheless, despite increasing the transportation cost for patients.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomeration economies.
+
+**Marshall's agglomeration principles**
+
+Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
+Firstly, knowledge spill over within industries or product specific technological knowledge.
+Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
+Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021Nielsen.
+
+There are solidly established conclusions regarding the excitanse of agglomeration economies @puga2010Puga.
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
+Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
+@duranton2004Duranton discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
+Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
+The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
+Hence, the larger the population sharing facilities the lower the cost per user @puga2010Puga.
+Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
+
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010Puga.
+Shopping malls require skilled workers to drive sales.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020Davis.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010Puga.
+Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
+That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010Puga.
+
+*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* @puga2010Puga
+
+**Urban location of dentist; The Alonso model**
+
+The model demonstrate geographical locations tied up to location.
+It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
+Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
+Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
+Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
+As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
+More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
+@capello2015.
+The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
+Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
+The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
+
+The illustration shows three categories: The most expensive rent closest to the centre is the Bank and finance district, followed by the retail agglomeration areas and industrial sites furthest away from the central point.
+Dental practices are service orientated thus depending on population density and central activities.
+Opening hours are relay outside of office hours.
+Production cost does not appear to be important in this case since dentist is a public service.
+Hence, not dependent of regular transportation of sales goods.
+The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
+From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
+Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
+The attractiveness might be high in these places due to economies of scope.
+As the positive benefits from the Marshallian agglomeration attributes play out.
+Subsequently, increasing profit by trading higher transportation cost for lower rent.
+In addition fewer dentists to compete with.
+The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
+
+**Gravitational models**
+
+The centre attracts firms, people and activities.
+Furthermore, it influence the central points in diverse ways such as commuter movements, diffusion of knowledge, corporations and network and personal relationship @capello2015.
+More-so, they are generated by the intensity of the flows generated between places @capello2015.
+The gravitational pull to a place of interest for a dentist will determine how much sales can be generated by locating there.
+Several locations can then be measures against each other to determine profit maximisation.
+These models has the potential to measures the future attractiveness of a place.
+Changes in infrastructure such as easier access to a place might change the attractiveness of a place or move the flow of people.
+An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
+Subserviently, increasing the gravitationally pull between the two places.
+Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
+Rogfast might move the population of Haugesund further south and a new town centre might occur competing with other retail clusters in years to come.
+On the other hand, the flow of people towards Raglamyr might increase until the potential new centre place has grown.
+These models have predictive capacity and can be used to estimate the potential impact.
+In particular, location of new productive activity in an area @capello2015.
+For instance predict where to build a shopping centre @capello2015.
+A new shopping centre further south in Haugaland or expansion of excising retail agglomeration areas might be of interest.
+As a result of increased population.
+By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
+
+**Retail Agglomeration and Economies of Shopping Centres**
+
+Town and cities have long had shopping districts in which stores are concentrated.
+To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
+Nevertheless, not coincidentally according to Hotelling's theory (1929).
+He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
+At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
+Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
+Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
+That being,intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
+In this case the dental industry information spillovers are bounded in geographical space.
+Hence, the need to locate in close proximity.
+That being the latest equipment, in order to work faster or the right personnel to attract more patients.
+Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
+
+Another important externalities expressed in Marshall (1920) theory is market transactions.
+Whereas, geographical proximity reduces transaction cost through several mechanism @nielsen2021.
+The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
+"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
+Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
+In both cases growth potential would be compromised.
+
+Marshal also points out the competition within the geographical location.
+Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
+The fact that owners of stores seem to prefer highly concentrated shopping.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
+
+**Shopping Malls**
+
+Owners of shopping malls "orchestrate" the process of retail agglomeration that happens naturally in towns and cities @brueckner2011.
+Naturally, the same retail agglomeration and Nash equilibrium may occur in shopping malls if not more intensely as the shops are in closer proximity.
+Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.
+When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
+Much so, because the price of rent is dependent on the overall revenue of a shopping mall.
+Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
+The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
+Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013Lai.
+For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
+Nevertheless, they are limited in numbers and needs government approval.
+Each municipality must apply to get the wine monopoly in their area.
+One of the factors considered in this process is the death of city centre.
+A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
+Perhaps more so if malls cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
+Not to mention the upper hand of easy road access and free parking.
+
 ## Data description
 
-```{r}
-kable(summary(Commune_Data))
-```
+## Econometric approach
 
-```{r}
-kable(summary(Dentist_Data[3:6]))
-```
+The questions addressed are related to the geographical space of Norway.
+Norway is a country with relatively small population and large heterogeneous land area.
+Furthermore, the population is decentralized with the majority of population living in urban areas.
+This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
+Whereas one of the assumptions are homogeneous landscape.
+Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
+Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
 
-```{r}
-kable(summary(Dentist_Data[8:12]))
-```
+The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
+Having an urban system of multiple centres and applying the same logic of the model to other retail agglomeration areas.
+Resulting in finding relationships based on three different types of distance measurements.
+Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
+Additionally, dentist also located close to each other making some sort of industrial cluster.
+Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
 
-```{r}
-Commune_Data |>
-  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
-  theme_bw()
-```
+Furthermore, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
+Furthermore, decreasing as the distance gets bigger.
+Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
+Hence supporting findings; when the distance reach around 10km from the mall the income starts to increase before it decrease again around 20km.
+Also in line with the Keynesian theory suggesting that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
+Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
 
-Might be more appropriate to explore the distance from CBDs to Malls compared to turnover capita in retail per inhabitant, rather than the distance from the geographical centroid of Norwegian Communes, because of the tendency of bigger population density around CBDs.
+According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
+Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.
+However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
+Our findings in QGIS shows that in areas of high population density, the number of dentist also increase.
+The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
+Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation cost to the second closes dentist are noticeable higher.
 
-```{r}
-Dentist_Data |>
-  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from CBD", y = "Income") +
-  theme_bw()
-```
+Furthermore, being aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
+This also takes away some of the competition and would support placement in more rural areas.
+More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
+Suggesting some dentist might seek closer to CBD areas to capture most of the remaining part of the population.
+Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
+Reducing the transportation cost for the customer as they are already located in these areas even if they do not live here.
 
-```{r}
-Dentist_Data_LIMIT <- Dentist_Data |>
-  filter(Sum.salgsi < 100000000, DistCBD < 250,
-         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
-```
+The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
+Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
+The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
+Bertrand model also take price as competitive advantage in to consideration.
 
-```{r}
-Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
-  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
-  theme_bw()
-```
+Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
+Tjener tannleger i større byer mer?
+Her kan vi også skrive om gravitasjons modeller.
 
-```{r}
-Dentist_Data |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "EBIT") +
-  theme_bw()
-```
+Another Mashiallian point to be made could be the advantage of industry clusters.
+Taken for granted that the data on Dentist includes both "normal Dentists" and specialist.
+Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
+The customer would then probably prefer to go to the closes specialist.
+The access to work force in these areas would also be good and will be within the same geographic area.
 
-```{r}
-Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
-  geom_point(lwd = .9) +
-  labs(x = "Distance from Mall", y = "EBIT") +
-  theme_bw()
-```
+Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
+Prior to the most retail activities took place in the city centre.
+Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway @home-n.
 
-```{r}
-Dentist_Data_LIMIT |>
-  ggplot(aes(x = X_sum, y = Sum.salgsi, colour = Sum.salgsi)) +
-  geom_point(lwd = .9) +
-  labs(x = "Pop density", y = "Income") +
-  theme_bw()
-```
+However, the growth of shopping malls in Norway has not been without controversy.
+Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments @kanvin2014.
 
-```{r}
-lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
-lm2 <- lm(Sum.salgsi ~ DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
-```
+Information about Wine monopoly locations has also been studied in this paper.
+As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
+The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
+More importantly, the government may take in to account the potential impact of exciting retail centres in the region.
+This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
 
-```{r}
-huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
-```
-
-## Econometric approach
+Hva viser informasjonen om vinmonopolet?

### 2023-02-19 Retail Agglomeration_Theory.qmd
...

@@
 It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities as disparities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
+Furthermore, the theory uses the concept of externalities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approach to economic growth @capello2011.
 
 Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
+It also involves generating geographical advantages, in terms of easy or difficult access to a particular area @capello2011.
 
-Keynesian economics also emphasizes the importance of consumer demand in driving economic growth.
-In regional economics, this may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
-Subsequently, By increasing consumer demand, supporting business in a specific region and promoting economic growth or preventing economic decline @capello2016.
+Furthermore, Keynesian economics emphasizes the importance of consumer demand in driving economic growth.
+This may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
+Subsequently, increasing consumer demand, supporting business in a specific region.
+Promoting economic growth or preventing economic decline @capello2016.
 
@@
 The model of potential has the capacity to measure the potential of attractiveness to a place.
-Subsequently, the gravitational force is higher in bigger cities or more heavily populated areas.
-Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
+Bigger cities or more heavily populated areas have a stronger gravitational force.
+A possible indicator to predict places of growth potential for dentist practice or shopping malls @capello2015.
 
@@
 Furthermore, no relocation costs and demand only depend on location choice @capello2015.
-
 The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
@@
 
+**Hotelling-Bertand model (1979)**
+
+The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
+Furthermore, the consumers choose the product with the lowest price.
+The model assumes that firms have identical production cost and consumers have perfect information about prices.
+
+By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979.
+In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products.
+
 **Marshall's agglomeration principles**
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
-Firstly, knowledge spill over within industries or product specific technological knowledge.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially drive sales.
+First, knowledge spill-over within industries or product specific technological knowledge.
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
@@
 
-There are solidly established conclusions regarding the excitanse of agglomeration economies @puga2010Puga.
-However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
+There are solidly established conclusions regarding the exsitance of agglomeration economies @puga2010Puga.
+However, less proof of their estimated magnitude.
+Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010Puga.
 Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
-@duranton2004Duranton discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+@duranton2004Duranton discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
-Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
+A larger market allows for a more efficient sharing of local infrastructure and facilities.
+Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
+Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients in order to drive sales.
 The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
@@
 
-*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* @puga2010Puga
+Interactions with experienced workers helps acquire valuable skills.
+Experienced workers remain in cities to share the rent of this learning process @puga2010.
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the unintended casual flow of information facilitated by big cities." @puga2010Puga
 
@@
 
-The model demonstrate geographical locations tied up to location.
-It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
-Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
+The Alsonso model demonstrate geographical locations tied up to location.
+Furthermore, It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
+In this case used to get an indicating of where the most profitable location for dental practise.
+
 Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
-Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
+Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain.
+In some cases different from the normal or average price competition @capello2015.
+
 As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
-More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
+More-so, the town or city has a single centre point.
+The city centre of business district (CBD) is generally defined as the most attractive for all firms.
 @capello2015.
 The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
-Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
+Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.
+
+Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
 The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
+
+![](images/bidrent.gpg.jpg)
 
@@
 The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
+
 From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
-Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
+Furthermore, they may also increase their profits further out in to other retail agglomeration area.
 The attractiveness might be high in these places due to economies of scope.
@@
 Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with.
-The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
+In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013McCann.
 
@@
 
-Marshal also points out the competition within the geographical location.
+Marshall also points out the competition within the geographical location.
 Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition@capello2015.
 
@@
 
-According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
+According to the Alons's model dentist should locate in the retail area for higher volume of sales.
 Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.

### 2023-02-19 Retail Agglomeration_Theory.qmd
...

@@
 
+`The topic of research (What was studied)`
+
+`The purpose (What was done)`
+
+`The methodology (how the research was done)`
+
+`Main findings (what was found)`
+
+`Policy and implications (significance of findings)`
+
 ## Introduction
@@
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
+
+`A brief description of the problem and its importance`
+
+`Statement of purpose or what was done`
+
+`A brief description of main results`
+
+`Description of methodology, main challenges and solutions`
+
+`Discretion of the papers contribution in relation to previous research`
+
+`Organization of paper`
 
@@
 Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially drive sales.
-First, knowledge spill-over within industries or product specific technological knowledge.
+Firstly, knowledge spill-over within industries or product specific technological knowledge.
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
@@
 
-There are solidly established conclusions regarding the exsitance of agglomeration economies @puga2010Puga.
+There are solidly established conclusions regarding the existence of agglomeration economies @puga2010Puga.
 However, less proof of their estimated magnitude.
@@
 Several locations can then be measures against each other to determine profit maximisation.
+
 These models has the potential to measures the future attractiveness of a place.
 Changes in infrastructure such as easier access to a place might change the attractiveness of a place or move the flow of people.
-An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
 Subserviently, increasing the gravitationally pull between the two places.
 Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
-Rogfast might move the population of Haugesund further south and a new town centre might occur competing with other retail clusters in years to come.
-On the other hand, the flow of people towards Raglamyr might increase until the potential new centre place has grown.
-These models have predictive capacity and can be used to estimate the potential impact.
+Furthermore, these models have predictive capacity and can be used to estimate the potential impact.
 In particular, location of new productive activity in an area @capello2015.
 For instance predict where to build a shopping centre @capello2015.
-A new shopping centre further south in Haugaland or expansion of excising retail agglomeration areas might be of interest.
-As a result of increased population.
 By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
@@
 The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
+"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." @bakerBaker.
+Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
 Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
@@
 Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
+
 The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.

### 2023-02-19 MSB106_Assignment.qmd
...

@@
 
+**Hotelling-Bertand model (1979)**
+
+The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
+Furthermore, the consumers choose the product with the lowest price.
+The model assumes that firms have identical production cost and consumers have perfect information about prices.
+
+By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979.
+In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products.
+
 **Marshall's agglomeration principles**
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
-Firstly, knowledge spill over within industries or product specific technological knowledge.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially drive sales.
+Firstly, knowledge spill-over within industries or product specific technological knowledge.
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
@@
 
-There are solidly established conclusions regarding the excitanse of agglomeration economies @puga2010Puga.
-However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
+There are solidly established conclusions regarding the existence of agglomeration economies @puga2010Puga.
+However, less proof of their estimated magnitude.
+Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010Puga.
 Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
-@duranton2004Duranton discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
+@duranton2004Duranton discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
-Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
+A larger market allows for a more efficient sharing of local infrastructure and facilities.
+Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
+Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients in order to drive sales.
 The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
@@
 
-Marshal also points out the competition within the geographical location.
+Marshall also points out the competition within the geographical location.
 Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
 

### 2023-02-19 Retail Agglomeration_Theory.qmd
Manlende referanser

@@
 The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
-Furthermore, the consumers choose the product with the lowest price.
-The model assumes that firms have identical production cost and consumers have perfect information about prices.
+Furthermore, the consumers choose the product with the lowest price@tolotti2020.
+The model assumes that firms have identical production cost and consumers have perfect information about prices @tolotti2020.
 
-By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979.
-In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products.
+By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979 @tolotti2020.
+In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products @tolotti2020.
 
@@
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition@capello2015.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
 

### 2023-02-19 MSB106_Assignment.qmd
Commit.

--- 2023-02-18 MSB106_Assignment.qmd
@@
   select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
-  rename("knr" = kommunenummer, "DistMal" = HubDist, "HubNameMal" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
+  rename("knr" = kommunenummer, "DistMal" = HubDist, "fid" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
+```
+
+```{r}
+Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
+  select(knr, DistMal, fid, Turnover_capita_retail, shop, Vinmonopolet) |>
+  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
+```
+
+```{r}
+Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
 ```
@@
   rename("DistComp" = distance)
+```
+
+```{r}
+Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
+  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
+  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
+```
+
+```{r}
+Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
 ```
@@
 ```{r}
-kable(summary(Dentist_Data[8:11]))
+kable(summary(Dentist_Data[7:10]))
 ```
@@
 ```{r}
-kable(summary(Dentist_Data[11:14]))
+kable(summary(Dentist_Data[11:12]))
 ```
@@
   filter(Sum.salgsi < 100000000, DistCBD < 250,
-         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
+         Sum.Drifts < 20000000, Sum.Drifts > -10000000, DistMal < 80)
 ```
@@
 Dentist_Data |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  ggplot(aes(x = DistMal, y = Sum.Drifts, colour = DistMal)) +
   geom_point(lwd = .9) +
@@
 Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  ggplot(aes(x = DistMal, y = Sum.Drifts, colour = DistMal)) +
   geom_point(lwd = .9) +
@@
 ```{r}
-lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
-lm2 <- lm(Sum.salgsi ~ DistComp + DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
+lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
 ```
@@
 ```{r}
-huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+huxreg(list("Dentist Sales" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
 ```

### 2023-02-19 Retail Agglomeration.qmd
Commit.

@@
 We firstly utilized QGIS to calculate different other statistics and variables.
-Regarding the Norwegian shopping malls layer we added the variable \"Wine-monopoly\" as a false/ true variable to determine which Norwegian malls, and also the number of stores which are in each mall.
+Regarding the Norwegian shopping malls layer we added the variable "Wine-monopoly" as a false/ true variable to determine which Norwegian malls, and also the number of stores which are in each mall.
 The malls opening hours have also been included in this layer.

### 2023-02-19 Retail Agglomeration_Theory.qmd
..

@@
 The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
-Furthermore, the consumers choose the product with the lowest price@tolotti2020.
+Furthermore, the consumers choose the product with the lowest price\@tolotti2020.
 The model assumes that firms have identical production cost and consumers have perfect information about prices @tolotti2020.
@@
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021Nielsen.
+Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021.
 
-There are solidly established conclusions regarding the existence of agglomeration economies @puga2010Puga.
-However, less proof of their estimated magnitude.
-Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010Puga.
-Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
-@duranton2004Duranton discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
+There are solidly established conclusions regarding the existence of agglomeration economies @puga2010. However, less proof of their estimated magnitude.
+Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010.
+Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010.
+@duranton2004 discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
 
 A larger market allows for a more efficient sharing of local infrastructure and facilities.
-Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
+Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010.
 Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients in order to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
-Hence, the larger the population sharing facilities the lower the cost per user @puga2010Puga.
+The attraction for the consumers and users of public facilities is overall cost reduction @puga2010.
+Hence, the larger the population sharing facilities the lower the cost per user @puga2010.
 Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
 
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010Puga.
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010.
 Shopping malls require skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020Davis.
-More-so, cites provide a constant market for specialized skills and more productive work force @puga2010Puga.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010.
 Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
 That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010Puga.
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010.
 
@@
 Experienced workers remain in cities to share the rent of this learning process @puga2010.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the unintended casual flow of information facilitated by big cities." @puga2010Puga
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the unintended casual flow of information facilitated by big cities." @puga2010
 
@@
 Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013McCann.
+In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013.
 
@@
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition @capello2015.
 

### 2023-02-19 Retail Agglomeration_Theory.qmd
...
sorry hvis eg har ødelagt...

@@
 
-There are solidly established conclusions regarding the existence of agglomeration economies @puga2010. However, less proof of their estimated magnitude.
+There are solidly established conclusions regarding the existence of agglomeration economies @puga2010.
+However, less proof of their estimated magnitude.
 Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010.
@@
 However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020.
-More-so, cites provide a constant market for specialized skills and more productive work force @puga2010.
-Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010P Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
 That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
@@
 Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013.
+In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013McCann.
 
@@
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition @capello2015.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
 

### 2023-02-19 Retail-Agglomeration_Theory.tex
sorry hvis eg har ødelagt...

--- 2023-02-19 Retail Agglomeration_Theory.qmd
@@
----
-title: "Regional and Urban Economics"
-subtitle: Dental industry
-author: "Sindre Halvorsen Øveraas , Sebastian Mena Fløysand, Alen Colakovic & Mona Lisa Jones"
-format:
- pdf:
-  fontsize: "10"
-  documentclass: article
-  margin-left: 2cm
-  margin-right: 2cm
-papersize: A4
-editor: visual
-bibliography: references.bib
----
+% Options for packages loaded elsewhere
+\PassOptionsToPackage{unicode}{hyperref}
+\PassOptionsToPackage{hyphens}{url}
+\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
+%
+\documentclass[
+  10,
+  a4paper,
+]{article}
 
-## Abstract
+\usepackage{amsmath,amssymb}
+\usepackage{lmodern}
+\usepackage{iftex}
+\ifPDFTeX
+  \usepackage[T1]{fontenc}
+  \usepackage[utf8]{inputenc}
+  \usepackage{textcomp} % provide euro and other symbols
+\else % if luatex or xetex
+  \usepackage{unicode-math}
+  \defaultfontfeatures{Scale=MatchLowercase}
+  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
+\fi
+% Use upquote if available, for straight quotes in verbatim environments
+\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
+\IfFileExists{microtype.sty}{% use microtype if available
+  \usepackage[]{microtype}
+  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
+}{}
+\makeatletter
+\@ifundefined{KOMAClassName}{% if non-KOMA class
+  \IfFileExists{parskip.sty}{%
+    \usepackage{parskip}
+  }{% else
+    \setlength{\parindent}{0pt}
+    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
+}{% if KOMA class
+  \KOMAoptions{parskip=half}}
+\makeatother
+\usepackage{xcolor}
+\usepackage[lmargin=2cm,rmargin=2cm]{geometry}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
+% Make \paragraph and \subparagraph free-standing
+\ifx\paragraph\undefined\else
+  \let\oldparagraph\paragraph
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+\fi
+\ifx\subparagraph\undefined\else
+  \let\oldsubparagraph\subparagraph
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+\fi
 
-`The topic of research (What was studied)`
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{calc} % for calculating minipage widths
+% Correct order of tables after \paragraph or \subparagraph
+\usepackage{etoolbox}
+\makeatletter
+\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
+\makeatother
+% Allow footnotes in longtable head/foot
+\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
+\makesavenoteenv{longtable}
+\usepackage{graphicx}
+\makeatletter
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+% Set default figure placement to htbp
+\makeatletter
+\def\fps@figure{htbp}
+\makeatother
+\newlength{\cslhangindent}
+\setlength{\cslhangindent}{1.5em}
+\newlength{\csllabelwidth}
+\setlength{\csllabelwidth}{3em}
+\newlength{\cslentryspacingunit} % times entry-spacing
+\setlength{\cslentryspacingunit}{\parskip}
+\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
+ {% don't indent paragraphs
+  \setlength{\parindent}{0pt}
+  % turn on hanging indent if param 1 is 1
+  \ifodd #1
+  \let\oldpar\par
+  \def\par{\hangindent=\cslhangindent\oldpar}
+  \fi
+  % set entry spacing
+  \setlength{\parskip}{#2\cslentryspacingunit}
+ }%
+ {}
+\usepackage{calc}
+\newcommand{\CSLBlock}[1]{#1\hfill\break}
+\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
+\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
+\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
 
-`The purpose (What was done)`
+\makeatletter
+\makeatother
+\makeatletter
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\AtBeginDocument{%
+\ifdefined\contentsname
+  \renewcommand*\contentsname{Table of contents}
+\else
+  \newcommand\contentsname{Table of contents}
+\fi
+\ifdefined\listfigurename
+  \renewcommand*\listfigurename{List of Figures}
+\else
+  \newcommand\listfigurename{List of Figures}
+\fi
+\ifdefined\listtablename
+  \renewcommand*\listtablename{List of Tables}
+\else
+  \newcommand\listtablename{List of Tables}
+\fi
+\ifdefined\figurename
+  \renewcommand*\figurename{Figure}
+\else
+  \newcommand\figurename{Figure}
+\fi
+\ifdefined\tablename
+  \renewcommand*\tablename{Table}
+\else
+  \newcommand\tablename{Table}
+\fi
+}
+\@ifpackageloaded{float}{}{\usepackage{float}}
+\floatstyle{ruled}
+\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
+\floatname{codelisting}{Listing}
+\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{caption}{}{\usepackage{caption}}
+\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
+\makeatother
+\makeatletter
+\@ifpackageloaded{tcolorbox}{}{\usepackage[many]{tcolorbox}}
+\makeatother
+\makeatletter
+\@ifundefined{shadecolor}{\definecolor{shadecolor}{rgb}{.97, .97, .97}}
+\makeatother
+\makeatletter
+\makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
+\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
+\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
+\urlstyle{same} % disable monospaced font for URLs
+\hypersetup{
+  pdftitle={Regional and Urban Economics},
+  pdfauthor={Sindre Halvorsen Øveraas , Sebastian Mena Fløysand, Alen Colakovic \& Mona Lisa Jones},
+  colorlinks=true,
+  linkcolor={blue},
+  filecolor={Maroon},
+  citecolor={Blue},
+  urlcolor={Blue},
+  pdfcreator={LaTeX via pandoc}}
 
-`The methodology (how the research was done)`
+\title{Regional and Urban Economics}
+\usepackage{etoolbox}
+\makeatletter
+\providecommand{\subtitle}[1]{% add subtitle to \maketitle
+  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
+}
+\makeatother
+\subtitle{Dental industry}
+\author{Sindre Halvorsen Øveraas , Sebastian Mena Fløysand, Alen
+Colakovic \& Mona Lisa Jones}
+\date{}
 
-`Main findings (what was found)`
+\begin{document}
+\maketitle
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
 
-`Policy and implications (significance of findings)`
+\hypertarget{abstract}{%
+\subsection{Abstract}\label{abstract}}
 
-## Introduction
+\texttt{The\ topic\ of\ research\ (What\ was\ studied)}
 
-Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
-More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
-As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
+\texttt{The\ purpose\ (What\ was\ done)}
 
-`A brief description of the problem and its importance`
+\texttt{The\ methodology\ (how\ the\ research\ was\ done)}
 
-`Statement of purpose or what was done`
+\texttt{Main\ findings\ (what\ was\ found)}
 
-`A brief description of main results`
+\texttt{Policy\ and\ implications\ (significance\ of\ findings)}
 
-`Description of methodology, main challenges and solutions`
+\hypertarget{introduction}{%
+\subsection{Introduction}\label{introduction}}
 
-`Discretion of the papers contribution in relation to previous research`
+Our main task in this assignment is to perform an analysis of geospatial
+determinants of firm activity. More specifically we are to focus on the
+Norwegian dental industry in this regard, and see how geospatial
+determinants such as distances to shopping malls and CBDs (Central
+Business Districts), as well as population density can determine dental
+businesses income and general financial operations. As an example,
+central questions in this assignment will be; ``Is it more beneficial to
+be highly centralized in urban areas with high population density and
+many competitors, or is it a greater advantage to be less centralized to
+the advantage that the nearest competing company is considerably further
+away?'', ``Which determinants appear to be most significant for economic
+benefit?''
 
-`Organization of paper`
+\texttt{A\ brief\ description\ of\ the\ problem\ and\ its\ importance}
 
-**Hypothesis**
+\texttt{Statement\ of\ purpose\ or\ what\ was\ done}
 
-The location choice of firms providing service to end consumers, significantly determine their ultimate growth potential.
+\texttt{A\ brief\ description\ of\ main\ results}
 
-**Research question**
+\texttt{Description\ of\ methodology,\ main\ challenges\ and\ solutions}
+
+\texttt{Discretion\ of\ the\ papers\ contribution\ in\ relation\ to\ previous\ research}
+
+\texttt{Organization\ of\ paper}
+
+\textbf{Hypothesis}
+
+The location choice of firms providing service to end consumers,
+significantly determine their ultimate growth potential.
+
+\textbf{Research question}
 
@@
 
-## Theoretical Framework
+\hypertarget{theoretical-framework}{%
+\subsection{Theoretical Framework}\label{theoretical-framework}}
 
-Location theory gives regional economics its scientific disciplinary identity and constitutes its theoretical methodological core @capello2011.
-It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approach to economic growth @capello2011.
+Location theory gives regional economics its scientific disciplinary
+identity and constitutes its theoretical methodological core
+(\textbf{capello2011?}). It has typically microeconomic foundation and
+uses theoretical models as well as adopting a statistical and
+geographical approach (\textbf{capello2011?}). Furthermore, the theory
+uses the concept of externalities in the spatial distribution of
+activities, thereby laying the territorial bases for dynamic approach to
+economic growth (\textbf{capello2011?}).
 
-Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generating geographical advantages, in terms of easy or difficult access to a particular area @capello2011.
+Regional growth theory involves spatial aspects of economic growth and
+territorial distribution of income (\textbf{capello2011?}). It also
+involves generating geographical advantages, in terms of easy or
+difficult access to a particular area (\textbf{capello2011?}).
 
-Furthermore, Keynesian economics emphasizes the importance of consumer demand in driving economic growth.
-This may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
-Subsequently, increasing consumer demand, supporting business in a specific region.
-Promoting economic growth or preventing economic decline @capello2016.
+Furthermore, Keynesian economics emphasizes the importance of consumer
+demand in driving economic growth. This may involve policies that
+involve consumer spending, such as incentives for buying local products
+or supporting small businesses or in this case preventing the the death
+of down town due to competition from shopping malls. Subsequently,
+increasing consumer demand, supporting business in a specific region.
+Promoting economic growth or preventing economic decline
+(\textbf{capello2016?}).
 
-Harald Hotelling's locational equilibrium is determined by a logic of profit maximization whereby each producer controls its own market area.
-Productivity advantages of cities and urban clusters with a high density of firms increase profit by attracting a larger number of potential customers, and more productive workers @capello2011.
-Furthermore, the attractiveness of a central location increase the cost of rent.
+Harald Hotelling's locational equilibrium is determined by a logic of
+profit maximization whereby each producer controls its own market area.
+Productivity advantages of cities and urban clusters with a high density
+of firms increase profit by attracting a larger number of potential
+customers, and more productive workers (\textbf{capello2011?}).
+Furthermore, the attractiveness of a central location increase the cost
+of rent.
 
-Alonso's bid rent model indicates the most profitable location for firms.
-Closer to the centre with agglomeration attributes or in rural areas with spacial monopoly and low rent.
-In gravitational models, the attractiveness of the retail location, represent the size of the retail centre.
-Furthermore, it depends on the variety of goods which can be purchased at the same location @mccann2013.
+Alonso's bid rent model indicates the most profitable location for
+firms. Closer to the centre with agglomeration attributes or in rural
+areas with spacial monopoly and low rent. In gravitational models, the
+attractiveness of the retail location, represent the size of the retail
+centre. Furthermore, it depends on the variety of goods which can be
+purchased at the same location (\textbf{mccann2013?}).
 
-The model of potential has the capacity to measure the potential of attractiveness to a place.
-Bigger cities or more heavily populated areas have a stronger gravitational force.
-A possible indicator to predict places of growth potential for dentist practice or shopping malls @capello2015.
+The model of potential has the capacity to measure the potential of
+attractiveness to a place. Bigger cities or more heavily populated areas
+have a stronger gravitational force. A possible indicator to predict
+places of growth potential for dentist practice or shopping malls
+(\textbf{capello2015?}).
 
-**Interdependent location choice, the Hotelling's model (1929)**
+\textbf{Interdependent location choice, the Hotelling's model (1929)}
 
-The model assume that given the location of producers, and given demand uniformly distributed geographically (in linear or circular form) the market is divided into areas within each of which there operates a single firm in a duopoly environment.
-Furthermore, no relocation costs and demand only depend on location choice @capello2015.
-The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
-One firm starts relocating closer to the other to take some of the customers in the other market area @capello2015.
-The other respond by doing the same and the game continues until both end up in the middle of the market on the broader of AB.
-The end of the game is the position where neither can increase sales volume by moving position @capello2015.
+The model assume that given the location of producers, and given demand
+uniformly distributed geographically (in linear or circular form) the
+market is divided into areas within each of which there operates a
+single firm in a duopoly environment. Furthermore, no relocation costs
+and demand only depend on location choice (\textbf{capello2015?}). The
+location game starts off with the total market of AB, firm A in the
+middle of location A and firm B starts in the middle of location B. One
+firm starts relocating closer to the other to take some of the customers
+in the other market area (\textbf{capello2015?}). The other respond by
+doing the same and the game continues until both end up in the middle of
+the market on the broader of AB. The end of the game is the position
+where neither can increase sales volume by moving position
+(\textbf{capello2015?}).
 
-A simple explanation of why two dentists providing the same service, at the same price might locate next to each other.
-Nevertheless, despite increasing the transportation cost for patients.
-Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomeration economies.
+A simple explanation of why two dentists providing the same service, at
+the same price might locate next to each other. Nevertheless, despite
+increasing the transportation cost for patients. Perhaps the simplest
+way to explain why there is a natural tendency for retailers to cluster
+in space; a tendency which may help explain the existence of larger
+agglomeration economies.
 
-**Hotelling-Bertand model (1979)**
+\textbf{Hotelling-Bertand model (1979)}
 
-The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
-Furthermore, the consumers choose the product with the lowest price@tolotti2020.
-The model assumes that firms have identical production cost and consumers have perfect information about prices @tolotti2020.
+The Bertrand model was introduced as early as in 1883 and demonstrate
+two firms competing by simultaneously setting prices for their
+homogeneous products. Furthermore, the consumers choose the product with
+the lowest price@tolotti2020. The model assumes that firms have
+identical production cost and consumers have perfect information about
+prices (\textbf{tolotti2020?}).
 
-By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979 @tolotti2020.
-In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products @tolotti2020.
+By combining elements from the two models a hybrid model of spatial
+completion was birthed by Salop in 1979 (\textbf{tolotti2020?}). In this
+model firms compete both in location and price, with consumers having a
+transportation cost and firms producing differentiated products
+(\textbf{tolotti2020?}).
 
-**Marshall's agglomeration principles**
+\textbf{Marshall's agglomeration principles}
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially drive sales.
-Firstly, knowledge spill-over within industries or product specific technological knowledge.
-Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021Nielsen.
+Marshall (1920) broadly divides externalities within agglomeration in
+three main categories potentially drive sales. Firstly, knowledge
+spill-over within industries or product specific technological
+knowledge. Furthermore, market transactions in terms of value chain
+transactions with industry-specialized buyers and suppliers. Lastly,
+competition for specialized production factors such as labour and
+product market competition (\textbf{nielsen2021?}).
 
-There are solidly established conclusions regarding the existence of agglomeration economies @puga2010Puga.
-However, less proof of their estimated magnitude.
-Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010Puga.
-Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
-@duranton2004Duranton discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
+There are solidly established conclusions regarding the existence of
+agglomeration economies (\textbf{puga2010?}). However, less proof of
+their estimated magnitude. Hence, identifying the causes of
+agglomeration economies, is proving more difficult (\textbf{puga2010?}).
+Nevertheless, there is a large theoretical literature that develops
+these mechanisms (\textbf{puga2010?}). (\textbf{duranton2004?}) discuss
+these classifications and identify \emph{learning, sharing} and
+\emph{matching} as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities.
-Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
-Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients in order to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
-Hence, the larger the population sharing facilities the lower the cost per user @puga2010Puga.
-Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
+A larger market allows for a more efficient sharing of local
+infrastructure and facilities. Therefore, a variety of intermediate
+input suppliers, or a pool of workers with similar skills
+(\textbf{puga2010?}). Despite higher rent the dental industry and
+shopping malls, seams to reap higher benefits in more populated areas as
+they are dependent on being located where there is a higher volume of
+patients in order to drive sales. The attraction for the consumers and
+users of public facilities is overall cost reduction
+(\textbf{puga2010?}). Hence, the larger the population sharing
+facilities the lower the cost per user (\textbf{puga2010?}). Presumably,
+industrial factories and business clusters are more dependent on being
+close to raw materials and industrial action.
 
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010Puga.
-Shopping malls require skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020Davis.
-More-so, cites provide a constant market for specialized skills and more productive work force @puga2010Puga.
-Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
-That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010Puga.
+Furthermore, a larger market also allows for a better matching between
+employers and employees, improved chances of finding suitable and better
+quality of matches (\textbf{puga2010?}). Shopping malls require skilled
+workers to drive sales. However, they are not so dependent on highly
+educated workers as dentists whom according to resent study, tend to
+prefer bigger cities (as they are highly educated)
+(\textbf{davis2020?}). More-so, cites provide a constant market for
+specialized skills and more productive work force (\textbf{puga2010P?})
+Perhaps, a possibility for higher wages being compensated by more
+efficient workers in bigger agglomeration economies. That said the
+Norwegian cities might differ from the american cites in the study, as
+the population and clusters are no where near the same size. Lastly, a
+larger market can also facilitate learning by promoting the development
+and widespread adaptation of new technologies and business practice
+(\textbf{puga2010?}).
 
 Interactions with experienced workers helps acquire valuable skills.
-Experienced workers remain in cities to share the rent of this learning process @puga2010.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the unintended casual flow of information facilitated by big cities." @puga2010Puga
+Experienced workers remain in cities to share the rent of this learning
+process (\textbf{puga2010?}). Besides this purposeful transition of
+knowledge, the information literature on learning in cities has also
+emphasized the unintended casual flow of information facilitated by big
+cities.'' (\textbf{puga2010?})
 
-**Urban location of dentist; The Alonso model**
+\textbf{Urban location of dentist; The Alonso model}
 
-The Alsonso model demonstrate geographical locations tied up to location.
-Furthermore, It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
-In this case used to get an indicating of where the most profitable location for dental practise.
+The Alsonso model demonstrate geographical locations tied up to
+location. Furthermore, It is an urbanized formulation of the nineteenth
+century von Thünen model (\textbf{capello2015?}). In this case used to
+get an indicating of where the most profitable location for dental
+practise.
 
-Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
-Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain.
-In some cases different from the normal or average price competition @capello2015.
+Alonso assumes the existence of a city that cannot be build
+instantaneously, and therefore of an effective rent-curve from the city
+centre to the periphery. Furthermore, it determines the location for a
+new firm willing to locate in the city and the profit the firm can
+obtain. In some cases different from the normal or average price
+competition (\textbf{capello2015?}).
 
-As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
-More-so, the town or city has a single centre point.
-The city centre of business district (CBD) is generally defined as the most attractive for all firms.
-@capello2015.
-The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
-Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.
+As the Von Thünen assume a uniformed space where all land is equally
+fertile, this model envisages a city, endowed with infrastructure which
+cover the entire city in all directions whereas unit transport costs are
+constant in space. More-so, the town or city has a single centre point.
+The city centre of business district (CBD) is generally defined as the
+most attractive for all firms. (\textbf{capello2015?}). The model also
+assume perfect competition and unlimited demand, in other words it is
+supply-orientated (\textbf{capello2015?}). Furthermore, demonstrating a
+specific production function with fixed coefficients and constant
+returns to scale.
 
-Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
-The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
+Rent obtained in the model is the remainder left after transport,
+production costs and profit has been subtracted from revenues
+(\textbf{capello2015?}). The bid-rent curve will in this case
+demonstrate the prise of rent within Bank and Financial locations,
+Retail locations, and industrial locations.
 
-![](images/bidrent.gpg.jpg)
+\includegraphics{images/bidrent.gpg.jpg}
 
-The illustration shows three categories: The most expensive rent closest to the centre is the Bank and finance district, followed by the retail agglomeration areas and industrial sites furthest away from the central point.
-Dental practices are service orientated thus depending on population density and central activities.
-Opening hours are relay outside of office hours.
-Production cost does not appear to be important in this case since dentist is a public service.
-Hence, not dependent of regular transportation of sales goods.
-The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
+The illustration shows three categories: The most expensive rent closest
+to the centre is the Bank and finance district, followed by the retail
+agglomeration areas and industrial sites furthest away from the central
+point. Dental practices are service orientated thus depending on
+population density and central activities. Opening hours are relay
+outside of office hours. Production cost does not appear to be important
+in this case since dentist is a public service. Hence, not dependent of
+regular transportation of sales goods. The rent realized by the
+landowner is the envelope of the three rent supply curves
+(\textbf{capello2015?}).
 
-From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
-Furthermore, they may also increase their profits further out in to other retail agglomeration area.
-The attractiveness might be high in these places due to economies of scope.
-As the positive benefits from the Marshallian agglomeration attributes play out.
-Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013McCann.
+From a Marshallian way of thinking dentists should cluster somewhere in
+this pink area to reap maximum sales volume. Furthermore, they may also
+increase their profits further out in to other retail agglomeration
+area. The attractiveness might be high in these places due to economies
+of scope. As the positive benefits from the Marshallian agglomeration
+attributes play out. Subsequently, increasing profit by trading higher
+transportation cost for lower rent. In addition fewer dentists to
+compete with supporting the hotellings game theory whereas companies
+more to achieve spatial monopoly (\textbf{mccann2013McCann?}).
 
-**Gravitational models**
+\textbf{Gravitational models}
 
-The centre attracts firms, people and activities.
-Furthermore, it influence the central points in diverse ways such as commuter movements, diffusion of knowledge, corporations and network and personal relationship @capello2015.
-More-so, they are generated by the intensity of the flows generated between places @capello2015.
-The gravitational pull to a place of interest for a dentist will determine how much sales can be generated by locating there.
-Several locations can then be measures against each other to determine profit maximisation.
+The centre attracts firms, people and activities. Furthermore, it
+influence the central points in diverse ways such as commuter movements,
+diffusion of knowledge, corporations and network and personal
+relationship (\textbf{capello2015?}). More-so, they are generated by the
+intensity of the flows generated between places (\textbf{capello2015?}).
+The gravitational pull to a place of interest for a dentist will
+determine how much sales can be generated by locating there. Several
+locations can then be measures against each other to determine profit
+maximisation.
 
-These models has the potential to measures the future attractiveness of a place.
-Changes in infrastructure such as easier access to a place might change the attractiveness of a place or move the flow of people.
-Subserviently, increasing the gravitationally pull between the two places.
-Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
-Furthermore, these models have predictive capacity and can be used to estimate the potential impact.
-In particular, location of new productive activity in an area @capello2015.
-For instance predict where to build a shopping centre @capello2015.
-By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
+These models has the potential to measures the future attractiveness of
+a place. Changes in infrastructure such as easier access to a place
+might change the attractiveness of a place or move the flow of people.
+Subserviently, increasing the gravitationally pull between the two
+places. Furthermore, The model uses population and distance to measure
+the attractiveness and relations between two places
+(\textbf{capello2015?}). Furthermore, these models have predictive
+capacity and can be used to estimate the potential impact. In
+particular, location of new productive activity in an area
+(\textbf{capello2015?}). For instance predict where to build a shopping
+centre (\textbf{capello2015?}). By taking these factors in to account
+dentist could predict the most profitable locations in years to come by
+predicting the amount of people that will move from one city to another
+or tow different points of interest within a city.
 
-**Retail Agglomeration and Economies of Shopping Centres**
+\textbf{Retail Agglomeration and Economies of Shopping Centres}
 
-Town and cities have long had shopping districts in which stores are concentrated.
-To some extent such centre points were the result of uncoordinated store-location decisions @brueckner2011.
+Town and cities have long had shopping districts in which stores are
+concentrated. To some extent such centre points were the result of
+uncoordinated store-location decisions (\textbf{brueckner2011?}).
 Nevertheless, not coincidentally according to Hotelling's theory (1929).
-He describes the concentration of retail stores by a market desire to create a Nash equilibrium.
-At this point all retailers expected sales maximisation outcome cannot be improved by changing position.
-Contradictory to the assumption that consumers will shop at the nearest retailer, they are willing to travel significantly further to reach certain central places.
-Marshall (1920) points out another explanation for this in his seminal work on agglomeration.
-That being,intra-industry knowledge spillovers as mechanisms leading to formation of industry cluster @nielsen2021.
-In this case the dental industry information spillovers are bounded in geographical space.
-Hence, the need to locate in close proximity.
-That being the latest equipment, in order to work faster or the right personnel to attract more patients.
-Another way to drive sales in this area could be by dentist specialisation and patients referrals between dentists.
+He describes the concentration of retail stores by a market desire to
+create a Nash equilibrium. At this point all retailers expected sales
+maximisation outcome cannot be improved by changing position.
+Contradictory to the assumption that consumers will shop at the nearest
+retailer, they are willing to travel significantly further to reach
+certain central places. Marshall (1920) points out another explanation
+for this in his seminal work on agglomeration. That being,intra-industry
+knowledge spillovers as mechanisms leading to formation of industry
+cluster (\textbf{nielsen2021?}). In this case the dental industry
+information spillovers are bounded in geographical space. Hence, the
+need to locate in close proximity. That being the latest equipment, in
+order to work faster or the right personnel to attract more patients.
+Another way to drive sales in this area could be by dentist
+specialisation and patients referrals between dentists.
 
-Another important externalities expressed in Marshall (1920) theory is market transactions.
-Whereas, geographical proximity reduces transaction cost through several mechanism @nielsen2021.
-The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." @bakerBaker.
-Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
-Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
-In both cases growth potential would be compromised.
+Another important externalities expressed in Marshall (1920) theory is
+market transactions. Whereas, geographical proximity reduces transaction
+cost through several mechanism (\textbf{nielsen2021?}). The market
+transactions can be extended to the endogenous location of specialized
+human capital (\textbf{nielsen2021?}). ``It is well known among
+economists and other social scientists that large cities have
+disproportionately large shares of highly educated workers, and the
+trend has been growing in recent decades.'' (\textbf{bakerBaker?}).
+Presumably, larger cities would then be more attractive to dentists in
+terms of cost and quality of specialised labour in smaller towns.
+Especially, those trying to achieve spatial monopoly end up loosing
+profit by overpaying for labour or in worst case scenario not finding
+enough dentists labour and end up compromising on quality. In both cases
+growth potential would be compromised.
 
-Marshall also points out the competition within the geographical location.
-Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
-The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
+Marshall also points out the competition within the geographical
+location. Hence, the marketplace and business districts give consumers
+low search and switching costs (\textbf{capello2015?}). The fact that
+owners of stores seem to prefer highly concentrated shopping. More-so,
+it suggests that the sales volumes outweigh the loss attributed to
+greater price competition@capello2015.
 
-**Shopping Malls**
+\textbf{Shopping Malls}
 
-Owners of shopping malls "orchestrate" the process of retail agglomeration that happens naturally in towns and cities @brueckner2011.
-Naturally, the same retail agglomeration and Nash equilibrium may occur in shopping malls if not more intensely as the shops are in closer proximity.
-Furthermore, the potential for more knowledge spillovers, higher volume of market transaction and more competition.
-When retail agglomeration is orchestrated by the owner of a shopping mall, the strength and direction of such externalities are taken into account @brueckner2011.
-Much so, because the price of rent is dependent on the overall revenue of a shopping mall.
-Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
+Owners of shopping malls ``orchestrate'' the process of retail
+agglomeration that happens naturally in towns and cities
+(\textbf{brueckner2011?}). Naturally, the same retail agglomeration and
+Nash equilibrium may occur in shopping malls if not more intensely as
+the shops are in closer proximity. Furthermore, the potential for more
+knowledge spillovers, higher volume of market transaction and more
+competition. When retail agglomeration is orchestrated by the owner of a
+shopping mall, the strength and direction of such externalities are
+taken into account (\textbf{brueckner2011?}). Much so, because the price
+of rent is dependent on the overall revenue of a shopping mall. Hence,
+the owner of the mall wants to choose the mix of stores and their sizes
+taking these externalities, attributes and anchor stores attractiveness
+into account (\textbf{brueckner2011?}).
 
-The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
-Anchor-stores such as the Wine-monopoly bring in a higher volume of consumers @lai2013Lai.
-For example, shoppers visiting the Wine-monopoly in a mall may also visit a clothing store, and even find it beneficial to use public services at the same time.
+The Norwegian retail market for alcoholic beverages is controlled by the
+state monopoly. Anchor-stores such as the Wine-monopoly bring in a
+higher volume of consumers (\textbf{lai2013Lai?}). For example, shoppers
+visiting the Wine-monopoly in a mall may also visit a clothing store,
+and even find it beneficial to use public services at the same time.
 Nevertheless, they are limited in numbers and needs government approval.
-Each municipality must apply to get the wine monopoly in their area.
-One of the factors considered in this process is the death of city centre.
-A possible reason for this is so called "predatory malls" which are purposely built to overtake the exciting local market.
-Perhaps more so if malls cater for cinema, restaurants, bars and other attributes in conjunction with the city centre.
-Not to mention the upper hand of easy road access and free parking.
+Each municipality must apply to get the wine monopoly in their area. One
+of the factors considered in this process is the death of city centre. A
+possible reason for this is so called ``predatory malls'' which are
+purposely built to overtake the exciting local market. Perhaps more so
+if malls cater for cinema, restaurants, bars and other attributes in
+conjunction with the city centre. Not to mention the upper hand of easy
+road access and free parking.
 
-## Data
+\hypertarget{data}{%
+\subsection{Data}\label{data}}
 
-`In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
-If the data was collected by the authors this section will describe how the sample was collected.
-If specific instruments were constructed to collect the data they will be described here and included in the appendix.
-Procedures used to measure variables are often described in sufficient detail to permit replication.
-In economic papers, this section describes the empirical model and estimation strategy used by the authors.`
+\texttt{In\ quantitative\ studies\ this\ section\ provides\ details\ about\ dataset,\ variables,\ instruments\ and\ measurement\ procedures\ used\ in\ the\ study.\ If\ the\ data\ was\ collected\ by\ the\ authors\ this\ section\ will\ describe\ how\ the\ sample\ was\ collected.\ If\ specific\ instruments\ were\ constructed\ to\ collect\ the\ data\ they\ will\ be\ described\ here\ and\ included\ in\ the\ appendix.\ Procedures\ used\ to\ measure\ variables\ are\ often\ described\ in\ sufficient\ detail\ to\ permit\ replication.\ In\ economic\ papers,\ this\ section\ describes\ the\ empirical\ model\ and\ estimation\ strategy\ used\ by\ the\ authors.}
 
-**Data description**
+\textbf{Data description}
 
@@
 
-**Empirical Specification**
+\textbf{Empirical Specification}
 
@@
 
-**Estimation strategy**
+\textbf{Estimation strategy}
 
-**Analytical framework**
+\textbf{Analytical framework}
 
@@
 
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
-The Journal of the statistics society define such data as being objects in space, which objects have physical location.
-Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
+Spatial data analysis is a rapidly growing area in Statistics
+(\textbf{haslett1992?}). The Journal of the statistics society define
+such data as being objects in space, which objects have physical
+location. Furthermore, the analyses are spatial if these locations are
+relevant to the interpretation of the data (\textbf{haining1998?}).
 
-The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
-Graphical techniques for examining such maps are thus a central part of the methodology.
-A number of such techniques are discussed."` @haslett1992.
+The examination of the data involves the examination of data collected
+from q-gis maps;
+\texttt{it\ will\ involve\ little\ more.\ Graphical\ techniques\ for\ examining\ such\ maps\ are\ thus\ a\ central\ part\ of\ the\ methodology.\ A\ number\ of\ such\ techniques\ are\ discussed."}
+(\textbf{haslett1992?}).
 
-`What data are used?`
+\texttt{What\ data\ are\ used?}
 
-`What has been calculated?`
+\texttt{What\ has\ been\ calculated?}
 
-## **Econometric Approach**
+\hypertarget{econometric-approach}{%
+\subsection{\texorpdfstring{\textbf{Econometric
+Approach}}{Econometric Approach}}\label{econometric-approach}}
 
 The questions addressed are related to the geographical space of Norway.
-Norway is a country with relatively small population and large heterogeneous land area.
-Furthermore, the population is decentralized with the majority of population living in urban areas.
-This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
-Whereas one of the assumptions are homogeneous landscape.
-Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
-Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
+Norway is a country with relatively small population and large
+heterogeneous land area. Furthermore, the population is decentralized
+with the majority of population living in urban areas. This unique
+spatial distribution of land might give a possible explanation to
+diversions from the simple theoretical models used in this paper.
+Whereas one of the assumptions are homogeneous landscape. Nevertheless,
+the way economies are organised comes partly from the emergence of
+centralities within a city. Transportation networks have the potential
+to increase sprawl depending on how the urban system evolves gradually.
 
-The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
-Having an urban system of multiple centres and applying the same logic of the model to other retail agglomeration areas.
-Resulting in finding relationships based on three different types of distance measurements.
-Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
-Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
-Additionally, dentist also located close to each other making some sort of industrial cluster.
-Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
+The results from the bid rent configuration from the city centre gave
+the opportunity to explore other locations in each municipality's.
+Having an urban system of multiple centres and applying the same logic
+of the model to other retail agglomeration areas. Resulting in finding
+relationships based on three different types of distance measurements.
+Firstly, the distance between Dental Practice and CBD then, dentist and
+other retail agglomeration areas and lastly amongst dentists. Finding
+that most dentist are either located close to a City, Central Business
+District or a shopping mall. Additionally, dentist also located close to
+each other making some sort of industrial cluster. Supporting,
+Marshall's agglomeration theory on knowledge spillovers, increasing the
+sills to attract more patients.
 
-Furthermore, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
-Furthermore, decreasing as the distance gets bigger.
-Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
-Hence supporting findings; when the distance reach around 10km from the mall the income starts to increase before it decrease again around 20km.
-Also in line with the Keynesian theory suggesting that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
-Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
+Furthermore, when using the dentist income (y) and comparing the
+distance from the mall (x), signs of patterns emerge showing that income
+is greater when the dentist is placed closer to the shopping mall.
+Furthermore, decreasing as the distance gets bigger. Faster
+transportation can increase convenience of commuting, increasing rent
+commuters from the city centre are willing to pay, and therefore
+increasing the area of developed land without this being necessary
+corresponding to an increase in population. Hence supporting findings;
+when the distance reach around 10km from the mall the income starts to
+increase before it decrease again around 20km. Also in line with the
+Keynesian theory suggesting that this pattern might take place as the
+transportation cost for individuals further away from the Shopping mall
+does not increase to much by travelling the extra distance. Hotelling
+also describes this case where the consumer are willing to travel
+further to reach a certain central place.
 
-According to the Alons's model dentist should locate in the retail area for higher volume of sales.
-Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.
-However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
-Our findings in QGIS shows that in areas of high population density, the number of dentist also increase.
-The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
-Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation cost to the second closes dentist are noticeable higher.
+According to the Alons's model dentist should locate in the retail area
+for higher volume of sales. Dentist being placed further out in to other
+retail agglomeration areas can also be a good strategy. However, the
+dentist would then rely on the area having less competition so the rate
+between competition and population density are proportional. Our
+findings in QGIS shows that in areas of high population density, the
+number of dentist also increase. The consumer/patient then has more to
+choose between and the dentist also has a bigger scale of potential
+customers. Where in smaller areas it is less competition and the dentist
+would appear to have somewhat of a monopoly since the transportation
+cost to the second closes dentist are noticeable higher.
 
-Furthermore, being aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
-This also takes away some of the competition and would support placement in more rural areas.
-More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
-Suggesting some dentist might seek closer to CBD areas to capture most of the remaining part of the population.
-Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
-Reducing the transportation cost for the customer as they are already located in these areas even if they do not live here.
+Furthermore, being aware of the fact that going to the dentist is a
+public services and most people get assigned to their closest dentist
+based on their living address. This also takes away some of the
+competition and would support placement in more rural areas. More-so, is
+important to notice that when passed 18 years of age, you could risk
+loosing your local public dentist due to lack of capacity. Suggesting
+some dentist might seek closer to CBD areas to capture most of the
+remaining part of the population. Business, public places and industry
+can have a positive effect on revenue since this is areas where most of
+the population spends their time during their day, either going to work,
+school and other errand. Reducing the transportation cost for the
+customer as they are already located in these areas even if they do not
+live here.
 
-The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
-Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
-The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
-Bertrand model also take price as competitive advantage in to consideration.
+The bid rent models are under the assumption that Dentists use spatial
+completion to maximise profit. Nevertheless, the cost of dental care in
+Norway can vary depending on a number of other factors such as the type
+of treatment, the location of the dental practice and the dentist
+experience and qualifications. The Norwegian Dental Association rate the
+average cost of check ups is between four to eight hundred Norwegian
+Krone's {``Tannlegeforeningen''} (n.d.). Bertrand model also take price
+as competitive advantage in to consideration.
 
-Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
-Tjener tannleger i større byer mer?
+Viser data noe om indikasjon på dyrere priser i nærheten av større byer
+of hvor det er stedsbestemt monopol? Tjener tannleger i større byer mer?
 Her kan vi også skrive om gravitasjons modeller.
 
-Another Mashiallian point to be made could be the advantage of industry clusters.
-Taken for granted that the data on Dentist includes both "normal Dentists" and specialist.
-Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
-The customer would then probably prefer to go to the closes specialist.
-The access to work force in these areas would also be good and will be within the same geographic area.
+Another Mashiallian point to be made could be the advantage of industry
+clusters. Taken for granted that the data on Dentist includes both
+``normal Dentists'' and specialist. Where in this case a ``normal
+dentist'' would for example recommend one of their customers to go see a
+specialist to get braces. The customer would then probably prefer to go
+to the closes specialist. The access to work force in these areas would
+also be good and will be within the same geographic area.
 
-Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
-Prior to the most retail activities took place in the city centre.
-Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
-In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway @home-n.
+Shopping malls have a relatively short history in Norway, with the first
+shopping mall opening in the late sixties. Prior to the most retail
+activities took place in the city centre. Furthermore, these malls
+typically featured a mix of shops, restaurants and entertainment
+options, and they were often located in suburban areas or on the
+outskirts of cities {``Home - NCSC''} (n.d.). In resent years, there has
+been a trend towards larger and more comprehensive shopping malls in
+Norway {``Home - NCSC''} (n.d.).
 
-However, the growth of shopping malls in Norway has not been without controversy.
-Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments @kanvin2014.
+However, the growth of shopping malls in Norway has not been without
+controversy. Some critics argue that the development of large shopping
+malls has led to the decline of traditional shopping streets and local
+businesses, while other raise concerns about the environmental impact of
+these large scale commercial developments {``Kan Vin Og Sprit Skape
+Byutvikling?''} (2014).
 
-Information about Wine monopoly locations has also been studied in this paper.
-As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
-The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
-More importantly, the government may take in to account the potential impact of exciting retail centres in the region.
-This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
+Information about Wine monopoly locations has also been studied in this
+paper. As the wine monopoly are anchor stores they may affect the
+attractiveness of the shopping mall {``Kan Vin Og Sprit Skape
+Byutvikling?''} (2014). The Norwegian government regulates the
+placements of wine monopoly stores in shopping malls to prevent
+completion with town centres omsorgsdepartementet (2022). More
+importantly, the government may take in to account the potential impact
+of exciting retail centres in the region. This regional economic
+approach aim to promote sustainable development and balance economic
+growth across the country {``Kan Vin Og Sprit Skape Byutvikling?''}
+(2014).
 
@@
 
-## Conclusion
+\hypertarget{conclusion}{%
+\subsection{Conclusion}\label{conclusion}}
 
-## References
+\hypertarget{references}{%
+\subsection*{References}\label{references}}
+\addcontentsline{toc}{subsection}{References}
+
+\hypertarget{refs}{}
+\begin{CSLReferences}{1}{0}
+\leavevmode\vadjust pre{\hypertarget{ref-home-n}{}}%
+{``Home - NCSC.''} n.d.
+\href{https://ncscnordic.org/,\%20https://ncscnordic.org/}{https://ncscnordic.org/,
+https://ncscnordic.org/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-kanvin2014}{}}%
+{``Kan Vin Og Sprit Skape Byutvikling?''} 2014.
+\url{https://www.aftenposten.no/kultur/i/oEy0/kan-vin-og-sprit-skape-byutvikling}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-omsorgsdepartementet2022}{}}%
+omsorgsdepartementet, Helse-og. 2022. {``Vinmonopolordningen.''}
+\emph{Regjeringen.no}. regjeringen.no.
+\url{https://www.regjeringen.no/no/tema/helse-og-omsorg/psykisk-helse/innsikt/vinmonopolordningen/id2528183/}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-tannlege}{}}%
+{``Tannlegeforeningen.''} n.d. \url{https://www.tannlegeforeningen.no/}.
+
+\end{CSLReferences}
+
+\end{document}

### 2023-02-19 Retail Agglomeration_Theory.qmd
Ferdig med abstract håper at resultatet stemmer lol

--- 2023-02-19 Retail Agglomeration_Theory.qmd
@@
 
-`The topic of research (What was studied)`
+There are several factors to take in to consideration when deciding where a dental practice should be located.
+Spatial competition is a critical consideration in regional economics.
+Some of the factors influencing the decision making is retail agglomeration attributes, spatial monopoly and the attractiveness of a particular place of interest.
+This study aims to examine the dental industry and find the strategy most profitable for the dental industry.
 
-`The purpose (What was done)`
+The study utilize a quantitative research approach, employing spacial explanatory data analysis techniques to determine the location decisions of firms.
+Geographical data was collected from locations in Norway showing all dentists, central places, retail agglomeration and other points of interest such as the government regulated wine monopoly.
+Spatial regression analyses was used to identify the factors influencing the location decisions of firms.
 
-`The methodology (how the research was done)`
+Furthermore, the result indicates the distance to retail agglomeration areas and central places playes a significant role in the location decisions for firms.
+Firms are more likely to locate closer to these areas, which offer access to a larger customer base and lower cost of transportation.
+However, the results also suggest other factors, such as market size and completion also influence the location decisions for firms.
 
-`Main findings (what was found)`
-
-`Policy and implications (significance of findings)`
+The studies findings have particular implications for policy-makers, developer, and dental practises seeking to maximize profits through strategic decisions.
+By understanding the importance of proximity to retail agglomeration areas and central places, these stakeholders can make informed decisions that benefit the local economy and improve competitiveness of the region.
+Perhaps more importantly prevent agrassive retail agglomeration areas from "killing down town".
 
@@
 
-`A brief description of the problem and its importance`
+`A brief description of the problem and its importance CHECK`
 

### 2023-02-19 Retail Agglomeration_Theory.qmd
...

@@
 Geographical data was collected from locations in Norway showing all dentists, central places, retail agglomeration and other points of interest such as the government regulated wine monopoly.
-Spatial regression analyses was used to identify the factors influencing the location decisions of firms.
+Combines with regional income data a spatial regression analyses was used to identify the factors influencing the location decisions of firms.
 
@@
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
-
-`A brief description of the problem and its importance CHECK`
-
-`Statement of purpose or what was done`
-
-`A brief description of main results`
-
-`Description of methodology, main challenges and solutions`
-
-`Discretion of the papers contribution in relation to previous research`
-
-`Organization of paper`
 

### 2023-02-19 MSB106_Assignment.qmd
Commit.

@@
 Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
-NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
+NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv") |>
+  rename("HubName" = fid)
 Zonal_Den <- read.csv("Zonal_Den.csv") |>
@@
   select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
-  rename("knr" = kommunenummer, "DistMal" = HubDist, "fid" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
+  rename("knr" = kommunenummer, "DistMal" = HubDist, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
 ```
@@
 ```{r}
-Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
-  select(knr, DistMal, fid, Turnover_capita_retail, shop, Vinmonopolet) |>
+Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
+  select(knr, DistMal, HubName, Turnover_capita_retail, shop, Vinmonopolet) |>
   rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
@@
 Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
+Commune_Data$Winemonopoly <- as.integer(as.logical(Commune_Data$Winemonopoly))
 ```
@@
   select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
-  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubName" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
 ```
@@
 ```{r}
-Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
-  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
+Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
+  select(fid, Juridisk.n, Antall.ans, Driftsres, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
   rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
@@
 Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
+Dentist_Data$Winemonopoly <- as.integer(as.logical(Dentist_Data$Winemonopoly))
 ```
-
-## Introduction
-
-Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
-More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
-As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
-
-## Theoretical Foundation of Hypothesis
-
-## Data description
 
@@
 Commune_Data |>
-  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
+  ggplot(aes(x = SizeMall, y = Turnover_capita_retail, colour = SizeMall)) +
   geom_point(lwd = .9) +
@@
 ```
-
-Might be more appropriate to explore the distance from CBDs to Malls compared to turnover capita in retail per inhabitant, rather than the distance from the geographical centroid of Norwegian Communes, because of the tendency of bigger population density around CBDs.
 
@@
   filter(Sum.salgsi < 100000000, DistCBD < 250,
-         Sum.Drifts < 20000000, Sum.Drifts > -10000000, DistMal < 80)
+         Driftsres < 20000000, Driftsres > -5000000, DistMal < 80)
 ```
@@
 Dentist_Data |>
-  ggplot(aes(x = DistMal, y = Sum.Drifts, colour = DistMal)) +
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
   geom_point(lwd = .9) +
@@
 Dentist_Data_LIMIT |>
-  ggplot(aes(x = DistMal, y = Sum.Drifts, colour = DistMal)) +
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
   geom_point(lwd = .9) +
@@
 ```{r}
-huxreg(list("Dentist Sales" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+```
+
+```{r}
+Den_W_Wine <- Dentist_Data_LIMIT |>
+  filter(Winemonopoly == 1)
+```
+
+```{r}
+Den_WO_Wine <- Dentist_Data_LIMIT |>
+  filter(Winemonopoly == 0)
+```
+
+```{r}
+lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
+```
+
+```{r}
+huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
 ```

### 2023-02-19 Retail Agglomeration.qmd
Commit.

--- 2023-02-18 Retail Agglomeration.qmd
@@
 For this assignment we where provided with different geospatial data in the geographic information software QGIS to calculate the relationship of the different variables that might determine firm activity.
-This geospatial data included Norwegian commune data which contained the administrative boundaries of the Communes, presented in the software as different divisions of lines that distinguish the different commune boundaties.
+This geospatial data included Norwegian commune data which contained the administrative boundaries of the Communes, presented in the software as different divisions of lines that distinguish the different commune boundaries.
 We where also provided with different point layers, such as the geometric centers of communes, urban city centers or Central Business Districts (CBD), all businesses in Norway which are regarded as dental services, and all Norwegian shopping malls.
@@
 
-To carry out a full-flegded data analysis of the influencing factors of the activity of Norwegian dentist services, we collected some more data than what we already had been presented with.
+To carry out a more complete data analysis of the influencing factors of the activity of Norwegian dentist services, we had to collect more data than what we already had been presented with.
 We firstly utilized QGIS to calculate different other statistics and variables.
-Regarding the Norwegian shopping malls layer we added the variable \"Wine-monopoly\" as a false/ true variable to determine which Norwegian malls, and also the number of stores which are in each mall.
+Regarding the Norwegian shopping malls layer we added the variable "Wine-monopoly" as a false/ true variable to determine which Norwegian malls had a hard liquor store.
+We added the number of stores which are in each mall, to use as an estimate for the malls sizes.
 The malls opening hours have also been included in this layer.
 From SSB, we gathered data on the average municipal sales per inhabitant in retailing for each Norwegian commune.
-We also created 2 km buffer zones around dentists and CBDs to determine population density in these areas.
+Lastly we created 2 kilometer buffer zones around dentists and CBDs to determine population density in these areas.
+
+After calculations in QGIS we transferred our data to RSTUDIO to do the primary statistical analysis of our data.
+The data is then merged into two different data frames.
+The first one containing Norwegian communes as the unit of observation, which includes geometric commune centers to the nearest mall, the retail per capita per commune, the different commune numbers and the size of the nearest mall, including the true/ false "Winemonopoly" variable.
+The second data frame use dentists as the unit of observations.
+This data frame include data on the distances from CBD to dentists, distances from dentists to the nearest mall, including variables of the mall size and the "Winemonopoly" variable, different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
+
+To obtain a first impression and a simple understanding of our data, we have decided to make use of scatterplots.
+The scatterplot presented in @fig-1 show the relationship of average spending on retail amongst inhabitants and the distance from geometrical commune centers to nearest malls, which indicate that most malls are in a 50 km radius of commune centers, and that Norwegians on average spend between 50 thousand to 120 thousand on retail.
+
+We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
+In the first scatterplot we see some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these figures are accounting figures for an entire group of companies under the same legal name.
+By ignoring these outliers we are then left with the scatterplot in @fig-3.
+This scatterplot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
+The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall.
+Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
+A decent number of dentists run at a loss, having negative earnings before interest and taxes.
 

### 2023-02-19 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
Add files via upload

--- 2023-02-19 Retail Agglomeration_Theory.qmd
@@
 Geographical data was collected from locations in Norway showing all dentists, central places, retail agglomeration and other points of interest such as the government regulated wine monopoly.
-Combines with regional income data a spatial regression analyses was used to identify the factors influencing the location decisions of firms.
+Combined with regional income data a spatial regression analyses was used to identify the factors influencing the location decisions of firms.
 
@@
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
+
+The theoretical framework for this study draws on Marshall's agglomeration principles, which emphasize spatial aspects of economic growth.
+Hotelling's game theory is used to explain spatial competition by location.
+Furthermore, adding the hybrid model whereas firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products.
+Alsonso's bid-rent model and gravitational models also provides insights into how dentists or other retailer makes location decisions in the context of regional growth.
+
+The dental industry provides an interesting industry to study for this research because of its unique characteristics.
+Dental practices require a high degree of specialization and expertise.
+At the same time, they need to be easily accessible to consumers.
+Shopping malls and central places offer a continent location for dental practices that said competition for prime locations can be intense and push some in to spatial monopoly.
+
+The result of this study indicates that dentist's with the highest income are usually located in central places in big cities.
+The findings are in line with already exciting theories especially supporting the gravitational models and agglomeration attributes.
+?????? Proving a central location within large cities are the best location to maximize profit for Dental practices.
+???
 

### 2023-02-19 MSB106_Assignment.qmd
...

--- 2023-02-19 MSB106_Assignment.qmd
@@
   select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
-  rename("knr" = kommunenummer, "DistMal" = HubDist, "fid" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
-```
-
-```{r}
-Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
-  select(knr, DistMal, fid, Turnover_capita_retail, shop, Vinmonopolet) |>
-  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
-```
-
-```{r}
-Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
+  rename("knr" = kommunenummer, "DistMal" = HubDist, "HubNameMal" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
 ```
@@
 
-```{r}
-Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "fid") |>
-  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
-  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
-```
+**Abstract**
 
-```{r}
-Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
-```
+There are several factors to take in to consideration when deciding where a dental practice should be located.
+Spatial competition is a critical consideration in regional economics.
+Some of the factors influencing the decision making is retail agglomeration attributes, spatial monopoly and the attractiveness of a particular place of interest.
+This study aims to examine the dental industry and find the strategy most profitable for the dental industry.
+
+The study utilize a quantitative research approach, employing spacial explanatory data analysis techniques to determine the location decisions of firms.
+Geographical data was collected from locations in Norway showing all dentists, central places, retail agglomeration and other points of interest such as the government regulated wine monopoly.
+Combined with regional income data a spatial regression analyses was used to identify the factors influencing the location decisions of firms.
+
+Furthermore, the result indicates the distance to retail agglomeration areas and central places playes a significant role in the location decisions for firms.
+Firms are more likely to locate closer to these areas, which offer access to a larger customer base and lower cost of transportation.
+However, the results also suggest other factors, such as market size and completion also influence the location decisions for firms.
+
+The studies findings have particular implications for policy-makers, developer, and dental practises seeking to maximize profits through strategic decisions.
+By understanding the importance of proximity to retail agglomeration areas and central places, these stakeholders can make informed decisions that benefit the local economy and improve competitiveness of the region.
+Perhaps more importantly prevent agrassive retail agglomeration areas from "killing down town".
 
@@
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
+
+The theoretical framework for this study draws on Marshall's agglomeration principles, which emphasize spatial aspects of economic growth.
+Hotelling's game theory is used to explain spatial competition by location.
+Furthermore, adding the hybrid model whereas firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products.
+Alsonso's bid-rent model and gravitational models also provides insights into how dentists or other retailer makes location decisions in the context of regional growth.
+
+The dental industry provides an interesting industry to study for this research because of its unique characteristics.
+Dental practices require a high degree of specialization and expertise.
+At the same time, they need to be easily accessible to consumers.
+Shopping malls and central places offer a continent location for dental practices that said competition for prime locations can be intense and push some in to spatial monopoly.
+
+The result of this study indicates that dentist's with the highest income are usually located in central places in big cities.
+The findings are in line with already exciting theories especially supporting the gravitational models and agglomeration attributes.
+Proving a central location within large cities are the best location to maximize profit for Dental practices.
 
@@
 ```{r}
-kable(summary(Dentist_Data[7:10]))
+kable(summary(Dentist_Data[8:11]))
 ```
@@
 ```{r}
-kable(summary(Dentist_Data[11:12]))
+kable(summary(Dentist_Data[11:14]))
 ```
@@
   filter(Sum.salgsi < 100000000, DistCBD < 250,
-         Sum.Drifts < 20000000, Sum.Drifts > -10000000, DistMal < 80)
+         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
 ```
@@
 ```{r}
-lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
+lm1 <- lm(Sum.salgsi ~ X_sum, data = Dentist_Data_LIMIT)
+lm2 <- lm(Sum.salgsi ~ DistComp + DistCBD + DistMal + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
 ```
@@
 ```{r}
-huxreg(list("Dentist Sales" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+huxreg(list("Model 1" = lm1, "Model 2" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
 ```

### 2023-02-19 Retail-Agglomeration_abstract_introduction_theori_econometric_approach.tex
...

--- 2023-02-19 Retail-Agglomeration_Theory.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, frame hidden, interior hidden, boxrule=0pt, enhanced, breakable]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
 
@@
 
-\texttt{The\ topic\ of\ research\ (What\ was\ studied)}
+There are several factors to take in to consideration when deciding
+where a dental practice should be located. Spatial competition is a
+critical consideration in regional economics. Some of the factors
+influencing the decision making is retail agglomeration attributes,
+spatial monopoly and the attractiveness of a particular place of
+interest. This study aims to examine the dental industry and find the
+strategy most profitable for the dental industry.
 
-\texttt{The\ purpose\ (What\ was\ done)}
+The study utilize a quantitative research approach, employing spacial
+explanatory data analysis techniques to determine the location decisions
+of firms. Geographical data was collected from locations in Norway
+showing all dentists, central places, retail agglomeration and other
+points of interest such as the government regulated wine monopoly.
+Combined with regional income data a spatial regression analyses was
+used to identify the factors influencing the location decisions of
+firms.
 
-\texttt{The\ methodology\ (how\ the\ research\ was\ done)}
+Furthermore, the result indicates the distance to retail agglomeration
+areas and central places playes a significant role in the location
+decisions for firms. Firms are more likely to locate closer to these
+areas, which offer access to a larger customer base and lower cost of
+transportation. However, the results also suggest other factors, such as
+market size and completion also influence the location decisions for
+firms.
 
-\texttt{Main\ findings\ (what\ was\ found)}
-
-\texttt{Policy\ and\ implications\ (significance\ of\ findings)}
+The studies findings have particular implications for policy-makers,
+developer, and dental practises seeking to maximize profits through
+strategic decisions. By understanding the importance of proximity to
+retail agglomeration areas and central places, these stakeholders can
+make informed decisions that benefit the local economy and improve
+competitiveness of the region. Perhaps more importantly prevent
+agrassive retail agglomeration areas from ``killing down town''.
 
@@
 
-\texttt{A\ brief\ description\ of\ the\ problem\ and\ its\ importance}
+The theoretical framework for this study draws on Marshall's
+agglomeration principles, which emphasize spatial aspects of economic
+growth. Hotelling's game theory is used to explain spatial competition
+by location. Furthermore, adding the hybrid model whereas firms compete
+both in location and price, with consumers having a transportation cost
+and firms producing differentiated products. Alsonso's bid-rent model
+and gravitational models also provides insights into how dentists or
+other retailer makes location decisions in the context of regional
+growth.
 
-\texttt{Statement\ of\ purpose\ or\ what\ was\ done}
+The dental industry provides an interesting industry to study for this
+research because of its unique characteristics. Dental practices require
+a high degree of specialization and expertise. At the same time, they
+need to be easily accessible to consumers. Shopping malls and central
+places offer a continent location for dental practices that said
+competition for prime locations can be intense and push some in to
+spatial monopoly.
 
-\texttt{A\ brief\ description\ of\ main\ results}
-
-\texttt{Description\ of\ methodology,\ main\ challenges\ and\ solutions}
-
-\texttt{Discretion\ of\ the\ papers\ contribution\ in\ relation\ to\ previous\ research}
-
-\texttt{Organization\ of\ paper}
+The result of this study indicates that dentist's with the highest
+income are usually located in central places in big cities. The findings
+are in line with already exciting theories especially supporting the
+gravitational models and agglomeration attributes. ?????? Proving a
+central location within large cities are the best location to maximize
+profit for Dental practices. ???
 
@@
 Location theory gives regional economics its scientific disciplinary
-identity and constitutes its theoretical methodological core
-(\textbf{capello2011?}). It has typically microeconomic foundation and
-uses theoretical models as well as adopting a statistical and
-geographical approach (\textbf{capello2011?}). Furthermore, the theory
-uses the concept of externalities in the spatial distribution of
-activities, thereby laying the territorial bases for dynamic approach to
-economic growth (\textbf{capello2011?}).
+identity and constitutes its theoretical methodological core Capello
+(2011). It has typically microeconomic foundation and uses theoretical
+models as well as adopting a statistical and geographical approach
+Capello (2011). Furthermore, the theory uses the concept of
+externalities in the spatial distribution of activities, thereby laying
+the territorial bases for dynamic approach to economic growth Capello
+(2011).
 
 Regional growth theory involves spatial aspects of economic growth and
-territorial distribution of income (\textbf{capello2011?}). It also
-involves generating geographical advantages, in terms of easy or
-difficult access to a particular area (\textbf{capello2011?}).
+territorial distribution of income Capello (2011). It also involves
+generating geographical advantages, in terms of easy or difficult access
+to a particular area Capello (2011).
 
@@
 of firms increase profit by attracting a larger number of potential
-customers, and more productive workers (\textbf{capello2011?}).
-Furthermore, the attractiveness of a central location increase the cost
-of rent.
+customers, and more productive workers Capello (2011). Furthermore, the
+attractiveness of a central location increase the cost of rent.
 
@@
 centre. Furthermore, it depends on the variety of goods which can be
-purchased at the same location (\textbf{mccann2013?}).
+purchased at the same location McCann (2013).
 
@@
 places of growth potential for dentist practice or shopping malls
-(\textbf{capello2015?}).
+Capello (2015).
 
@@
 single firm in a duopoly environment. Furthermore, no relocation costs
-and demand only depend on location choice (\textbf{capello2015?}). The
-location game starts off with the total market of AB, firm A in the
-middle of location A and firm B starts in the middle of location B. One
-firm starts relocating closer to the other to take some of the customers
-in the other market area (\textbf{capello2015?}). The other respond by
-doing the same and the game continues until both end up in the middle of
-the market on the broader of AB. The end of the game is the position
-where neither can increase sales volume by moving position
-(\textbf{capello2015?}).
+and demand only depend on location choice Capello (2015). The location
+game starts off with the total market of AB, firm A in the middle of
+location A and firm B starts in the middle of location B. One firm
+starts relocating closer to the other to take some of the customers in
+the other market area Capello (2015). The other respond by doing the
+same and the game continues until both end up in the middle of the
+market on the broader of AB. The end of the game is the position where
+neither can increase sales volume by moving position Capello (2015).
 
@@
 competition for specialized production factors such as labour and
-product market competition (\textbf{nielsen2021?}).
+product market competition Nielsen et al. (2021).
 
 There are solidly established conclusions regarding the existence of
-agglomeration economies (\textbf{puga2010?}). However, less proof of
-their estimated magnitude. Hence, identifying the causes of
-agglomeration economies, is proving more difficult (\textbf{puga2010?}).
-Nevertheless, there is a large theoretical literature that develops
-these mechanisms (\textbf{puga2010?}). (\textbf{duranton2004?}) discuss
-these classifications and identify \emph{learning, sharing} and
-\emph{matching} as the main causes of agglomeration economies.
+agglomeration economies Puga (2010). However, less proof of their
+estimated magnitude. Hence, identifying the causes of agglomeration
+economies, is proving more difficult Puga (2010). Nevertheless, there is
+a large theoretical literature that develops these mechanisms Puga
+(2010). Duranton and Puga (2004) discuss these classifications and
+identify \emph{learning, sharing} and \emph{matching} as the main causes
+of agglomeration economies.
 
@@
 infrastructure and facilities. Therefore, a variety of intermediate
-input suppliers, or a pool of workers with similar skills
-(\textbf{puga2010?}). Despite higher rent the dental industry and
-shopping malls, seams to reap higher benefits in more populated areas as
-they are dependent on being located where there is a higher volume of
-patients in order to drive sales. The attraction for the consumers and
-users of public facilities is overall cost reduction
-(\textbf{puga2010?}). Hence, the larger the population sharing
-facilities the lower the cost per user (\textbf{puga2010?}). Presumably,
-industrial factories and business clusters are more dependent on being
-close to raw materials and industrial action.
+input suppliers, or a pool of workers with similar skills Puga (2010).
+Despite higher rent the dental industry and shopping malls, seams to
+reap higher benefits in more populated areas as they are dependent on
+being located where there is a higher volume of patients in order to
+drive sales. The attraction for the consumers and users of public
+facilities is overall cost reduction Puga (2010). Hence, the larger the
+population sharing facilities the lower the cost per user Puga (2010).
+Presumably, industrial factories and business clusters are more
+dependent on being close to raw materials and industrial action.
 
@@
 employers and employees, improved chances of finding suitable and better
-quality of matches (\textbf{puga2010?}). Shopping malls require skilled
-workers to drive sales. However, they are not so dependent on highly
-educated workers as dentists whom according to resent study, tend to
-prefer bigger cities (as they are highly educated)
-(\textbf{davis2020?}). More-so, cites provide a constant market for
-specialized skills and more productive work force (\textbf{puga2010P?})
-Perhaps, a possibility for higher wages being compensated by more
-efficient workers in bigger agglomeration economies. That said the
-Norwegian cities might differ from the american cites in the study, as
-the population and clusters are no where near the same size. Lastly, a
-larger market can also facilitate learning by promoting the development
-and widespread adaptation of new technologies and business practice
-(\textbf{puga2010?}).
+quality of matches Puga (2010). Shopping malls require skilled workers
+to drive sales. However, they are not so dependent on highly educated
+workers as dentists whom according to resent study, tend to prefer
+bigger cities (as they are highly educated) Davis and Dingel (2020).
+More-so, cites provide a constant market for specialized skills and more
+productive work force Puga (2010). Perhaps, a possibility for higher
+wages being compensated by more efficient workers in bigger
+agglomeration economies. That said the Norwegian cities might differ
+from the american cites in the study, as the population and clusters are
+no where near the same size. Lastly, a larger market can also facilitate
+learning by promoting the development and widespread adaptation of new
+technologies and business practice Puga (2010).
 
@@
 Experienced workers remain in cities to share the rent of this learning
-process (\textbf{puga2010?}). Besides this purposeful transition of
-knowledge, the information literature on learning in cities has also
-emphasized the unintended casual flow of information facilitated by big
-cities.'' (\textbf{puga2010?})
+process Puga (2010). Besides this purposeful transition of knowledge,
+the information literature on learning in cities has also emphasized the
+unintended casual flow of information facilitated by big cities.'' Puga
+(2010)
 
@@
 location. Furthermore, It is an urbanized formulation of the nineteenth
-century von Thünen model (\textbf{capello2015?}). In this case used to
-get an indicating of where the most profitable location for dental
-practise.
+century von Thünen model Capello (2015). In this case used to get an
+indicating of where the most profitable location for dental practise.
 
@@
 obtain. In some cases different from the normal or average price
-competition (\textbf{capello2015?}).
+competition Capello (2015).
 
@@
 The city centre of business district (CBD) is generally defined as the
-most attractive for all firms. (\textbf{capello2015?}). The model also
-assume perfect competition and unlimited demand, in other words it is
-supply-orientated (\textbf{capello2015?}). Furthermore, demonstrating a
-specific production function with fixed coefficients and constant
-returns to scale.
+most attractive for all firms. Capello (2015). The model also assume
+perfect competition and unlimited demand, in other words it is
+supply-orientated Capello (2015). Furthermore, demonstrating a specific
+production function with fixed coefficients and constant returns to
+scale.
 
 Rent obtained in the model is the remainder left after transport,
-production costs and profit has been subtracted from revenues
-(\textbf{capello2015?}). The bid-rent curve will in this case
-demonstrate the prise of rent within Bank and Financial locations,
-Retail locations, and industrial locations.
+production costs and profit has been subtracted from revenues Capello
+(2015). The bid-rent curve will in this case demonstrate the prise of
+rent within Bank and Financial locations, Retail locations, and
+industrial locations.
 
@@
 regular transportation of sales goods. The rent realized by the
-landowner is the envelope of the three rent supply curves
-(\textbf{capello2015?}).
+landowner is the envelope of the three rent supply curves Capello
+(2015).
 
@@
 compete with supporting the hotellings game theory whereas companies
-more to achieve spatial monopoly (\textbf{mccann2013McCann?}).
+more to achieve spatial monopoly McCann (2013).
 
@@
 diffusion of knowledge, corporations and network and personal
-relationship (\textbf{capello2015?}). More-so, they are generated by the
-intensity of the flows generated between places (\textbf{capello2015?}).
-The gravitational pull to a place of interest for a dentist will
-determine how much sales can be generated by locating there. Several
-locations can then be measures against each other to determine profit
-maximisation.
+relationship Capello (2015). More-so, they are generated by the
+intensity of the flows generated between places Capello (2015). The
+gravitational pull to a place of interest for a dentist will determine
+how much sales can be generated by locating there. Several locations can
+then be measures against each other to determine profit maximisation.
 
@@
 places. Furthermore, The model uses population and distance to measure
-the attractiveness and relations between two places
-(\textbf{capello2015?}). Furthermore, these models have predictive
-capacity and can be used to estimate the potential impact. In
-particular, location of new productive activity in an area
-(\textbf{capello2015?}). For instance predict where to build a shopping
-centre (\textbf{capello2015?}). By taking these factors in to account
+the attractiveness and relations between two places Capello (2015).
+Furthermore, these models have predictive capacity and can be used to
+estimate the potential impact. In particular, location of new productive
+activity in an area Capello (2015). For instance predict where to build
+a shopping centre Capello (2015). By taking these factors in to account
 dentist could predict the most profitable locations in years to come by
@@
 concentrated. To some extent such centre points were the result of
-uncoordinated store-location decisions (\textbf{brueckner2011?}).
-Nevertheless, not coincidentally according to Hotelling's theory (1929).
-He describes the concentration of retail stores by a market desire to
-create a Nash equilibrium. At this point all retailers expected sales
-maximisation outcome cannot be improved by changing position.
-Contradictory to the assumption that consumers will shop at the nearest
-retailer, they are willing to travel significantly further to reach
-certain central places. Marshall (1920) points out another explanation
-for this in his seminal work on agglomeration. That being,intra-industry
-knowledge spillovers as mechanisms leading to formation of industry
-cluster (\textbf{nielsen2021?}). In this case the dental industry
-information spillovers are bounded in geographical space. Hence, the
-need to locate in close proximity. That being the latest equipment, in
-order to work faster or the right personnel to attract more patients.
-Another way to drive sales in this area could be by dentist
-specialisation and patients referrals between dentists.
+uncoordinated store-location decisions Brueckner (2011). Nevertheless,
+not coincidentally according to Hotelling's theory (1929). He describes
+the concentration of retail stores by a market desire to create a Nash
+equilibrium. At this point all retailers expected sales maximisation
+outcome cannot be improved by changing position. Contradictory to the
+assumption that consumers will shop at the nearest retailer, they are
+willing to travel significantly further to reach certain central places.
+Marshall (1920) points out another explanation for this in his seminal
+work on agglomeration. That being,intra-industry knowledge spillovers as
+mechanisms leading to formation of industry cluster Nielsen et al.
+(2021). In this case the dental industry information spillovers are
+bounded in geographical space. Hence, the need to locate in close
+proximity. That being the latest equipment, in order to work faster or
+the right personnel to attract more patients. Another way to drive sales
+in this area could be by dentist specialisation and patients referrals
+between dentists.
 
@@
 market transactions. Whereas, geographical proximity reduces transaction
-cost through several mechanism (\textbf{nielsen2021?}). The market
+cost through several mechanism Nielsen et al. (2021). The market
 transactions can be extended to the endogenous location of specialized
-human capital (\textbf{nielsen2021?}). ``It is well known among
-economists and other social scientists that large cities have
-disproportionately large shares of highly educated workers, and the
-trend has been growing in recent decades.'' (\textbf{bakerBaker?}).
-Presumably, larger cities would then be more attractive to dentists in
-terms of cost and quality of specialised labour in smaller towns.
-Especially, those trying to achieve spatial monopoly end up loosing
-profit by overpaying for labour or in worst case scenario not finding
-enough dentists labour and end up compromising on quality. In both cases
-growth potential would be compromised.
+human capital Nielsen et al. (2021). ``It is well known among economists
+and other social scientists that large cities have disproportionately
+large shares of highly educated workers, and the trend has been growing
+in recent decades.'' (\textbf{bakerBaker?}). Presumably, larger cities
+would then be more attractive to dentists in terms of cost and quality
+of specialised labour in smaller towns. Especially, those trying to
+achieve spatial monopoly end up loosing profit by overpaying for labour
+or in worst case scenario not finding enough dentists labour and end up
+compromising on quality. In both cases growth potential would be
+compromised.
 
@@
 location. Hence, the marketplace and business districts give consumers
-low search and switching costs (\textbf{capello2015?}). The fact that
-owners of stores seem to prefer highly concentrated shopping. More-so,
-it suggests that the sales volumes outweigh the loss attributed to
-greater price competition@capello2015.
+low search and switching costs Capello (2015). The fact that owners of
+stores seem to prefer highly concentrated shopping. More-so, it suggests
+that the sales volumes outweigh the loss attributed to greater price
+competition Capello (2015).
 
@@
 Owners of shopping malls ``orchestrate'' the process of retail
-agglomeration that happens naturally in towns and cities
-(\textbf{brueckner2011?}). Naturally, the same retail agglomeration and
-Nash equilibrium may occur in shopping malls if not more intensely as
-the shops are in closer proximity. Furthermore, the potential for more
-knowledge spillovers, higher volume of market transaction and more
-competition. When retail agglomeration is orchestrated by the owner of a
-shopping mall, the strength and direction of such externalities are
-taken into account (\textbf{brueckner2011?}). Much so, because the price
-of rent is dependent on the overall revenue of a shopping mall. Hence,
-the owner of the mall wants to choose the mix of stores and their sizes
-taking these externalities, attributes and anchor stores attractiveness
-into account (\textbf{brueckner2011?}).
+agglomeration that happens naturally in towns and cities Brueckner
+(2011). Naturally, the same retail agglomeration and Nash equilibrium
+may occur in shopping malls if not more intensely as the shops are in
+closer proximity. Furthermore, the potential for more knowledge
+spillovers, higher volume of market transaction and more competition.
+When retail agglomeration is orchestrated by the owner of a shopping
+mall, the strength and direction of such externalities are taken into
+account Brueckner (2011). Much so, because the price of rent is
+dependent on the overall revenue of a shopping mall. Hence, the owner of
+the mall wants to choose the mix of stores and their sizes taking these
+externalities, attributes and anchor stores attractiveness into account
+Brueckner (2011).
 
@@
 
-Spatial data analysis is a rapidly growing area in Statistics
-(\textbf{haslett1992?}). The Journal of the statistics society define
-such data as being objects in space, which objects have physical
-location. Furthermore, the analyses are spatial if these locations are
-relevant to the interpretation of the data (\textbf{haining1998?}).
+Spatial data analysis is a rapidly growing area in Statistics Haslett
+(1992). The Journal of the statistics society define such data as being
+objects in space, which objects have physical location. Furthermore, the
+analyses are spatial if these locations are relevant to the
+interpretation of the data Haining, Wise, and Ma (1998).
 
@@
 \texttt{it\ will\ involve\ little\ more.\ Graphical\ techniques\ for\ examining\ such\ maps\ are\ thus\ a\ central\ part\ of\ the\ methodology.\ A\ number\ of\ such\ techniques\ are\ discussed."}
-(\textbf{haslett1992?}).
+Haslett (1992).
 
@@
 average cost of check ups is between four to eight hundred Norwegian
-Krone's {``Tannlegeforeningen''} (n.d.). Bertrand model also take price
-as competitive advantage in to consideration.
+Krone's (\textbf{tannlege?}). Bertrand model also take price as
+competitive advantage in to consideration.
 
@@
 options, and they were often located in suburban areas or on the
-outskirts of cities {``Home - NCSC''} (n.d.). In resent years, there has
-been a trend towards larger and more comprehensive shopping malls in
-Norway {``Home - NCSC''} (n.d.).
+outskirts of cities (\textbf{home-n?}). In resent years, there has been
+a trend towards larger and more comprehensive shopping malls in Norway
+(\textbf{home-n?}).
 
@@
 businesses, while other raise concerns about the environmental impact of
-these large scale commercial developments {``Kan Vin Og Sprit Skape
-Byutvikling?''} (2014).
+these large scale commercial developments (\textbf{kanvin2014?}).
 
@@
 paper. As the wine monopoly are anchor stores they may affect the
-attractiveness of the shopping mall {``Kan Vin Og Sprit Skape
-Byutvikling?''} (2014). The Norwegian government regulates the
-placements of wine monopoly stores in shopping malls to prevent
-completion with town centres omsorgsdepartementet (2022). More
-importantly, the government may take in to account the potential impact
-of exciting retail centres in the region. This regional economic
-approach aim to promote sustainable development and balance economic
-growth across the country {``Kan Vin Og Sprit Skape Byutvikling?''}
-(2014).
+attractiveness of the shopping mall (\textbf{kanvin2014?}). The
+Norwegian government regulates the placements of wine monopoly stores in
+shopping malls to prevent completion with town centres
+(\textbf{omsorgsdepartementet2022?}). More importantly, the government
+may take in to account the potential impact of exciting retail centres
+in the region. This regional economic approach aim to promote
+sustainable development and balance economic growth across the country
+(\textbf{kanvin2014?}).
 
@@
 \begin{CSLReferences}{1}{0}
-\leavevmode\vadjust pre{\hypertarget{ref-home-n}{}}%
-{``Home - NCSC.''} n.d.
-\href{https://ncscnordic.org/,\%20https://ncscnordic.org/}{https://ncscnordic.org/,
-https://ncscnordic.org/}.
+\leavevmode\vadjust pre{\hypertarget{ref-brueckner2011}{}}%
+Brueckner, Jan K. 2011. \emph{Lectures on Urban Economics}. MIT Press.
 
-\leavevmode\vadjust pre{\hypertarget{ref-kanvin2014}{}}%
-{``Kan Vin Og Sprit Skape Byutvikling?''} 2014.
-\url{https://www.aftenposten.no/kultur/i/oEy0/kan-vin-og-sprit-skape-byutvikling}.
+\leavevmode\vadjust pre{\hypertarget{ref-capello2011}{}}%
+Capello, Roberta. 2011. {``{[}PDF{]}location, Regional Growth and Local
+Development Theories,''} January.
+\url{https://www.academia.edu/48547951/_PDF_Location_Regional_Growth_and_Local_Development_Theories}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-omsorgsdepartementet2022}{}}%
-omsorgsdepartementet, Helse-og. 2022. {``Vinmonopolordningen.''}
-\emph{Regjeringen.no}. regjeringen.no.
-\url{https://www.regjeringen.no/no/tema/helse-og-omsorg/psykisk-helse/innsikt/vinmonopolordningen/id2528183/}.
+\leavevmode\vadjust pre{\hypertarget{ref-capello2015}{}}%
+---------. 2015. \emph{Regional Economics}. 2nd ed. Routledge.
+\url{https://doi.org/10.4324/9781315720074}.
 
-\leavevmode\vadjust pre{\hypertarget{ref-tannlege}{}}%
-{``Tannlegeforeningen.''} n.d. \url{https://www.tannlegeforeningen.no/}.
+\leavevmode\vadjust pre{\hypertarget{ref-davis2020}{}}%
+Davis, Donald R., and Jonathan I. Dingel. 2020. {``The Comparative
+Advantage of Cities.''} \emph{Journal of International Economics} 123
+(March): 103291. \url{https://doi.org/10.1016/j.jinteco.2020.103291}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-duranton2004}{}}%
+Duranton, Gilles, and Diego Puga. 2004. {``Chapter 48 -
+Micro-Foundations of Urban Agglomeration Economies.''} In \emph{Handbook
+of Regional and Urban Economics}, edited by J. Vernon Henderson and
+Jacques-François Thisse, 4:2063--2117. Cities and Geography. Elsevier.
+\url{https://doi.org/10.1016/S1574-0080(04)80005-1}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-haining1998}{}}%
+Haining, R., S. Wise, and J. Ma. 1998. {``Exploratory Spatial Data
+Analysis.''} \emph{Journal of the Royal Statistical Society: Series D
+(The Statistician)} 47 (3): 457--69.
+\url{https://doi.org/10.1111/1467-9884.00147}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-haslett1992}{}}%
+Haslett, John. 1992. {``Spatial Data Analysis---Challenges.''}
+\emph{Journal of the Royal Statistical Society: Series D (The
+Statistician)} 41 (3): 271--84. \url{https://doi.org/10.2307/2348549}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-mccann2013}{}}%
+McCann, Philip. 2013. \emph{Modern Urban and Regional Economics}. Second
+edition. Oxford University Press.
+
+\leavevmode\vadjust pre{\hypertarget{ref-nielsen2021}{}}%
+Nielsen, Bo Bernhard, Christian Geisler Asmussen, Cecilie Dohlmann
+Weatherall, and Ditte Håkonsson Lyngemark. 2021. {``Marshall Vs Jacobs
+Agglomeration and the Micro-Location of Foreign and Domestic Firms.''}
+\emph{Cities} 117 (October): 103322.
+\url{https://doi.org/10.1016/j.cities.2021.103322}.
+
+\leavevmode\vadjust pre{\hypertarget{ref-puga2010}{}}%
+Puga, Diego. 2010. {``THE MAGNITUDE AND CAUSES OF AGGLOMERATION
+ECONOMIES.''} \emph{Journal of Regional Science} 50 (1): 203--19.
+\url{https://doi.org/10.1111/j.1467-9787.2009.00657.x}.
 

### 2023-02-19 MSB106_Assignment.qmd
.

--- 2023-02-19 MSB106_Assignment.qmd
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, echo=FALSE}
 library(rgdal)
@@
 
-```{r, message=FALSE}
+```{r, message=FALSE, echo=FALSE}
 NOR_CBD <- read_csv("NOR_CBD.csv")
@@
 
-```{r}
+```{r, echo=FALSE}
 Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
@@
 
-```{r}
+```{r, echo=FALSE, warning=FALSE}
 Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
@@
 
-```{r}
+```{r, echo=FALSE}
 kable(summary(Commune_Data))
@@
 
-```{r}
+```{r, echo=FALSE}
 kable(summary(Dentist_Data[3:6]))
@@
 
-```{r}
+```{r, echo=FALSE}
 kable(summary(Dentist_Data[7:10]))
@@
 
-```{r}
+```{r, echo=FALSE}
 kable(summary(Dentist_Data[11:12]))
@@
 
-```{r}
+```{r, echo=FALSE}
+#| label: fig-1
+#| tbl-cap: "Capita in retail compared to mall distance"
 Commune_Data |>
@@
 
-```{r}
+```{r,echo=FALSE}
+#| label: fig-2
+#| tbl-cap: "Dentist income compared to CBD distance"
 Dentist_Data |>
@@
 
-```{r}
+```{r,echo=FALSE}
 Dentist_Data_LIMIT <- Dentist_Data |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data_LIMIT |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Dentist_Data |>
@@
 
-```{r}
+```{r, echo=FALSE}
 lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
@@
 
-```{r}
+```{r, echo=FALSE}
 huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
@@
 
-```{r}
+```{r, echo=FALSE}
 Den_W_Wine <- Dentist_Data_LIMIT |>
@@
 
-```{r}
+```{r, echo=FALSE}
 Den_WO_Wine <- Dentist_Data_LIMIT |>
@@
 
-```{r}
+```{r, echo=FALSE}
 lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
@@
 
-```{r}
+```{r, echo=FALSE}
 huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")

### 2023-02-19 Retail Agglomeration.qmd
.

@@
 We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
-In the first scatterplot we see some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these figures are accounting figures for an entire group of companies under the same legal name.
+The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
 By ignoring these outliers we are then left with the scatterplot in @fig-3.
-This scatterplot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
-The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall.
+This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
+The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
 Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.

### 2023-02-19 Retail Agglomeration_Theory.qmd
.

--- 2023-02-19 Retail Agglomeration_Theory.qmd
@@
 
-`The topic of research (What was studied)`
-
-`The purpose (What was done)`
-
-`The methodology (how the research was done)`
-
-`Main findings (what was found)`
-
-`Policy and implications (significance of findings)`
-
 ## Introduction
@@
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
-
-`A brief description of the problem and its importance`
-
-`Statement of purpose or what was done`
-
-`A brief description of main results`
-
-`Description of methodology, main challenges and solutions`
-
-`Discretion of the papers contribution in relation to previous research`
-
-`Organization of paper`
 
@@
 It has typically microeconomic foundation and uses theoretical models as well as adopting a statistical and geographical approach @capello2011.
-Furthermore, the theory uses the concept of externalities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approach to economic growth @capello2011.
+Furthermore, the theory uses the concept of externalities as disparities in the spatial distribution of activities, thereby laying the territorial bases for dynamic approaches @capello2011.
 
 Regional growth theory involves spatial aspects of economic growth and territorial distribution of income @capello2011.
-It also involves generating geographical advantages, in terms of easy or difficult access to a particular area @capello2011.
+It also involves generating geographical advantages, in terms of easy or difficult access of a particular area @capello2011.
 
-Furthermore, Keynesian economics emphasizes the importance of consumer demand in driving economic growth.
-This may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
-Subsequently, increasing consumer demand, supporting business in a specific region.
-Promoting economic growth or preventing economic decline @capello2016.
+Keynesian economics also emphasizes the importance of consumer demand in driving economic growth.
+In regional economics, this may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
+Subsequently, By increasing consumer demand, supporting business in a specific region and promoting economic growth or preventing economic decline @capello2016.
 
@@
 The model of potential has the capacity to measure the potential of attractiveness to a place.
-Bigger cities or more heavily populated areas have a stronger gravitational force.
-A possible indicator to predict places of growth potential for dentist practice or shopping malls @capello2015.
+Subsequently, the gravitational force is higher in bigger cities or more heavily populated areas.
+Furthermore, predicting places of future growth potential for firms, shopping malls or public service more specifically dentist practice @capello2015.
 
@@
 Furthermore, no relocation costs and demand only depend on location choice @capello2015.
+
 The location game starts off with the total market of AB, firm A in the middle of location A and firm B starts in the middle of location B.
@@
 Nevertheless, despite increasing the transportation cost for patients.
-Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomeration economies.
-
-**Hotelling-Bertand model (1979)**
-
-The Bertrand model was introduced as early as in 1883 and demonstrate two firms competing by simultaneously setting prices for their homogeneous products.
-Furthermore, the consumers choose the product with the lowest price\@tolotti2020.
-The model assumes that firms have identical production cost and consumers have perfect information about prices @tolotti2020.
-
-By combining elements from the two models a hybrid model of spatial completion was birthed by Salop in 1979 @tolotti2020.
-In this model firms compete both in location and price, with consumers having a transportation cost and firms producing differentiated products @tolotti2020.
+Perhaps the simplest way to explain why there is a natural tendency for retailers to cluster in space; a tendency which may help explain the existence of larger agglomerations.
 
@@
 
-Marshall (1920) broadly divides externalities within agglomeration in three main categories potentially drive sales.
-Firstly, knowledge spill-over within industries or product specific technological knowledge.
+Marshall (1920) broadly divides externalities within agglomeration in three main categories with potential to drive sales.
+Firstly, knowledge spill over within industries or product specific technological knowledge.
 Furthermore, market transactions in terms of value chain transactions with industry-specialized buyers and suppliers.
-Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021.
+Lastly, competition for specialized production factors such as labour and product market competition @nielsen2021Nielsen.
 
-There are solidly established conclusions regarding the existence of agglomeration economies @puga2010.
-However, less proof of their estimated magnitude.
-Hence, identifying the causes of agglomeration economies, is proving more difficult @puga2010.
-Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010.
-@duranton2004 discuss these classifications and identify *learning, sharing* and *matching* as the main causes of agglomeration economies.
+There are solidly established conclusions regarding the excitanse of agglomeration economies @puga2010Puga.
+However, less proof of their estimated magnitude therefore identifying the causes of agglomeration economies is proving more difficult @puga2010Puga.
+Nevertheless, there is a large theoretical literature that develops these mechanisms @puga2010Puga.
+@duranton2004Duranton discuss these classifications and identify learning, sharing and matching as the main causes of agglomeration economies.
 
-A larger market allows for a more efficient sharing of local infrastructure and facilities.
-Therefore, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010.
-Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients in order to drive sales.
-The attraction for the consumers and users of public facilities is overall cost reduction @puga2010.
-Hence, the larger the population sharing facilities the lower the cost per user @puga2010.
+A larger market allows for a more efficient sharing of local infrastructure and facilities, a variety of intermediate input suppliers, or a pool of workers with similar skills @puga2010Puga.
+Despite higher rent the dental industry and shopping malls, seams to reap higher benefits in more populated areas as they are dependent on being located where there is a higher volume of patients to drive sales.
+The attraction for the consumers and users of public facilities is overall cost reduction @puga2010Puga.
+Hence, the larger the population sharing facilities the lower the cost per user @puga2010Puga.
 Presumably, industrial factories and business clusters are more dependent on being close to raw materials and industrial action.
 
-Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010.
+Furthermore, a larger market also allows for a better matching between employers and employees, improved chances of finding suitable and better quality of matches @puga2010Puga.
 Shopping malls require skilled workers to drive sales.
-However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020.
-More-so, cites provide a constant market for specialized skills and more productive work force @puga2010P Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
+However, they are not so dependent on highly educated workers as dentists whom according to resent study, tend to prefer bigger cities (as they are highly educated) @davis2020Davis.
+More-so, cites provide a constant market for specialized skills and more productive work force @puga2010Puga.
+Perhaps, a possibility for higher wages being compensated by more efficient workers in bigger agglomeration economies.
 That said the Norwegian cities might differ from the american cites in the study, as the population and clusters are no where near the same size.
-Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010.
+Lastly, a larger market can also facilitate learning by promoting the development and widespread adaptation of new technologies and business practice @puga2010Puga.
 
-Interactions with experienced workers helps acquire valuable skills.
-Experienced workers remain in cities to share the rent of this learning process @puga2010.
-Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the unintended casual flow of information facilitated by big cities." @puga2010
+*"Interactions with experienced workers helps acquire valuable skills and experienced workers remain in cities to share the rent of this learning process.
+Besides this purposeful transition of knowledge, the information literature on learning in cities has also emphasized the casual and unintended flow of information facilitated by big cities."* @puga2010Puga
 
@@
 
-The Alsonso model demonstrate geographical locations tied up to location.
-Furthermore, It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
-In this case used to get an indicating of where the most profitable location for dental practise.
-
+The model demonstrate geographical locations tied up to location.
+It is an urbanized formulation of the nineteenth century von Thünen model @capello2015.
+Moreover, in this case used to get an indicating of where the most profitable location for dental practise might be located.
 Alonso assumes the existence of a city that cannot be build instantaneously, and therefore of an effective rent-curve from the city centre to the periphery.
-Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain.
-In some cases different from the normal or average price competition @capello2015.
-
+Furthermore, it determines the location for a new firm willing to locate in the city and the profit the firm can obtain, which might be different from the normal or average price competition @capello2015.
 As the Von Thünen assume a uniformed space where all land is equally fertile, this model envisages a city, endowed with infrastructure which cover the entire city in all directions whereas unit transport costs are constant in space.
-More-so, the town or city has a single centre point.
-The city centre of business district (CBD) is generally defined as the most attractive for all firms.
+More-so, the town or city has a single city centre, and the city centre of business district (CBD) is generally defined as the most attractive for all firms.
 @capello2015.
 The model also assume perfect competition and unlimited demand, in other words it is supply-orientated @capello2015.
-Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.
-
-Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
+Furthermore, demonstrating a specific production function with fixed coefficients and constant returns to scale.Rent obtained in the model is the remainder left after transport, production costs and profit has been subtracted from revenues @capello2015.
 The bid-rent curve will in this case demonstrate the prise of rent within Bank and Financial locations, Retail locations, and industrial locations.
-
-![](images/bidrent.gpg.jpg)
 
@@
 The rent realized by the landowner is the envelope of the three rent supply curves @capello2015.
-
 From a Marshallian way of thinking dentists should cluster somewhere in this pink area to reap maximum sales volume.
-Furthermore, they may also increase their profits further out in to other retail agglomeration area.
+Other than locating in the inner circle dentists may also increase their profits further out in to other retail agglomeration area.
 The attractiveness might be high in these places due to economies of scope.
@@
 Subsequently, increasing profit by trading higher transportation cost for lower rent.
-In addition fewer dentists to compete with supporting the hotellings game theory whereas companies more to achieve spatial monopoly @mccann2013McCann.
+In addition fewer dentists to compete with.
+The attendances to break out of the clusters support Hotelling's game theory of spatial competition which suggests that the spatial monopoly power provides an incentive for firms to use location as a competitive weapon to achieve spatial monopoly @mccann2013McCann.
 
@@
 Several locations can then be measures against each other to determine profit maximisation.
-
 These models has the potential to measures the future attractiveness of a place.
 Changes in infrastructure such as easier access to a place might change the attractiveness of a place or move the flow of people.
+An example of this could be Rogfast, improving the connection between Haugalandet and Stavanger.
 Subserviently, increasing the gravitationally pull between the two places.
 Furthermore, The model uses population and distance to measure the attractiveness and relations between two places @capello2015.
-Furthermore, these models have predictive capacity and can be used to estimate the potential impact.
+Rogfast might move the population of Haugesund further south and a new town centre might occur competing with other retail clusters in years to come.
+On the other hand, the flow of people towards Raglamyr might increase until the potential new centre place has grown.
+These models have predictive capacity and can be used to estimate the potential impact.
 In particular, location of new productive activity in an area @capello2015.
 For instance predict where to build a shopping centre @capello2015.
+A new shopping centre further south in Haugaland or expansion of excising retail agglomeration areas might be of interest.
+As a result of increased population.
 By taking these factors in to account dentist could predict the most profitable locations in years to come by predicting the amount of people that will move from one city to another or tow different points of interest within a city.
@@
 The market transactions can be extended to the endogenous location of specialized human capital @nielsen2021.
-"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." @bakerBaker.
-Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
+"It is well known among economists and other social scientists that large cities have disproportionately large shares of highly educated workers, and the trend has been growing in recent decades." Baker (n.d.) Presumably, larger cities would then be more attractive to dentists in terms of cost and quality of specialised labour in smaller towns.
 Especially, those trying to achieve spatial monopoly end up loosing profit by overpaying for labour or in worst case scenario not finding enough dentists labour and end up compromising on quality.
@@
 
-Marshall also points out the competition within the geographical location.
+Marshal also points out the competition within the geographical location.
 Hence, the marketplace and business districts give consumers low search and switching costs @capello2015.
 The fact that owners of stores seem to prefer highly concentrated shopping.
-More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition\@capello2015.
+More-so, it suggests that the sales volumes outweigh the loss attributed to greater price competition.
 
@@
 Hence, the owner of the mall wants to choose the mix of stores and their sizes taking these externalities, attributes and anchor stores attractiveness into account @brueckner2011.
-
 The Norwegian retail market for alcoholic beverages is controlled by the state monopoly.
@@
 
-(Data and how it was collected)
+For this assignment we where provided with different geospatial data in the geographic information software QGIS to calculate the relationship of the different variables that might determine firm activity.
+This geospatial data included Norwegian commune data which contained the administrative boundaries of the Communes, presented in the software as different divisions of lines that distinguish the different commune boundaries.
+We where also provided with different point layers, such as the geometric centers of communes, urban city centers or Central Business Districts (CBD), all businesses in Norway which are regarded as dental services, and all Norwegian shopping malls.
+We where also provided with a QGIS layer with population density estimates, visualized with 100x100 raster graphics.
 
-**Empirical Specification**
+To carry out a more complete data analysis of the influencing factors of the activity of Norwegian dentist services, we had to collect more data than what we already had been presented with.
+We firstly utilized QGIS to calculate different other statistics and variables.
+Regarding the Norwegian shopping malls layer we added the variable "Wine-monopoly" as a false/ true variable to determine which Norwegian malls had a hard liquor store.
+We added the number of stores which are in each mall, to use as an estimate for the malls sizes.
+The malls opening hours have also been included in this layer.
+From SSB, we gathered data on the average municipal sales per inhabitant in retailing for each Norwegian commune.
+Lastly we created 2 kilometer buffer zones around dentists and CBDs to determine population density in these areas.
 
-Included why and why not outliner????
+After calculations in QGIS we transferred our data to RSTUDIO to do the primary statistical analysis of our data.
+The data is then merged into two different data frames.
+The first one containing Norwegian communes as the unit of observation, which includes geometric commune centers to the nearest mall, the retail per capita per commune, the different commune numbers and the size of the nearest mall, including the true/ false "Winemonopoly" variable.
+The second data frame use dentists as the unit of observations.
+This data frame include data on the distances from CBD to dentists, distances from dentists to the nearest mall, including variables of the mall size and the "Winemonopoly" variable, different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
 
-**Estimation strategy**
+To obtain a first impression and a simple understanding of our data, we have decided to make use of scatterplots.
+The scatterplot presented in @fig-1 show the relationship of average spending on retail amongst inhabitants and the distance from geometrical commune centers to nearest malls, which indicate that most malls are in a 50 km radius of commune centers, and that Norwegians on average spend between 50 thousand to 120 thousand on retail.
 
-**Analytical framework**
-
-SEDA (Spatial Explanatory Data Analyses)
-
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
-The Journal of the statistics society define such data as being objects in space, which objects have physical location.
-Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
-
-The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
-Graphical techniques for examining such maps are thus a central part of the methodology.
-A number of such techniques are discussed."` @haslett1992.
-
-`What data are used?`
-
-`What has been calculated?`
+We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
+The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
+By ignoring these outliers we are then left with the scatterplot in @fig-3.
+This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
+The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
+Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
+A decent number of dentists run at a loss, having negative earnings before interest and taxes.
 
@@
 
-According to the Alons's model dentist should locate in the retail area for higher volume of sales.
+According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
 Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.

### 2023-02-19 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
.

@@
 
-`In quantitative studies this section provides details about dataset, variables, instruments and measurement procedures used in the study.
-If the data was collected by the authors this section will describe how the sample was collected.
-If specific instruments were constructed to collect the data they will be described here and included in the appendix.
-Procedures used to measure variables are often described in sufficient detail to permit replication.
-In economic papers, this section describes the empirical model and estimation strategy used by the authors.`
+For this assignment we where provided with different geospatial data in the geographic information software QGIS to calculate the relationship of the different variables that might determine firm activity.
+This geospatial data included Norwegian commune data which contained the administrative boundaries of the Communes, presented in the software as different divisions of lines that distinguish the different commune boundaries.
+We where also provided with different point layers, such as the geometric centers of communes, urban city centers or Central Business Districts (CBD), all businesses in Norway which are regarded as dental services, and all Norwegian shopping malls.
+We where also provided with a QGIS layer with population density estimates, visualized with 100x100 raster graphics.
 
-**Data description**
+To carry out a more complete data analysis of the influencing factors of the activity of Norwegian dentist services, we had to collect more data than what we already had been presented with.
+We firstly utilized QGIS to calculate different other statistics and variables.
+Regarding the Norwegian shopping malls layer we added the variable "Wine-monopoly" as a false/ true variable to determine which Norwegian malls had a hard liquor store.
+We added the number of stores which are in each mall, to use as an estimate for the malls sizes.
+The malls opening hours have also been included in this layer.
+From SSB, we gathered data on the average municipal sales per inhabitant in retailing for each Norwegian commune.
+Lastly we created 2 kilometer buffer zones around dentists and CBDs to determine population density in these areas.
 
-(Data and how it was collected)
+After calculations in QGIS we transferred our data to RSTUDIO to do the primary statistical analysis of our data.
+The data is then merged into two different data frames.
+The first one containing Norwegian communes as the unit of observation, which includes geometric commune centers to the nearest mall, the retail per capita per commune, the different commune numbers and the size of the nearest mall, including the true/ false "Winemonopoly" variable.
+The second data frame use dentists as the unit of observations.
+This data frame include data on the distances from CBD to dentists, distances from dentists to the nearest mall, including variables of the mall size and the "Winemonopoly" variable, different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
+A summary of all the different variables can be found in @tbl-1, @tbl-2, @tbl-3 and @tbl-4.
 
-**Empirical Specification**
+To obtain a first impression and a simple understanding of our data, we have decided to make use of scatterplots.
+The scatterplot presented in @fig-1 show the relationship of average spending on retail amongst inhabitants and the distance from geometrical commune centers to nearest malls, which indicate that most malls are in a 50 km radius of commune centers, and that Norwegians on average spend between 50 thousand to 120 thousand on retail.
 
-Included why and why not outliner????
-
-**Estimation strategy**
-
-**Analytical framework**
-
-SEDA (Spatial Explanatory Data Analyses)
-
-Spatial data analysis is a rapidly growing area in Statistics @haslett1992.
-The Journal of the statistics society define such data as being objects in space, which objects have physical location.
-Furthermore, the analyses are spatial if these locations are relevant to the interpretation of the data @haining1998.
-
-The examination of the data involves the examination of data collected from q-gis maps; `it will involve little more.
-Graphical techniques for examining such maps are thus a central part of the methodology.
-A number of such techniques are discussed."` @haslett1992.
-
-`What data are used?`
-
-`What has been calculated?`
+We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
+The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
+By ignoring these outliers we are then left with the scatterplot in @fig-3.
+This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
+The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
+Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
+A decent number of dentists run at a loss, having negative earnings before interest and taxes.
 

### 2023-02-19 Retail-Agglomeration.tex
.

--- 2023-02-19 Retail-Agglomeration_abstract_introduction_theori_econometric_approach.tex
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, borderline west={3pt}{0pt}{shadecolor}, enhanced, breakable, interior hidden, frame hidden, boxrule=0pt]}{\end{tcolorbox}}\fi
 
@@
 \subsection{Abstract}\label{abstract}}
-
-There are several factors to take in to consideration when deciding
-where a dental practice should be located. Spatial competition is a
-critical consideration in regional economics. Some of the factors
-influencing the decision making is retail agglomeration attributes,
-spatial monopoly and the attractiveness of a particular place of
-interest. This study aims to examine the dental industry and find the
-strategy most profitable for the dental industry.
-
-The study utilize a quantitative research approach, employing spacial
-explanatory data analysis techniques to determine the location decisions
-of firms. Geographical data was collected from locations in Norway
-showing all dentists, central places, retail agglomeration and other
-points of interest such as the government regulated wine monopoly.
-Combined with regional income data a spatial regression analyses was
-used to identify the factors influencing the location decisions of
-firms.
-
-Furthermore, the result indicates the distance to retail agglomeration
-areas and central places playes a significant role in the location
-decisions for firms. Firms are more likely to locate closer to these
-areas, which offer access to a larger customer base and lower cost of
-transportation. However, the results also suggest other factors, such as
-market size and completion also influence the location decisions for
-firms.
-
-The studies findings have particular implications for policy-makers,
-developer, and dental practises seeking to maximize profits through
-strategic decisions. By understanding the importance of proximity to
-retail agglomeration areas and central places, these stakeholders can
-make informed decisions that benefit the local economy and improve
-competitiveness of the region. Perhaps more importantly prevent
-agrassive retail agglomeration areas from ``killing down town''.
 
@@
 
-The theoretical framework for this study draws on Marshall's
-agglomeration principles, which emphasize spatial aspects of economic
-growth. Hotelling's game theory is used to explain spatial competition
-by location. Furthermore, adding the hybrid model whereas firms compete
-both in location and price, with consumers having a transportation cost
-and firms producing differentiated products. Alsonso's bid-rent model
-and gravitational models also provides insights into how dentists or
-other retailer makes location decisions in the context of regional
-growth.
-
-The dental industry provides an interesting industry to study for this
-research because of its unique characteristics. Dental practices require
-a high degree of specialization and expertise. At the same time, they
-need to be easily accessible to consumers. Shopping malls and central
-places offer a continent location for dental practices that said
-competition for prime locations can be intense and push some in to
-spatial monopoly.
-
-The result of this study indicates that dentist's with the highest
-income are usually located in central places in big cities. The findings
-are in line with already exciting theories especially supporting the
-gravitational models and agglomeration attributes. ?????? Proving a
-central location within large cities are the best location to maximize
-profit for Dental practices. ???
-
 \textbf{Hypothesis}
@@
 
+\hypertarget{literature-review}{%
+\subsection{Literature Review}\label{literature-review}}
+
+\hypertarget{previous-literature-on-shopping-malls}{%
+\paragraph{Previous literature on Shopping
+malls}\label{previous-literature-on-shopping-malls}}
+
 \hypertarget{theoretical-framework}{%
-\subsection{Theoretical Framework}\label{theoretical-framework}}
+\subsubsection{Theoretical Framework}\label{theoretical-framework}}
 
@@
 Capello (2011). Furthermore, the theory uses the concept of
-externalities in the spatial distribution of activities, thereby laying
-the territorial bases for dynamic approach to economic growth Capello
+externalities as disparities in the spatial distribution of activities,
+thereby laying the territorial bases for dynamic approaches Capello
 (2011).
@@
 generating geographical advantages, in terms of easy or difficult access
-to a particular area Capello (2011).
-
-Furthermore, Keynesian economics emphasizes the importance of consumer
-demand in driving economic growth. This may involve policies that
-involve consumer spending, such as incentives for buying local products
-or supporting small businesses or in this case preventing the the death
-of down town due to competition from shopping malls. Subsequently,
-increasing consumer demand, supporting business in a specific region.
-Promoting economic growth or preventing economic decline
-(\textbf{capello2016?}).
+of a particular area Capello (2011).
 
@@
 customers, and more productive workers Capello (2011). Furthermore, the
-attractiveness of a central location increase the cost of rent.
-
-Alonso's bid rent model indicates the most profitable location for
-firms. Closer to the centre with agglomeration attributes or in rural
-areas with spacial monopoly and low rent. In gravitational models, the
+attractiveness of a central location increase the cost of rent. Alonso's
+bid rent model indicates the most profitable location for firms. Closer
+to the centre with agglomeration attributes or in rural areas with
+spacial monopoly and low rent. In gravitational models, the
 attractiveness of the retail location, represent the size of the retail
@@
 
-The model of potential has the capacity to measure the potential of
-attractiveness to a place. Bigger cities or more heavily populated areas
-have a stronger gravitational force. A possible indicator to predict
-places of growth potential for dentist practice or shopping malls
-Capello (2015).
+A generalized Models based on Newton law of gravity can be applied to
+estimate flows between two points such as, cities or other points of
+interest Capello (2015). The model of potential has the capacity to
+measure the potential of attractiveness to a place. Furthermore,
+predicting places of future growth potential for firms, shopping malls
+or public service more specifically dentist practice Capello (2015).
 
@@
 single firm in a duopoly environment. Furthermore, no relocation costs
-and demand only depend on location choice Capello (2015). The location
-game starts off with the total market of AB, firm A in the middle of
-location A and firm B starts in the middle of location B. One firm
-starts relocating closer to the other to take some of the customers in
-the other market area Capello (2015). The other respond by doing the
+and demand only depend on location choice Capello (2015).
+
+The location game starts off with the total market of AB, firm A in the
+middle of location A and firm B starts in the middle of location B. One
+firm starts relocating closer to the other to take some of the customers
+in the other market area Capello (2015). The other respond by doing the
 same and the game continues until both end up in the middle of the
@@
 in space; a tendency which may help explain the existence of larger
-agglomeration economies.
-
-\textbf{Hotelling-Bertand model (1979)}
-
-The Bertrand model was introduced as early as in 1883 and demonstrate
-two firms competing by simultaneously setting prices for their
-homogeneous products. Furthermore, the consumers choose the product with
-the lowest price@tolotti2020. The model assumes that firms have
-identical production cost and consumers have perfect information about
-prices (\textbf{tolotti2020?}).
-
-By combining elements from the two models a hybrid model of spatial
-completion was birthed by Salop in 1979 (\textbf{tolotti2020?}). In this
-model firms compete both in location and price, with consumers having a
-transportation cost and firms producing differentiated products
-(\textbf{tolotti2020?}).
+agglomerations.
 
@@
 Marshall (1920) broadly divides externalities within agglomeration in
-three main categories potentially drive sales. Firstly, knowledge
-spill-over within industries or product specific technological
+three main categories with potential to drive sales. Firstly, knowledge
+spill over within industries or product specific technological
 knowledge. Furthermore, market transactions in terms of value chain
@@
 competition for specialized production factors such as labour and
-product market competition Nielsen et al. (2021).
+product market competition (\textbf{nielsen2021Nielsen?}).
 
-There are solidly established conclusions regarding the existence of
-agglomeration economies Puga (2010). However, less proof of their
-estimated magnitude. Hence, identifying the causes of agglomeration
-economies, is proving more difficult Puga (2010). Nevertheless, there is
-a large theoretical literature that develops these mechanisms Puga
-(2010). Duranton and Puga (2004) discuss these classifications and
-identify \emph{learning, sharing} and \emph{matching} as the main causes
-of agglomeration economies.
+There are solidly established conclusions regarding the excistanse of
+agglomeration economies (\textbf{puga2010Puga?}). However, less proof of
+their estimated magnitude therefore identifying the causes of
+agglomeration economies is proving more difficult
+(\textbf{puga2010Puga?}). Nevertheless, there is a large theoretical
+literature that develops these mechanisms (\textbf{puga2010Puga?}).
+(\textbf{duranton2004Duranton?}) discuss these classifications and
+identify learning, sharing and matching as the main causes of
+agglomeration economies.
 
 A larger market allows for a more efficient sharing of local
-infrastructure and facilities. Therefore, a variety of intermediate
-input suppliers, or a pool of workers with similar skills Puga (2010).
-Despite higher rent the dental industry and shopping malls, seams to
-reap higher benefits in more populated areas as they are dependent on
-being located where there is a higher volume of patients in order to
-drive sales. The attraction for the consumers and users of public
-facilities is overall cost reduction Puga (2010). Hence, the larger the
-population sharing facilities the lower the cost per user Puga (2010).
-Presumably, industrial factories and business clusters are more
-dependent on being close to raw materials and industrial action.
+infrastructure and facilities, a variety of intermediate input
+suppliers, or a pool of workers with similar skills
+(\textbf{puga2010Puga?}). Dispite higer rent the dental industry and
+shopping malls, seams to reap higher benefits in more populated areas as
+they are dependent on being located where there is a higher volume of
+patiants to drive sales. The attraction for the consumers and users of
+public facilities is overall cost reduction (\textbf{puga2010Puga?}).
+Hence, the larger the population sharing facilities the lower the cost
+per user (\textbf{puga2010Puga?}). Presumably, industrial factories and
+business clusters are more dependent on being close to raw materials and
+industrial action.
 
@@
 employers and employees, improved chances of finding suitable and better
-quality of matches Puga (2010). Shopping malls require skilled workers
-to drive sales. However, they are not so dependent on highly educated
-workers as dentists whom according to resent study, tend to prefer
-bigger cities (as they are highly educated) Davis and Dingel (2020).
-More-so, cites provide a constant market for specialized skills and more
-productive work force Puga (2010). Perhaps, a possibility for higher
-wages being compensated by more efficient workers in bigger
-agglomeration economies. That said the Norwegian cities might differ
-from the american cites in the study, as the population and clusters are
-no where near the same size. Lastly, a larger market can also facilitate
-learning by promoting the development and widespread adaptation of new
-technologies and business practice Puga (2010).
+quality of matches (\textbf{puga2010Puga?}). Shopping malls require
+skilled workers to drive sales. However, they are not so dependent on
+highly educated workers as dentists whom according to resent study, tend
+to prefer bigger cities (as they are highly educated)
+(\textbf{davis2020Davis?}). More-so, cites provide a constant market for
+specialized skills and more productive work force
+(\textbf{puga2010Puga?}). Perhaps, a possibility for higher wages being
+compensated by more efficient workers in bigger agglomeration economies.
+That said the Norwegian cities might differ from the american cites in
+the study, as the population and clusters are no where near the same
+size. Lastly, a larger market can also facilitate learning by promoting
+the development and widespread adaptation of new technologies and
+business practice (\textbf{puga2010Puga?}).
 
-Interactions with experienced workers helps acquire valuable skills.
-Experienced workers remain in cities to share the rent of this learning
-process Puga (2010). Besides this purposeful transition of knowledge,
+\emph{``Interactions with experienced workers helps acquire valuable
+skills and experienced workers remain in cities to share the rent of
+this learning process. Besides this purposeful transition of knowledge,
 the information literature on learning in cities has also emphasized the
-unintended casual flow of information facilitated by big cities.'' Puga
-(2010)
+casual and unintended flow of information facilitated by big cities.''}
+(\textbf{puga2010Puga?})
 
@@
 
-The Alsonso model demonstrate geographical locations tied up to
-location. Furthermore, It is an urbanized formulation of the nineteenth
-century von Thünen model Capello (2015). In this case used to get an
-indicating of where the most profitable location for dental practise.
-
+The model demonstrate geographical locations tied up to location. It is
+an urbanized formulation of the nineteenth century von Thünen model
+Capello (2015). Moreover, in this case used to get an indicating of
+where the most profitable location for dental practise might be located.
 Alonso assumes the existence of a city that cannot be build
@@
 new firm willing to locate in the city and the profit the firm can
-obtain. In some cases different from the normal or average price
-competition Capello (2015).
+obtain, which might be different from the normal or average price
+competition Capello (2015). As the Von Thünen assume a uniformed space
+where all land is equally fertile, this model envisages a city, endowed
+with infrastructure which cover the entire city in all directions
+whereas unit transport costs are constant in space. More-so, the town or
+city has a single city centre, and the city centre of business district
+(CBD) is generally defined as the most attractive for all firms. Capello
+(2015). The model also assume perfect competition and unlimited demand,
+in other words it is supply-orientated Capello (2015). Furthermore,
+demonstrating a specific production function with fixed coefficients and
+constant returns to scale.Rent obtained in the model is the remainder
+left after transport, production costs and profit has been subtracted
+from revenues Capello (2015). The bid-rent curve will in this case
+demonstrate the prise of rent within Bank and Financial locations,
+Retail locations, and industrial locations.
 
-As the Von Thünen assume a uniformed space where all land is equally
-fertile, this model envisages a city, endowed with infrastructure which
-cover the entire city in all directions whereas unit transport costs are
-constant in space. More-so, the town or city has a single centre point.
-The city centre of business district (CBD) is generally defined as the
-most attractive for all firms. Capello (2015). The model also assume
-perfect competition and unlimited demand, in other words it is
-supply-orientated Capello (2015). Furthermore, demonstrating a specific
-production function with fixed coefficients and constant returns to
-scale.
-
-Rent obtained in the model is the remainder left after transport,
-production costs and profit has been subtracted from revenues Capello
-(2015). The bid-rent curve will in this case demonstrate the prise of
-rent within Bank and Financial locations, Retail locations, and
-industrial locations.
-
-\includegraphics{images/bidrent.gpg.jpg}
+\includegraphics{images/bidrent.png}
 
@@
 landowner is the envelope of the three rent supply curves Capello
-(2015).
-
-From a Marshallian way of thinking dentists should cluster somewhere in
-this pink area to reap maximum sales volume. Furthermore, they may also
-increase their profits further out in to other retail agglomeration
-area. The attractiveness might be high in these places due to economies
-of scope. As the positive benefits from the Marshallian agglomeration
-attributes play out. Subsequently, increasing profit by trading higher
-transportation cost for lower rent. In addition fewer dentists to
-compete with supporting the hotellings game theory whereas companies
-more to achieve spatial monopoly McCann (2013).
+(2015). From a Marshallian way of thinking dentists should cluster
+somewhere in this pink area to reap maximum sales volume. Other than
+locating in the inner circle dentists may also increase their profits
+further out in to other retail agglomeration area. The attractiveness
+might be high in these places due to economies of scope. As the positive
+benefits from the Marshallian agglomeration attributes play out.
+Subsequently, increasing profit by trading higher transportation cost
+for lower rent. In addition fewer dentists to compete with. The
+attendances to break out of the clusters support Hotelling's game theory
+of spatial competition which suggests that the spatial monopoly power
+provides an incentive for firms to use location as a competitive weapon
+to achieve spatial monopoly (\textbf{mccann2013McCann?}).
 
@@
 influence the central points in diverse ways such as commuter movements,
-diffusion of knowledge, corporations and network and personal
+disfustion of knowledge, corporations and network and personal
 relationship Capello (2015). More-so, they are generated by the
@@
 then be measures against each other to determine profit maximisation.
-
 These models has the potential to measures the future attractiveness of
-a place. Changes in infrastructure such as easier access to a place
-might change the attractiveness of a place or move the flow of people.
-Subserviently, increasing the gravitationally pull between the two
-places. Furthermore, The model uses population and distance to measure
-the attractiveness and relations between two places Capello (2015).
-Furthermore, these models have predictive capacity and can be used to
-estimate the potential impact. In particular, location of new productive
-activity in an area Capello (2015). For instance predict where to build
-a shopping centre Capello (2015). By taking these factors in to account
-dentist could predict the most profitable locations in years to come by
-predicting the amount of people that will move from one city to another
-or tow different points of interest within a city.
+a place. Changes in infrastructure such easier access to a place might
+change the attractiveness of a place or move the flow of people. An
+example of this could be Rogfast, improving the connection between
+Haugalandet and Stavanger. Subserviently, increasing the gravitationally
+pull between the two places. Furthermore, The model uses population and
+distance to measure the attractiveness and relations between two places
+Capello (2015). Rogfast might move the population of Haugesund further
+south and a new town centre might occur competing with other retail
+clusters in years to come. On the other hand the flow of people towards
+Raglamyr might increase until the potential new centre place has grown.
+These models have predictive capacity and can be used to estimate the
+potential impact. In particular, location of new productive activity in
+an area Capello (2015). For instance predict where to build a shopping
+centre Capello (2015). A new shopping centre further south in Haugaland
+or expansion of excising retail agglomeration areas might be of
+interest. As a result of increased population. By taking these factors
+in to account dentist could predict the most profitable locations in
+years to come by predicting the amount of people that will move from one
+city to another or tow different points of interest within a city.
 
@@
 large shares of highly educated workers, and the trend has been growing
-in recent decades.'' (\textbf{bakerBaker?}). Presumably, larger cities
-would then be more attractive to dentists in terms of cost and quality
-of specialised labour in smaller towns. Especially, those trying to
-achieve spatial monopoly end up loosing profit by overpaying for labour
-or in worst case scenario not finding enough dentists labour and end up
+in recent decades.'' Baker (n.d.) Presumably, larger cities would then
+be more attractive to dentists in terms of cost and quality of
+specialised labour in smaller towns. Especially, those trying to achieve
+spatial monopoly end up loosing profit by overpaying for labour or in
+worst case scenario not finding enough dentists labour and end up
 compromising on quality. In both cases growth potential would be
@@
 
-Marshall also points out the competition within the geographical
+Marshal also points out the competition within the geographical
 location. Hence, the marketplace and business districts give consumers
@@
 that the sales volumes outweigh the loss attributed to greater price
-competition Capello (2015).
+competition.
 
@@
 externalities, attributes and anchor stores attractiveness into account
-Brueckner (2011).
-
-The Norwegian retail market for alcoholic beverages is controlled by the
-state monopoly. Anchor-stores such as the Wine-monopoly bring in a
-higher volume of consumers (\textbf{lai2013Lai?}). For example, shoppers
-visiting the Wine-monopoly in a mall may also visit a clothing store,
-and even find it beneficial to use public services at the same time.
-Nevertheless, they are limited in numbers and needs government approval.
-Each municipality must apply to get the wine monopoly in their area. One
-of the factors considered in this process is the death of city centre. A
-possible reason for this is so called ``predatory malls'' which are
-purposely built to overtake the exciting local market. Perhaps more so
-if malls cater for cinema, restaurants, bars and other attributes in
-conjunction with the city centre. Not to mention the upper hand of easy
-road access and free parking.
+Brueckner (2011). The Norwegian retail market for alcoholic beverages is
+controlled by the state monopoly. Anchor-stores such as the
+Wine-monopoly bring in a higher volume of consumers
+(\textbf{lai2013Lai?}). For example, shoppers visiting the Wine-monopoly
+in a mall may also visit a clothing store, and even find it beneficial
+to use public services at the same time. Nevertheless, they are limited
+in numbers and needs government approval. Each municipality must apply
+to get the wine monopoly in their area. One of the factors considered in
+this process is the death of city centre. A possible reason for this is
+so called ``predatory malls'' which are purposely built to overtake the
+exciting local market. Perhaps more so if malls cater for cinema,
+restaurants, bars and other attributes in conjunction with the city
+centre. Not to mention the upper hand of easy road access and free
+parking.
 
@@
 
-(Data and how it was collected)
+For this assignment we where provided with different geospatial data in
+the geographic information software QGIS to calculate the relationship
+of the different variables that might determine firm activity. This
+geospatial data included Norwegian commune data which contained the
+administrative boundaries of the Communes, presented in the software as
+different divisions of lines that distinguish the different commune
+boundaries. We where also provided with different point layers, such as
+the geometric centers of communes, urban city centers or Central
+Business Districts (CBD), all businesses in Norway which are regarded as
+dental services, and all Norwegian shopping malls. We where also
+provided with a QGIS layer with population density estimates, visualized
+with 100x100 raster graphics.
+
+To carry out a more complete data analysis of the influencing factors of
+the activity of Norwegian dentist services, we had to collect more data
+than what we already had been presented with. We firstly utilized QGIS
+to calculate different other statistics and variables. Regarding the
+Norwegian shopping malls layer we added the variable ``Wine-monopoly''
+as a false/ true variable to determine which Norwegian malls had a hard
+liquor store. We added the number of stores which are in each mall, to
+use as an estimate for the malls sizes. The malls opening hours have
+also been included in this layer. From SSB, we gathered data on the
+average municipal sales per inhabitant in retailing for each Norwegian
+commune. Lastly we created 2 kilometer buffer zones around dentists and
+CBDs to determine population density in these areas.
+
+After calculations in QGIS we transferred our data to RSTUDIO to do the
+primary statistical analysis of our data. The data is then merged into
+two different data frames. The first one containing Norwegian communes
+as the unit of observation, which includes geometric commune centers to
+the nearest mall, the retail per capita per commune, the different
+commune numbers and the size of the nearest mall, including the true/
+false ``Winemonopoly'' variable. The second data frame use dentists as
+the unit of observations. This data frame include data on the distances
+from CBD to dentists, distances from dentists to the nearest mall,
+including variables of the mall size and the ``Winemonopoly'' variable,
+different significant accounting figures for the different dentist
+businesses, such as total sales revenue and EBIT (Earnings Before
+Interests and Taxes).
+
+To obtain a first impression and a simple understanding of our data, we
+have decided to make use of scatterplots. The scatterplot presented in
+\textbf{?@fig-1} show the relationship of average spending on retail
+amongst inhabitants and the distance from geometrical commune centers to
+nearest malls, which indicate that most malls are in a 50 km radius of
+commune centers, and that Norwegians on average spend between 50
+thousand to 120 thousand on retail.
+
+We furthermore have used plots to look at the relationship between the
+income of dentist's businesses and the distance from CBDs. The
+scatterplot in \textbf{?@fig-2} have some extreme figures of 1 billion+
+income which we choose to ignore and deem as outliers, as we notice that
+these observations are accounting figures for an entire group of
+companies under the same legal name, and not individual companies. By
+ignoring these outliers we are then left with the scatterplot in
+\textbf{?@fig-3}. This plot shows a tendency of a negative relationship
+between the variables, which might suggest that dentists might earn more
+closer to CBDs. The last scatterplots we have chosen to calculate is
+concerning the relationship of dentists EBIT which is earnings before
+interest and taxes, compared to their distances to the nearest shopping
+mall in \textbf{?@fig-4}. Just as the previous scatterplot, we also see
+here a tendency of a negative relationship between the variables. A
+decent number of dentists run at a loss, having negative earnings before
+interest and taxes.
 
@@
 
-Included why and why not outliner????
+\textbf{Econometric Approach}
 
@@
 
-\hypertarget{econometric-approach}{%
-\subsection{\texorpdfstring{\textbf{Econometric
-Approach}}{Econometric Approach}}\label{econometric-approach}}
+\hypertarget{results}{%
+\subsection{Results}\label{results}}
 
-The questions addressed are related to the geographical space of Norway.
-Norway is a country with relatively small population and large
-heterogeneous land area. Furthermore, the population is decentralized
-with the majority of population living in urban areas. This unique
-spatial distribution of land might give a possible explanation to
-diversions from the simple theoretical models used in this paper.
-Whereas one of the assumptions are homogeneous landscape. Nevertheless,
-the way economies are organised comes partly from the emergence of
-centralities within a city. Transportation networks have the potential
-to increase sprawl depending on how the urban system evolves gradually.
-
-The results from the bid rent configuration from the city centre gave
-the opportunity to explore other locations in each municipality's.
-Having an urban system of multiple centres and applying the same logic
-of the model to other retail agglomeration areas. Resulting in finding
-relationships based on three different types of distance measurements.
-Firstly, the distance between Dental Practice and CBD then, dentist and
-other retail agglomeration areas and lastly amongst dentists. Finding
-that most dentist are either located close to a City, Central Business
-District or a shopping mall. Additionally, dentist also located close to
-each other making some sort of industrial cluster. Supporting,
-Marshall's agglomeration theory on knowledge spillovers, increasing the
-sills to attract more patients.
-
-Furthermore, when using the dentist income (y) and comparing the
-distance from the mall (x), signs of patterns emerge showing that income
-is greater when the dentist is placed closer to the shopping mall.
-Furthermore, decreasing as the distance gets bigger. Faster
-transportation can increase convenience of commuting, increasing rent
-commuters from the city centre are willing to pay, and therefore
-increasing the area of developed land without this being necessary
-corresponding to an increase in population. Hence supporting findings;
-when the distance reach around 10km from the mall the income starts to
-increase before it decrease again around 20km. Also in line with the
-Keynesian theory suggesting that this pattern might take place as the
-transportation cost for individuals further away from the Shopping mall
-does not increase to much by travelling the extra distance. Hotelling
-also describes this case where the consumer are willing to travel
-further to reach a certain central place.
-
-According to the Alons's model dentist should locate in the retail area
-for higher volume of sales. Dentist being placed further out in to other
-retail agglomeration areas can also be a good strategy. However, the
-dentist would then rely on the area having less competition so the rate
-between competition and population density are proportional. Our
-findings in QGIS shows that in areas of high population density, the
-number of dentist also increase. The consumer/patient then has more to
-choose between and the dentist also has a bigger scale of potential
-customers. Where in smaller areas it is less competition and the dentist
-would appear to have somewhat of a monopoly since the transportation
-cost to the second closes dentist are noticeable higher.
-
-Furthermore, being aware of the fact that going to the dentist is a
-public services and most people get assigned to their closest dentist
-based on their living address. This also takes away some of the
-competition and would support placement in more rural areas. More-so, is
-important to notice that when passed 18 years of age, you could risk
-loosing your local public dentist due to lack of capacity. Suggesting
-some dentist might seek closer to CBD areas to capture most of the
-remaining part of the population. Business, public places and industry
-can have a positive effect on revenue since this is areas where most of
-the population spends their time during their day, either going to work,
-school and other errand. Reducing the transportation cost for the
-customer as they are already located in these areas even if they do not
-live here.
-
-The bid rent models are under the assumption that Dentists use spatial
-completion to maximise profit. Nevertheless, the cost of dental care in
-Norway can vary depending on a number of other factors such as the type
-of treatment, the location of the dental practice and the dentist
-experience and qualifications. The Norwegian Dental Association rate the
-average cost of check ups is between four to eight hundred Norwegian
-Krone's (\textbf{tannlege?}). Bertrand model also take price as
-competitive advantage in to consideration.
-
-Viser data noe om indikasjon på dyrere priser i nærheten av større byer
-of hvor det er stedsbestemt monopol? Tjener tannleger i større byer mer?
-Her kan vi også skrive om gravitasjons modeller.
-
-Another Mashiallian point to be made could be the advantage of industry
-clusters. Taken for granted that the data on Dentist includes both
-``normal Dentists'' and specialist. Where in this case a ``normal
-dentist'' would for example recommend one of their customers to go see a
-specialist to get braces. The customer would then probably prefer to go
-to the closes specialist. The access to work force in these areas would
-also be good and will be within the same geographic area.
-
-Shopping malls have a relatively short history in Norway, with the first
-shopping mall opening in the late sixties. Prior to the most retail
-activities took place in the city centre. Furthermore, these malls
-typically featured a mix of shops, restaurants and entertainment
-options, and they were often located in suburban areas or on the
-outskirts of cities (\textbf{home-n?}). In resent years, there has been
-a trend towards larger and more comprehensive shopping malls in Norway
-(\textbf{home-n?}).
-
-However, the growth of shopping malls in Norway has not been without
-controversy. Some critics argue that the development of large shopping
-malls has led to the decline of traditional shopping streets and local
-businesses, while other raise concerns about the environmental impact of
-these large scale commercial developments (\textbf{kanvin2014?}).
-
-Information about Wine monopoly locations has also been studied in this
-paper. As the wine monopoly are anchor stores they may affect the
-attractiveness of the shopping mall (\textbf{kanvin2014?}). The
-Norwegian government regulates the placements of wine monopoly stores in
-shopping malls to prevent completion with town centres
-(\textbf{omsorgsdepartementet2022?}). More importantly, the government
-may take in to account the potential impact of exciting retail centres
-in the region. This regional economic approach aim to promote
-sustainable development and balance economic growth across the country
-(\textbf{kanvin2014?}).
-
-Hva viser informasjonen om vinmonopolet?
+\hypertarget{discussion}{%
+\subsection{Discussion}\label{discussion}}
 
@@
 \subsection{Conclusion}\label{conclusion}}
+
+\hypertarget{acknowledgements}{%
+\subsection{Acknowledgements}\label{acknowledgements}}
 
@@
 \url{https://doi.org/10.4324/9781315720074}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-davis2020}{}}%
-Davis, Donald R., and Jonathan I. Dingel. 2020. {``The Comparative
-Advantage of Cities.''} \emph{Journal of International Economics} 123
-(March): 103291. \url{https://doi.org/10.1016/j.jinteco.2020.103291}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-duranton2004}{}}%
-Duranton, Gilles, and Diego Puga. 2004. {``Chapter 48 -
-Micro-Foundations of Urban Agglomeration Economies.''} In \emph{Handbook
-of Regional and Urban Economics}, edited by J. Vernon Henderson and
-Jacques-François Thisse, 4:2063--2117. Cities and Geography. Elsevier.
-\url{https://doi.org/10.1016/S1574-0080(04)80005-1}.
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-puga2010}{}}%
-Puga, Diego. 2010. {``THE MAGNITUDE AND CAUSES OF AGGLOMERATION
-ECONOMIES.''} \emph{Journal of Regional Science} 50 (1): 203--19.
-\url{https://doi.org/10.1111/j.1467-9787.2009.00657.x}.
-
 \end{CSLReferences}

### 2023-02-19 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
Commit.

@@
 ---
+
+```{r, message=FALSE, echo=FALSE}
+library(rgdal)
+library(dplyr)
+library(RSQLite)
+library(sf)
+library(tidyverse)
+library(readr)
+library(ggplot2)
+library(kableExtra)
+library(huxtable)
+```
+
+```{r, message=FALSE, echo=FALSE}
+NOR_CBD <- read_csv("NOR_CBD.csv")
+Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
+Zonal_CBD <- read_csv("Zonal_CBD.csv")
+Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
+NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv") |>
+  rename("HubName" = fid)
+Zonal_Den <- read.csv("Zonal_Den.csv") |>
+  select(fid, X_sum)
+Dist_Comp_Den <- read.csv("Dist_Comp_Den.csv") |>
+  select(fid, distance)
+NOR_KOMM <- read_csv("NOR_KOMM.csv")
+NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
+```
+
+```{r, echo=FALSE}
+Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
+  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
+  rename("knr" = kommunenummer, "DistMal" = HubDist, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
+```
+
+```{r, echo=FALSE}
+Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
+  select(knr, DistMal, HubName, Turnover_capita_retail, shop, Vinmonopolet) |>
+  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
+```
+
+```{r, echo=FALSE, warning=FALSE}
+Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
+Commune_Data$Winemonopoly <- as.integer(as.logical(Commune_Data$Winemonopoly))
+```
+
+```{r, echo=FALSE}
+Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
+  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
+  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubName" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
+```
+
+```{r, echo=FALSE}
+Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
+```
+
+```{r, echo=FALSE}
+Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
+  rename("DistComp" = distance)
+```
+
+```{r, echo=FALSE}
+Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
+  select(fid, Juridisk.n, Antall.ans, Driftsres, Sum.salgsi, latlong, DistCBD, DistMal, X_sum, DistComp, shop, Vinmonopolet) |>
+  rename("SizeMall" = shop, "Winemonopoly" = Vinmonopolet)
+```
+
+```{r, echo=FALSE, warning=FALSE}
+Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
+Dentist_Data$Winemonopoly <- as.integer(as.logical(Dentist_Data$Winemonopoly))
+```
+
+```{r,echo=FALSE}
+Dentist_Data_LIMIT <- Dentist_Data |>
+  filter(Sum.salgsi < 100000000, DistCBD < 250,
+         Driftsres < 20000000, Driftsres > -5000000, DistMal < 80)
+```
 
@@
 
+```{r, echo=FALSE}
+lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
+```
+
+```{r, echo=FALSE}
+#|label: tbl-4
+#|tbl-cap: "Dentist Revenue MRM"
+huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
+```
+
 ## Conclusion
@@
 ## References
+
+## Appendix
+
+```{r, echo=FALSE}
+#|label: tbl-1
+#|tbl-cap: "Commune data summary"
+kable(summary(Commune_Data))
+```
+
+```{r, echo=FALSE}
+#|label: tbl-2
+#|tbl-cap: "Dentist data summary"
+kable(summary(Dentist_Data[3:6]))
+```
+
+```{r, echo=FALSE}
+#|label: tbl-3
+#|tbl-cap: "Dentist data summary"
+kable(summary(Dentist_Data[7:10]))
+```
+
+```{r, echo=FALSE}
+#|label: tbl-4
+#|tbl-cap: "Dentist data summary"
+kable(summary(Dentist_Data[11:12]))
+```
+
+```{r, echo=FALSE}
+#| label: fig-1
+#| fig-cap: "Capita in retail compared to mall distance"
+Commune_Data |>
+  ggplot(aes(x = SizeMall, y = Turnover_capita_retail, colour = SizeMall)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
+  theme_bw()
+```
+
+```{r,echo=FALSE}
+#| label: fig-2
+#| fig-cap: "Dentist income compared to CBD distance"
+Dentist_Data |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r, echo=FALSE}
+#| label: fig-3
+#| fig-cap: "Dentist income compared to CBD distance"
+Dentist_Data_LIMIT |>
+  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
+  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
+  theme_bw()
+```
+
+```{r, echo=FALSE}
+#| label: fig-4
+#| fig-cap: "Dentist EBIT compared to Mall distance"
+Dentist_Data |>
+  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
+  geom_point(lwd = .9) +
+  labs(x = "Distance from Mall", y = "EBIT") +
+  theme_bw()
+```

### 2023-02-19 Retail-Agglomeration_abstract_introduction_theori_econometric_approach.tex
Commit.

@@
 
+\usepackage{booktabs}
+\usepackage{longtable}
+\usepackage{array}
+\usepackage{multirow}
+\usepackage{wrapfig}
+\usepackage{float}
+\usepackage{colortbl}
+\usepackage{pdflscape}
+\usepackage{tabu}
+\usepackage{threeparttable}
+\usepackage{threeparttablex}
+\usepackage[normalem]{ulem}
+\usepackage{makecell}
+\usepackage{xcolor}
+\usepackage{caption}
+\usepackage{graphicx}
+\usepackage{siunitx}
+\usepackage{hhline}
+\usepackage{calc}
+\usepackage{tabularx}
+\usepackage{adjustbox}
+\usepackage{hyperref}
 \makeatletter
@@
 \maketitle
-\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, enhanced, boxrule=0pt, frame hidden, interior hidden, breakable, sharp corners]}{\end{tcolorbox}}\fi
+\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, frame hidden, breakable, boxrule=0pt, borderline west={3pt}{0pt}{shadecolor}, enhanced, interior hidden]}{\end{tcolorbox}}\fi
+
+\begin{verbatim}
+Warning: One or more parsing issues, see `problems()` for details
+\end{verbatim}
 
@@
 
-\texttt{In\ quantitative\ studies\ this\ section\ provides\ details\ about\ dataset,\ variables,\ instruments\ and\ measurement\ procedures\ used\ in\ the\ study.\ If\ the\ data\ was\ collected\ by\ the\ authors\ this\ section\ will\ describe\ how\ the\ sample\ was\ collected.\ If\ specific\ instruments\ were\ constructed\ to\ collect\ the\ data\ they\ will\ be\ described\ here\ and\ included\ in\ the\ appendix.\ Procedures\ used\ to\ measure\ variables\ are\ often\ described\ in\ sufficient\ detail\ to\ permit\ replication.\ In\ economic\ papers,\ this\ section\ describes\ the\ empirical\ model\ and\ estimation\ strategy\ used\ by\ the\ authors.}
+For this assignment we where provided with different geospatial data in
+the geographic information software QGIS to calculate the relationship
+of the different variables that might determine firm activity. This
+geospatial data included Norwegian commune data which contained the
+administrative boundaries of the Communes, presented in the software as
+different divisions of lines that distinguish the different commune
+boundaries. We where also provided with different point layers, such as
+the geometric centers of communes, urban city centers or Central
+Business Districts (CBD), all businesses in Norway which are regarded as
+dental services, and all Norwegian shopping malls. We where also
+provided with a QGIS layer with population density estimates, visualized
+with 100x100 raster graphics.
 
-\textbf{Data description}
+To carry out a more complete data analysis of the influencing factors of
+the activity of Norwegian dentist services, we had to collect more data
+than what we already had been presented with. We firstly utilized QGIS
+to calculate different other statistics and variables. Regarding the
+Norwegian shopping malls layer we added the variable ``Wine-monopoly''
+as a false/ true variable to determine which Norwegian malls had a hard
+liquor store. We added the number of stores which are in each mall, to
+use as an estimate for the malls sizes. The malls opening hours have
+also been included in this layer. From SSB, we gathered data on the
+average municipal sales per inhabitant in retailing for each Norwegian
+commune. Lastly we created 2 kilometer buffer zones around dentists and
+CBDs to determine population density in these areas.
 
-(Data and how it was collected)
+After calculations in QGIS we transferred our data to RSTUDIO to do the
+primary statistical analysis of our data. The data is then merged into
+two different data frames. The first one containing Norwegian communes
+as the unit of observation, which includes geometric commune centers to
+the nearest mall, the retail per capita per commune, the different
+commune numbers and the size of the nearest mall, including the true/
+false ``Winemonopoly'' variable. The second data frame use dentists as
+the unit of observations. This data frame include data on the distances
+from CBD to dentists, distances from dentists to the nearest mall,
+including variables of the mall size and the ``Winemonopoly'' variable,
+different significant accounting figures for the different dentist
+businesses, such as total sales revenue and EBIT (Earnings Before
+Interests and Taxes). A summary of all the different variables can be
+found in \textbf{?@tbl-1}, \textbf{?@tbl-2}, \textbf{?@tbl-3} and
+\textbf{?@tbl-4}.
 
-\textbf{Empirical Specification}
+To obtain a first impression and a simple understanding of our data, we
+have decided to make use of scatterplots. The scatterplot presented in
+Figure~\ref{fig-1} show the relationship of average spending on retail
+amongst inhabitants and the distance from geometrical commune centers to
+nearest malls, which indicate that most malls are in a 50 km radius of
+commune centers, and that Norwegians on average spend between 50
+thousand to 120 thousand on retail.
 
-Included why and why not outliner????
-
-\textbf{Estimation strategy}
-
-\textbf{Analytical framework}
-
-SEDA (Spatial Explanatory Data Analyses)
-
-Spatial data analysis is a rapidly growing area in Statistics Haslett
-(1992). The Journal of the statistics society define such data as being
-objects in space, which objects have physical location. Furthermore, the
-analyses are spatial if these locations are relevant to the
-interpretation of the data Haining, Wise, and Ma (1998).
-
-The examination of the data involves the examination of data collected
-from q-gis maps;
-\texttt{it\ will\ involve\ little\ more.\ Graphical\ techniques\ for\ examining\ such\ maps\ are\ thus\ a\ central\ part\ of\ the\ methodology.\ A\ number\ of\ such\ techniques\ are\ discussed."}
-Haslett (1992).
-
-\texttt{What\ data\ are\ used?}
-
-\texttt{What\ has\ been\ calculated?}
+We furthermore have used plots to look at the relationship between the
+income of dentist's businesses and the distance from CBDs. The
+scatterplot in Figure~\ref{fig-2} have some extreme figures of 1
+billion+ income which we choose to ignore and deem as outliers, as we
+notice that these observations are accounting figures for an entire
+group of companies under the same legal name, and not individual
+companies. By ignoring these outliers we are then left with the
+scatterplot in Figure~\ref{fig-3}. This plot shows a tendency of a
+negative relationship between the variables, which might suggest that
+dentists might earn more closer to CBDs. The last scatterplots we have
+chosen to calculate is concerning the relationship of dentists EBIT
+which is earnings before interest and taxes, compared to their distances
+to the nearest shopping mall in Figure~\ref{fig-4}. Just as the previous
+scatterplot, we also see here a tendency of a negative relationship
+between the variables. A decent number of dentists run at a loss, having
+negative earnings before interest and taxes.
 
@@
 
+  \providecommand{\huxb}[2]{\arrayrulecolor[RGB]{#1}\global\arrayrulewidth=#2pt}
+  \providecommand{\huxvb}[2]{\color[RGB]{#1}\vrule width #2pt}
+  \providecommand{\huxtpad}[1]{\rule{0pt}{#1}}
+  \providecommand{\huxbpad}[1]{\rule[-#1]{0pt}{#1}}
+
+\begin{table}[ht]
+\begin{centerbox}
+\begin{threeparttable}
+
+\setlength{\tabcolsep}{0pt}
+\begin{tabular}{l l}
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{c!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\centering \hspace{6pt} Dentist Revenue \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} (Intercept) \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 3616068.187 *** \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (373568.663)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} DistComp \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -105.228\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (152.737)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} DistMal \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -13619.260\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (32749.952)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} SizeMall \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 10063.209 *\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (4523.296)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} DistCBD \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 7969.413\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (9756.588)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} X\_sum \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 13.257\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (7.925)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} DistMal:SizeMall \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -106.204\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (724.299)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} DistCBD:X\_sum \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} -1.105\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt}  \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} (1.065)\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{255, 255, 255}{0.4}}->{\huxb{0, 0, 0}{0.4}}-}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} N \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 1364\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+
+\multicolumn{1}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} R2 \hspace{6pt}\huxbpad{6pt}} &
+\multicolumn{1}{r!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedleft \hspace{6pt} 0.008\hphantom{0}\hphantom{0}\hphantom{0}\hphantom{0} \hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{>{\huxb{0, 0, 0}{0.8}}->{\huxb{0, 0, 0}{0.8}}-}
+\arrayrulecolor{black}
+
+\multicolumn{2}{!{\huxvb{0, 0, 0}{0}}l!{\huxvb{0, 0, 0}{0}}}{\huxtpad{6pt + 1em}\raggedright \hspace{6pt} Note:  *** p $<$ 0.001;  ** p $<$ 0.01;  * p $<$ 0.05 T statistics in brackets.
+\hspace{6pt}\huxbpad{6pt}} \tabularnewline[-0.5pt]
+
+\hhline{}
+\arrayrulecolor{black}
+\end{tabular}
+\end{threeparttable}\par\end{centerbox}
+
+\end{table}
+
 \hypertarget{conclusion}{%
@@
 \hypertarget{references}{%
-\subsection*{References}\label{references}}
-\addcontentsline{toc}{subsection}{References}
+\subsection{References}\label{references}}
+
+\hypertarget{appendix}{%
+\subsection{Appendix}\label{appendix}}
+
+\begin{tabular}{l|l|l|l|l|l|l}
+\hline
+  &     knr &    DistMal &    HubName & Turnover\_capita\_retail &    SizeMall &  Winemonopoly\\
+\hline
+ & Length:435 & Min.   : 0.01854 & Min.   :  6.0 & Min.   : 17366 & Min.   :  2.00 & Min.   :0.0000\\
+\hline
+ & Class :character & 1st Qu.: 9.41403 & 1st Qu.:220.0 & 1st Qu.: 59250 & 1st Qu.: 13.00 & 1st Qu.:0.0000\\
+\hline
+ & Mode  :character & Median :18.87373 & Median :291.0 & Median : 83203 & Median : 25.00 & Median :1.0000\\
+\hline
+ & NA & Mean   :23.82982 & Mean   :280.8 & Mean   : 84719 & Mean   : 37.22 & Mean   :0.5701\\
+\hline
+ & NA & 3rd Qu.:32.67127 & 3rd Qu.:363.0 & 3rd Qu.:106257 & 3rd Qu.: 50.00 & 3rd Qu.:1.0000\\
+\hline
+ & NA & Max.   :98.37355 & Max.   :424.0 & Max.   :218728 & Max.   :206.00 & Max.   :1.0000\\
+\hline
+ & NA & NA & NA & NA & NA's   :124 & NA's   :107\\
+\hline
+\end{tabular}
+
+\begin{tabular}{l|l|l|l|l}
+\hline
+  &   Antall.ans &   Driftsres &   Sum.salgsi &   latlong\\
+\hline
+ & Min.   :   0.00 & Min.   :-30319000 & Min.   :    -28000 & Length:5740\\
+\hline
+ & 1st Qu.:   1.00 & 1st Qu.:    -5000 & 1st Qu.:    658250 & Class :character\\
+\hline
+ & Median :   1.00 & Median :   330000 & Median :   3176500 & Mode  :character\\
+\hline
+ & Mean   :  39.26 & Mean   :  4842620 & Mean   :  50123495 & NA\\
+\hline
+ & 3rd Qu.:   3.00 & 3rd Qu.:  1201750 & 3rd Qu.:   6248250 & NA\\
+\hline
+ & Max.   :6870.00 & Max.   :108100000 & Max.   :1058102000 & NA\\
+\hline
+ & NA & NA's   :3838 & NA's   :3838 & NA\\
+\hline
+\end{tabular}
+
+\begin{tabular}{l|l|l|l|l}
+\hline
+  &    DistCBD &    DistMal &     X\_sum &    DistComp\\
+\hline
+ & Min.   :  0.0046 & Min.   :  0.0008 & Min.   :    0 & Min.   :    0.0\\
+\hline
+ & 1st Qu.:  1.3912 & 1st Qu.:  0.3145 & 1st Qu.: 6129 & 1st Qu.:    0.0\\
+\hline
+ & Median :  6.7067 & Median :  0.9164 & Median :14602 & Median :    0.0\\
+\hline
+ & Mean   : 20.5000 & Mean   :  4.6331 & Mean   :22455 & Mean   :  614.4\\
+\hline
+ & 3rd Qu.: 19.8885 & 3rd Qu.:  2.5220 & 3rd Qu.:28904 & 3rd Qu.:  197.7\\
+\hline
+ & Max.   :470.1005 & Max.   :105.8867 & Max.   :92591 & Max.   :60577.8\\
+\hline
+\end{tabular}
+
+\begin{tabular}{l|l|l}
+\hline
+  &    SizeMall &  Winemonopoly\\
+\hline
+ & Min.   :  1.0 & Min.   :0.0000\\
+\hline
+ & 1st Qu.: 21.0 & 1st Qu.:0.0000\\
+\hline
+ & Median : 40.0 & Median :1.0000\\
+\hline
+ & Mean   : 51.1 & Mean   :0.6429\\
+\hline
+ & 3rd Qu.: 70.0 & 3rd Qu.:1.0000\\
+\hline
+ & Max.   :206.0 & Max.   :1.0000\\
+\hline
+ & NA's   :1388 & NA's   :1274\\
+\hline
+\end{tabular}
+
+\begin{verbatim}
+Warning: Removed 124 rows containing missing values (geom_point).
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics{Retail-Agglomeration_abstract_introduction_theori_econometric_approach_files/figure-pdf/fig-1-1.pdf}
+
+}
+
+\caption{\label{fig-1}Capita in retail compared to mall distance}
+
+\end{figure}
+
+\begin{verbatim}
+Warning: Removed 3838 rows containing missing values (geom_point).
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics{Retail-Agglomeration_abstract_introduction_theori_econometric_approach_files/figure-pdf/fig-2-1.pdf}
+
+}
+
+\caption{\label{fig-2}Dentist income compared to CBD distance}
+
+\end{figure}
+
+\begin{figure}
+
+{\centering \includegraphics{Retail-Agglomeration_abstract_introduction_theori_econometric_approach_files/figure-pdf/fig-3-1.pdf}
+
+}
+
+\caption{\label{fig-3}Dentist income compared to CBD distance}
+
+\end{figure}
+
+\begin{verbatim}
+Warning: Removed 3838 rows containing missing values (geom_point).
+\end{verbatim}
+
+\begin{figure}
+
+{\centering \includegraphics{Retail-Agglomeration_abstract_introduction_theori_econometric_approach_files/figure-pdf/fig-4-1.pdf}
+
+}
+
+\caption{\label{fig-4}Dentist EBIT compared to Mall distance}
+
+\end{figure}
 
@@
 
-\leavevmode\vadjust pre{\hypertarget{ref-haining1998}{}}%
-Haining, R., S. Wise, and J. Ma. 1998. {``Exploratory Spatial Data
-Analysis.''} \emph{Journal of the Royal Statistical Society: Series D
-(The Statistician)} 47 (3): 457--69.
-\url{https://doi.org/10.1111/1467-9884.00147}.
-
-\leavevmode\vadjust pre{\hypertarget{ref-haslett1992}{}}%
-Haslett, John. 1992. {``Spatial Data Analysis---Challenges.''}
-\emph{Journal of the Royal Statistical Society: Series D (The
-Statistician)} 41 (3): 271--84. \url{https://doi.org/10.2307/2348549}.
-
 \leavevmode\vadjust pre{\hypertarget{ref-mccann2013}{}}%

### 2023-02-19 MSB106_Assignment.qmd
.

@@
 ```{r, echo=FALSE}
+#|label: tbl-1
+#|tbl-cap: "Commune data summary"
 kable(summary(Commune_Data))
@@
 ```{r, echo=FALSE}
+#|label: tbl-2
+#|tbl-cap: "Dentist data summary"
 kable(summary(Dentist_Data[3:6]))
@@
 ```{r, echo=FALSE}
+#|label: tbl-3
+#|tbl-cap: "Dentist data summary"
 kable(summary(Dentist_Data[7:10]))
@@
 ```{r, echo=FALSE}
+#|label: tbl-4
+#|tbl-cap: "Dentist data summary"
 kable(summary(Dentist_Data[11:12]))
@@
 #| label: fig-1
-#| tbl-cap: "Capita in retail compared to mall distance"
+#| fig-cap: "Capita in retail compared to mall distance"
 Commune_Data |>
@@
 #| label: fig-2
-#| tbl-cap: "Dentist income compared to CBD distance"
+#| fig-cap: "Dentist income compared to CBD distance"
 Dentist_Data |>
@@
 ```{r, echo=FALSE}
+#| label: fig-3
+#| fig-cap: "Dentist income compared to CBD distance"
 Dentist_Data_LIMIT |>
@@
 ```{r, echo=FALSE}
+#| label: fig-4
+#| fig-cap: "Dentist EBIT compared to Mall distance"
 Dentist_Data |>
@@
 ```{r, echo=FALSE}
+#|label: tbl-4
+#|tbl-cap: "Dentist Revenue MRM"
 huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
@@
 ```{r, echo=FALSE}
+#|label: tbl-5
+#|tbl-cap: "Turnover per capita retail MRM"
 huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")

### 2023-02-19 Retail Agglomeration_Theory.qmd

@@
 editor: visual
-bibliography: references.bib
+bibliography: [references.bib, Mena.bib]
 ---
@@
 
-We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
+We furthermore have used plots to look at the relationship between the income of dentist's businesses and the distance from CBDs.
 The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
@@
 
+The size of the mall also has a positive effect on revenue for dentist practice, as larger malls tends to have have positive effect on the revenue.
+Larger malls gives the customer more options to choose from and are often placed closer to a larger CBD which might be more attractive for customers to visit, and then also explaining why customers could be more interested in traveling further to visit these malls instead of visiting the local mall in a smaller CBD.
+
+More so, the reasoning for a more profitable business close to a shopping malls can be explain by customers beeing able to combine their dentis visit with shopping afterwards, maybe in a way to threat them self.
+You can find some similiarities in marketing strategies for grocery stores in the sense where they alwyas have a well planned system of where all their products are.
+Where for example gums and grab and go snacks, will always be placed next to the cashier.
+these are usually cheap, small and quick to consume, and something that you dont usually look for, but since you are already there, you might aswell.
+Same thoughts goes for people who visit the dentists in or fairly close to a mall.
+they are already there and might aswell be done with their shopping.
+This tradeoff can be highly beneficial for the mall itself, considering the patients going to the dentists then also become customers of the mall by default.
+
+\
 According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
@@
 
+Findings suggest that locating further away from a CBD has a positive effect on the revenue for dentist and also that moving further away from the competition has a negative effect on revenue.
+These findings could come across as surprising at first, but taking in ti comparison that some of the CBD´s obtained from QGIS could be categorized as relatively small, one explanation to this would possibly be that the data contains a larger number of smaller CBD´s with fewer dentist offices that also has less potential customers and a weaker income which pulls the result in one direction.
+The fact that the population density (X_sum) also shows a positive effect on the revenue strengthen this theory.
+Areas with a higher population density tends to increase the revenue for dentist.
+
+\
 Furthermore, being aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
@@
 Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
-The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's @tannlege.
+The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's \@tannlege.
 Bertrand model also take price as competitive advantage in to consideration.
-
-Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
-Tjener tannleger i større byer mer?
-Her kan vi også skrive om gravitasjons modeller.
 
@@
 
+\
 Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
 Prior to the most retail activities took place in the city centre.
-Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
-In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway @home-n.
+Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities \@home-n.
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway \@home-n.
 
 However, the growth of shopping malls in Norway has not been without controversy.
-Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments @kanvin2014.
+Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments \@kanvin2014.
 
 Information about Wine monopoly locations has also been studied in this paper.
-As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall @kanvin2014.
-The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres @omsorgsdepartementet2022.
+As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall \@kanvin2014.
+The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres \@omsorgsdepartementet2022.
 More importantly, the government may take in to account the potential impact of exciting retail centres in the region.
-This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
+This regional economic approach aim to promote sustainable development and balance economic growth across the country \@kanvin2014.
 
-Hva viser informasjonen om vinmonopolet?
+Findings around Wine monopoly shows a negative effect on turnover per capita in retail.
+In other words, the Wine monopoly has a bad influence for the turnover in retail.
+The explanation could be as mentioned, since Wine monopoly could be categorized as an anchor store, it drag customers from malls and areas where there is no winemonoply over to the areas and malls where a wine monopoly exist, only helping out the stores surrounding.
+Another reasoning could be that the winemonopoly drags attention away from the other shopping options, resulting in a negative effect for the turnover for other retailers.
+
+is important to notice that the municipal council is more or less in charge of where Wine monopoly is allowed to operate.
+There are several criterias and factors that are taken into consideration before anything gets approved to prevent areas to dry out.
+[@kvikstadSeListenDisse2019]
 
@@
 
+There are several critical factors that effect the income for dental practice where our result shows that the income tends to be higher located close to a mall or inside the mall itself.
+A larger mall also has a positive effect of the income.
+Theorys crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice.
+However, it is important to notice that a high revenue does not equal good economic results as more populated and central areas tends to have higher fees for example rent.
+The Wine monopoly also appears to have negative effect on turnover per capita in retail.
+
 ## References
+
+::: {#ref}
+:::

### 2023-02-19 Retail Agglomeration_Theory.qmd
mona bib

@@
 editor: visual
-bibliography: [references.bib, Mena.bib]
+bibliography: [references.bib, Mena.bib, Mona_Retail_Agglomeration.bib]
 ---

### 2023-02-20 Retail Agglomeration_Theory.qmd
idk

@@
 A larger mall also has a positive effect of the income.
-Theorys crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice.
+Theorys crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice and why there are certain areas where there is a cluster of dentist.
 However, it is important to notice that a high revenue does not equal good economic results as more populated and central areas tends to have higher fees for example rent.

### 2023-02-20 Retail Agglomeration_Theory.qmd
kildr

@@
 Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
-The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's \@tannlege.
+The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's [@TannlegeforeningenTannlegeforeningen].
 Bertrand model also take price as competitive advantage in to consideration.
@@
 Prior to the most retail activities took place in the city centre.
-Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities \@home-n.
-In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway \@home-n.
+Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities [@NCSCNCSC2018].
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway [@NCSCNCSC2018].
 

### 2023-02-20 MSB106_Assignment.qmd
last one

--- 2023-02-19 MSB106_Assignment.qmd
@@
 
-```{r, message=FALSE, echo=FALSE}
+```{r, message=FALSE}
 library(rgdal)
@@
 
-```{r, message=FALSE, echo=FALSE}
+```{r, message=FALSE}
 NOR_CBD <- read_csv("NOR_CBD.csv")
@@
 
-```{r, echo=FALSE}
+```{r}
 Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Commune_Data <- inner_join(Commune_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
@@
 
-```{r, echo=FALSE, warning=FALSE}
+```{r}
 Commune_Data$SizeMall <- as.numeric(Commune_Data$SizeMall)
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data <- inner_join(Dentist_Data, Zonal_Den, by = "fid")
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data <- inner_join(Dentist_Data, Dist_Comp_Den, by = "fid") |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data <- left_join(Dentist_Data, NOR_OSM_SHOP_MAL, by = "HubName") |>
@@
 
-```{r, echo=FALSE, warning=FALSE}
+```{r}
 Dentist_Data$SizeMall <- as.numeric(Dentist_Data$SizeMall)
@@
 
-```{r, echo=FALSE}
+```{r}
 kable(summary(Commune_Data))
@@
 
-```{r, echo=FALSE}
+```{r}
 kable(summary(Dentist_Data[3:6]))
@@
 
-```{r, echo=FALSE}
+```{r}
 kable(summary(Dentist_Data[7:10]))
@@
 
-```{r, echo=FALSE}
+```{r}
 kable(summary(Dentist_Data[11:12]))
@@
 
-```{r, echo=FALSE}
-#| label: fig-1
-#| tbl-cap: "Capita in retail compared to mall distance"
+```{r}
 Commune_Data |>
@@
 
-```{r,echo=FALSE}
-#| label: fig-2
-#| tbl-cap: "Dentist income compared to CBD distance"
+```{r}
 Dentist_Data |>
@@
 
-```{r,echo=FALSE}
+```{r}
 Dentist_Data_LIMIT <- Dentist_Data |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data_LIMIT |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Dentist_Data |>
@@
 
-```{r, echo=FALSE}
+```{r}
 lm1 <- lm(Sum.salgsi ~ DistComp + DistMal + SizeMall + (DistMal*SizeMall) + DistCBD + X_sum + (DistCBD*X_sum), data = Dentist_Data_LIMIT)
@@
 
-```{r, echo=FALSE}
+```{r}
 huxreg(list("Dentist Revenue" = lm1), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
@@
 
-```{r, echo=FALSE}
+```{r}
 Den_W_Wine <- Dentist_Data_LIMIT |>
@@
 
-```{r, echo=FALSE}
+```{r}
 Den_WO_Wine <- Dentist_Data_LIMIT |>
@@
 
-```{r, echo=FALSE}
+```{r}
 lm2 <- lm(Turnover_capita_retail ~ DistMal + SizeMall + (DistMal*SizeMall) + Winemonopoly, data = Commune_Data)
@@
 
-```{r, echo=FALSE}
+```{r}
 huxreg(list("Turnover per capita in retail" = lm2), statistics = c(N = "nobs", R2 = "r.squared"), const. = "(Intercept)", note = "Note: {stars} T statistics in brackets.")
@@
 ## Econometric approach
+
+The questions addressed are related to the geographical space of Norway.
+Norway is a country with relatively small population and large heterogeneous land area.
+Furthermore, the population is decentralized with the majority of population living in urban areas.
+This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
+Whereas one of the assumptions are homogeneous landscape.
+Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
+Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
+
+The results from the bid rent configuration from the city centre gave the opportunity to explore other locations in each municipality's.
+Having an urban system of multiple centres and applying the same logic of the model to other retail agglomeration areas.
+Resulting in finding relationships based on three different types of distance measurements.
+Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
+Additionally, dentist also located close to each other making some sort of industrial cluster.
+Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
+
+Furthermore, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
+Furthermore, decreasing as the distance gets bigger.
+Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
+Hence supporting findings; when the distance reach around 10km from the mall the income starts to increase before it decrease again around 20km.
+Also in line with the Keynesian theory suggesting that this pattern might take place as the transportation cost for individuals further away from the Shopping mall does not increase to much by travelling the extra distance.
+Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.
+
+\
+\
+
+The size of the mall also has a positive effect on revenue for dentist practice, as larger malls tends to have have positive effect on the revenue.
+Larger malls gives the customer more options to choose from and are often placed closer to a larger CBD which might be more attractive for customers to visit, and then also explaining why customers could be more interested in traveling further to visit these malls instead of visiting the local mall in a smaller CBD.
+
+More so, the reasoning for a more profitable business close to a shopping malls can be explain by customers beeing able to combine their dentis visit with shopping afterwards, maybe in a way to threat them self.
+You can find some similiarities in marketing strategies for grocery stores in the sense where they alwyas have a well planned system of where all their products are.
+Where for example gums and grab and go snacks, will always be placed next to the cashier.
+these are usually cheap, small and quick to consume, and something that you dont usually look for, but since you are already there, you might aswell.
+Same thoughts goes for people who visit the dentists in or fairly close to a mall.
+they are already there and might aswell be done with their shopping.
+This tradeoff can be highly beneficial for the mall itself, considering the patients going to the dentists then also become customers of the mall by default.
+
+\
+According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
+Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.
+However, the dentist would then rely on the area having less competition so the rate between competition and population density are proportional.
+Our findings in QGIS shows that in areas of high population density, the number of dentist also increase.
+The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
+Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation cost to the second closes dentist are noticeable higher.
+
+Findings suggest that locating further away from a CBD has a positive effect on the revenue for dentist and also that moving further away from the competition has a negative effect on revenue.
+These findings could come across as surprising at first, but taking in ti comparison that some of the CBD´s obtained from QGIS could be categorized as relatively small, one explanation to this would possibly be that the data contains a larger number of smaller CBD´s with fewer dentist offices that also has less potential customers and a weaker income which pulls the result in one direction.
+The fact that the population density (X_sum) also shows a positive effect on the revenue strengthen this theory.
+Areas with a higher population density tends to increase the revenue for dentist.
+
+\
+Furthermore, being aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentist based on their living address.
+This also takes away some of the competition and would support placement in more rural areas.
+More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
+Suggesting some dentist might seek closer to CBD areas to capture most of the remaining part of the population.
+Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
+Reducing the transportation cost for the customer as they are already located in these areas even if they do not live here.
+
+The bid rent models are under the assumption that Dentists use spatial completion to maximise profit.
+Nevertheless, the cost of dental care in Norway can vary depending on a number of other factors such as the type of treatment, the location of the dental practice and the dentist experience and qualifications.
+The Norwegian Dental Association rate the average cost of check ups is between four to eight hundred Norwegian Krone's \@tannlege.
+Bertrand model also take price as competitive advantage in to consideration.
+
+Another Mashiallian point to be made could be the advantage of industry clusters.
+Taken for granted that the data on Dentist includes both "normal Dentists" and specialist.
+Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
+The customer would then probably prefer to go to the closes specialist.
+The access to work force in these areas would also be good and will be within the same geographic area.
+
+\
+Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
+Prior to the most retail activities took place in the city centre.
+Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities \@home-n.
+In resent years, there has been a trend towards larger and more comprehensive shopping malls in Norway \@home-n.
+
+However, the growth of shopping malls in Norway has not been without controversy.
+Some critics argue that the development of large shopping malls has led to the decline of traditional shopping streets and local businesses, while other raise concerns about the environmental impact of these large scale commercial developments \@kanvin2014.
+
+Information about Wine monopoly locations has also been studied in this paper.
+As the wine monopoly are anchor stores they may affect the attractiveness of the shopping mall \@kanvin2014.
+The Norwegian government regulates the placements of wine monopoly stores in shopping malls to prevent completion with town centres \@omsorgsdepartementet2022.
+More importantly, the government may take in to account the potential impact of exciting retail centres in the region.
+This regional economic approach aim to promote sustainable development and balance economic growth across the country \@kanvin2014.
+
+Findings around Wine monopoly shows a negative effect on turnover per capita in retail.
+In other words, the Wine monopoly has a bad influence for the turnover in retail.
+The explanation could be as mentioned, since Wine monopoly could be categorized as an anchor store, it drag customers from malls and areas where there is no winemonoply over to the areas and malls where a wine monopoly exist, only helping out the stores surrounding.
+Another reasoning could be that the winemonopoly drags attention away from the other shopping options, resulting in a negative effect for the turnover for other retailers.
+
+is important to notice that the municipal council is more or less in charge of where vinmonopolet is allowed to operate.
+There are several criterias and factors that are taken into consideration before anything gets approved to prevent areas to dry out.
+
+\
+
+\
+
+\
+
+\

### 2023-02-20 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
rettet litt

@@
 
-Furthermore, the result indicates the distance to retail agglomeration areas and central places playes a significant role in the location decisions for firms.
-Firms are more likely to locate closer to these areas, which offer access to a larger customer base and lower cost of transportation.
+Furthermore, the result indicates the distance to retail agglomeration areas and central places plays a significant role in the location decisions for firms.
+Retailers are more likely to locate closer to these areas, which offer access to a larger customer base and lower cost of transportation.
 However, the results also suggest other factors, such as market size and completion also influence the location decisions for firms.
@@
 By understanding the importance of proximity to retail agglomeration areas and central places, these stakeholders can make informed decisions that benefit the local economy and improve competitiveness of the region.
-Perhaps more importantly prevent agrassive retail agglomeration areas from "killing down town".
+Perhaps more importantly prevent aggressive retail agglomeration areas from "killing down town".
 
@@
 
-Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
-More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
+Our main task in this assignment is to perform an analysis of geospatial determinants of firms activity.
+More specifically we are to focus on the Norwegian dental industry in this regard, and to look in to how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
 As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"
@@
 The findings are in line with already exciting theories especially supporting the gravitational models and agglomeration attributes.
-?????? Proving a central location within large cities are the best location to maximize profit for Dental practices.
-???
+Proving a central location within large cities are the best location to maximize profit for Dental practices.
 
@@
 This may involve policies that involve consumer spending, such as incentives for buying local products or supporting small businesses or in this case preventing the the death of down town due to competition from shopping malls.
-Subsequently, increasing consumer demand, supporting business in a specific region.
+Subsequently, dental practices may increasing consumer demand and support business in a specific region.
 Promoting economic growth or preventing economic decline @capello2016.
@@
 The data is then merged into two different data frames.
-The first one containing Norwegian communes as the unit of observation, which includes geometric commune centers to the nearest mall, the retail per capita per commune, the different commune numbers and the size of the nearest mall, including the true/ false "Winemonopoly" variable.
+The first one containing Norwegian communes as the unit of observation, which includes geometric commune centres to the nearest mall, the retail per capita per commune, the different commune numbers and the size of the nearest mall, including the true/ false "Wine monopoly" variable.
 The second data frame use dentists as the unit of observations.
-This data frame include data on the distances from CBD to dentists, distances from dentists to the nearest mall, including variables of the mall size and the "Winemonopoly" variable, different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
+This data frame include data on the distances from CBD to dentists, distances from dentists to the nearest mall, including variables of the mall size and the "Wine monopoly" variable, different significant accounting figures for the different dentist businesses, such as total sales revenue and EBIT (Earnings Before Interests and Taxes).
 A summary of all the different variables can be found in @tbl-1, @tbl-2, @tbl-3 and @tbl-4.
 
-To obtain a first impression and a simple understanding of our data, we have decided to make use of scatterplots.
-The scatterplot presented in @fig-1 show the relationship of average spending on retail amongst inhabitants and the distance from geometrical commune centers to nearest malls, which indicate that most malls are in a 50 km radius of commune centers, and that Norwegians on average spend between 50 thousand to 120 thousand on retail.
+To obtain a first impression and a simple understanding of our data, we have decided to make use of scatter-plots.
+The scatter-plot presented in @fig-1 show the relationship of average spending on retail amongst inhabitants and the distance from geometrical commune centres to nearest malls, which indicate that most malls are in a 50 km radius of commune centres, and that Norwegians on average spend between 50 thousand to 120 thousand on retail.
 
-We furthermore have used plots to look at the relationship between the income of dentist\'s businesses and the distance from CBDs.
-The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
-By ignoring these outliers we are then left with the scatterplot in @fig-3.
+Furthermore, plots were used to look at the relationship between the income of dentist's businesses and the distance from CBDs.
+The scatter-plot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
+By ignoring these outliers we are then left with the scatter-plot in @fig-3.
 This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
-The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
-Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
+The last scatter-plots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
+Just as the previous scatter-plot, we also see here a tendency of a negative relationship between the variables.
 A decent number of dentists run at a loss, having negative earnings before interest and taxes.
@@
 Furthermore, the population is decentralized with the majority of population living in urban areas.
-This unique spatial distribution of land might give a possible explanation to diversions from the simple theoretical models used in this paper.
-Whereas one of the assumptions are homogeneous landscape.
-Nevertheless, the way economies are organised comes partly from the emergence of centralities within a city.
+Despite this unique spatial distribution of land the simple theoretical models used in this paper the results show similar patterns.
+Nevertheless, one of the assumptions are homogeneous landscape.
+However, the way economies are organised comes partly from the emergence of centralities within a city.
 Transportation networks have the potential to increase sprawl depending on how the urban system evolves gradually.
@@
 Resulting in finding relationships based on three different types of distance measurements.
-Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas and lastly amongst dentists.
+Firstly, the distance between Dental Practice and CBD then, dentist and other retail agglomeration areas.
 Finding that most dentist are either located close to a City, Central Business District or a shopping mall.
 Additionally, dentist also located close to each other making some sort of industrial cluster.
-Supporting, Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
+Furthermore, supporting Marshall's agglomeration theory on knowledge spillovers, increasing the sills to attract more patients.
 
-Furthermore, when using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
-Furthermore, decreasing as the distance gets bigger.
+When using the dentist income (y) and comparing the distance from the mall (x), signs of patterns emerge showing that income is greater when the dentist is placed closer to the shopping mall.
+Furthermore, decreasing as the distance increase.
 Faster transportation can increase convenience of commuting, increasing rent commuters from the city centre are willing to pay, and therefore increasing the area of developed land without this being necessary corresponding to an increase in population.
@@
 Our findings in QGIS shows that in areas of high population density, the number of dentist also increase.
-The consumer/patient then has more to choose between and the dentist also has a bigger scale of potential customers.
+The consumer/patient then has more dentists to choose from and the dentist also has a larger scale of potential customers.
 Where in smaller areas it is less competition and the dentist would appear to have somewhat of a monopoly since the transportation cost to the second closes dentist are noticeable higher.
@@
 This also takes away some of the competition and would support placement in more rural areas.
-More-so, is important to notice that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
-Suggesting some dentist might seek closer to CBD areas to capture most of the remaining part of the population.
+Dispute this the findings still support the Marshallian agglomeration attributes and bid-rent placement theory of central location.
+Nevertheless, some dentist seek closer to CBD areas to capture most of the remaining part of the population.
 Business, public places and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
@@
 
-Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
-Tjener tannleger i større byer mer?
-Her kan vi også skrive om gravitasjons modeller.
-
 Another Mashiallian point to be made could be the advantage of industry clusters.
@@
 Shopping malls have a relatively short history in Norway, with the first shopping mall opening in the late sixties.
-Prior to the most retail activities took place in the city centre.
+Prior to this most retail activities took place in the city centre.
 Furthermore, these malls typically featured a mix of shops, restaurants and entertainment options, and they were often located in suburban areas or on the outskirts of cities @home-n.
@@
 This regional economic approach aim to promote sustainable development and balance economic growth across the country @kanvin2014.
-
-Hva viser informasjonen om vinmonopolet?
 

### 2023-02-20 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
wrong document

--- 2023-02-19 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
@@
 
-According to the Alons's model dentist should locate in the retail area for higher volume of sales.
+The size of the mall also has a positive effect on revenue for dentist practice, as larger malls tends to have have positive effect on the revenue.
+Larger malls gives the customer more options to choose from and are often placed closer to a larger CBD which might be more attractive for customers to visit, and then also explaining why customers could be more interested in traveling further to visit these malls instead of visiting the local mall in a smaller CBD.
+
+More so, the reasoning for a more profitable business close to a shopping malls can be explain by customers beeing able to combine their dentis visit with shopping afterwards, maybe in a way to threat them self.
+You can find some similiarities in marketing strategies for grocery stores in the sense where they alwyas have a well planned system of where all their products are.
+Where for example gums and grab and go snacks, will always be placed next to the cashier.
+these are usually cheap, small and quick to consume, and something that you dont usually look for, but since you are already there, you might aswell.
+Same thoughts goes for people who visit the dentists in or fairly close to a mall.
+they are already there and might aswell be done with their shopping.
+This tradeoff can be highly beneficial for the mall itself, considering the patients going to the dentists then also become customers of the mall by default.
+
+\
+According to the Alons's model dentist should locate close to the CBD for higher volume of sales.
 Dentist being placed further out in to other retail agglomeration areas can also be a good strategy.
@@
 Bertrand model also take price as competitive advantage in to consideration.
-
-Viser data noe om indikasjon på dyrere priser i nærheten av større byer of hvor det er stedsbestemt monopol?
-Tjener tannleger i større byer mer?
-Her kan vi også skrive om gravitasjons modeller.
 
@@
 
-Hva viser informasjonen om vinmonopolet?
+Findings around Wine monopoly shows a negative effect on turnover per capita in retail.
+In other words, the Wine monopoly has a bad influence for the turnover in retail.
+The explanation could be as mentioned, since Wine monopoly could be categorized as an anchor store, it drag customers from malls and areas where there is no winemonoply over to the areas and malls where a wine monopoly exist, only helping out the stores surrounding.
+Another reasoning could be that the winemonopoly drags attention away from the other shopping options, resulting in a negative effect for the turnover for other retailers.
+
+is important to notice that the municipal council is more or less in charge of where Wine monopoly is allowed to operate.
+There are several criterias and factors that are taken into consideration before anything gets approved to prevent areas to dry out.
+[@kvikstadSeListenDisse2019]
 
@@
 
+There are several critical factors that effect the income for dental practice where our result shows that the income tends to be higher located close to a mall or inside the mall itself.
+A larger mall also has a positive effect of the income.
+Theorys crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice and why there are certain areas where there is a cluster of dentist.
+However, it is important to notice that a high revenue does not equal good economic results as more populated and central areas tends to have higher fees for example rent.
+The Wine monopoly also appears to have negative effect on turnover per capita in retail.
+
+##
+
 ## References
+
+::: {#ref}
+:::
 

### 2023-02-20 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd

--- 2023-02-20 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
@@
 
+We furthermore have used plots to look at the relationship between the income of dentist's businesses and the distance from CBDs.
+The scatterplot in @fig-2 have some extreme figures of 1 billion+ income which we choose to ignore and deem as outliers, as we notice that these observations are accounting figures for an entire group of companies under the same legal name, and not individual companies.
+By ignoring these outliers we are then left with the scatterplot in @fig-3.
+This plot shows a tendency of a negative relationship between the variables, which might suggest that dentists might earn more closer to CBDs.
+The last scatterplots we have chosen to calculate is concerning the relationship of dentists EBIT which is earnings before interest and taxes, compared to their distances to the nearest shopping mall in @fig-4.
+Just as the previous scatterplot, we also see here a tendency of a negative relationship between the variables.
+
 Furthermore, plots were used to look at the relationship between the income of dentist's businesses and the distance from CBDs.
@@
 Just as the previous scatter-plot, we also see here a tendency of a negative relationship between the variables.
+
 A decent number of dentists run at a loss, having negative earnings before interest and taxes.
@@
 
+Findings around Wine monopoly shows a negative effect on turnover per capita in retail.
+In other words, the Wine monopoly has a bad influence for the turnover in retail.
+The explanation could be as mentioned, since Wine monopoly could be categorized as an anchor store, it drag customers from malls and areas where there is no winemonoply over to the areas and malls where a wine monopoly exist, only helping out the stores surrounding.
+Another reasoning could be that the winemonopoly drags attention away from the other shopping options, resulting in a negative effect for the turnover for other retailers.
+
+is important to notice that the municipal council is more or less in charge of where Wine monopoly is allowed to operate.
+There are several criterias and factors that are taken into consideration before anything gets approved to prevent areas to dry out.
+[@kvikstadSeListenDisse2019]
+
 ```{r, echo=FALSE}

### 2023-02-20 Retail Agglomeration_abstract_introduction_theori_econometric_approach.qmd
idk

@@
 A larger mall also has a positive effect of the income.
-Theorys crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice and why there are certain areas where there is a cluster of dentist.
+Theory´s crated by Marshall, Alonso and Keynesian helps us explain some of our findings on why malls shows a pattern to be a better location for dental practice and why there are certain areas where there is a cluster of dentist.
 However, it is important to notice that a high revenue does not equal good economic results as more populated and central areas tends to have higher fees for example rent.
